### 2015-04-14 paper.tex
Struttura articolo

\documentclass{llncs}
%
\begin{document}
\title{Tampering detection for low-power smart camera}

\author{Author 1, Author 2}

\institute{Politecnico di Milano}
\maketitle

\begin{abstract}

	Abstract is here.

\end{abstract}

\section{Introduction}\label{sec:introduction}

Introduction is here.

\section{Related works}\label{sec:relWorks}

Other sections are here.

\section{Problem formulation}\label{sec:probForm}

\section{Proposed solution}\label{sec:propSol}

\section{Experiments}\label{sec:experiments}

\section{Conclusion}\label{sec:Conclusion}

Conclusions are here.

\section*{Acknowledgments}\label{sec:Acknowledgments}

Authors would like to thank YYYYY.

\bibliography{bibl_tesi}

%\begin{thebibliography}{1}
%	
%	\bibitem{Einstein}
%	A. Einstein, On the movement of small particles suspended in stationary liquids required by the molecular-kinetic theory of heat, Annalen der Physik 17, pp. 549-560, 1905.
%	
%\end{thebibliography}
\end{document}

### 2015-05-14 paper.tex
Aggiunta bibliografia
Inserite immagini in pdf
Inserite affiliations autori

@@
 \documentclass{llncs}
-%
+
+\usepackage{amssymb}
+\setcounter{tocdepth}{3}
+\usepackage{graphicx}
+\usepackage[ruled]{algorithm2e}
+%\usepackage[lined,boxed,commentsnumbered]{algorithm2e}
+\usepackage{amssymb}
+\usepackage{amsmath,graphicx,color,doi}
+\usepackage{algorithm2e}
+%\usepackage{algorithmic}	
+\usepackage{multirow}
+\usepackage{subfigure}
+\usepackage{cite}
+
+\newcommand{\gi}[1]{{\textcolor{red}{[\small \textbf{Giacomo}: #1]}}}
+\newcommand{\ad}[1]{{\textcolor{red}{[\small \textbf{Adriano}: #1]}}}
+\newcommand{\cl}[1]{{\textcolor{red}{[\small \textbf{Claudio}: #1]}}}
+
 \begin{document}
@@
 
-\author{Author 1, Author 2}
+\author{Adriano Gaibotti\inst{1} \and Claudio Marchisio\inst{1} \and Alexandro Sentinelli\inst{1} \and Giacomo Boracchi\inst{2}}
 
-\institute{Politecnico di Milano}
+\institute{ STMicroelectronics, Advanced System Technology, Via Camillo Olivetti 2, 20864, Agrate Brianza (MB), Italy\\
+	\email{\{adriano.gaibotti, claudio.marchisio, alexandro.sentinelli\}@st.com}
+	\and
+	Politecnico di Milano, Dipartimento di Elettronica, Informazione e Bioingegneria (DEIB), Via Ponzio 34/5, 20133, Milano (MI), Italy\\
+	\email{giacomo.boracchi@polimi.it}}
 \maketitle
@@
 \begin{abstract}
-
-	Abstract is here.
-
+\gi{Adriano: riesci a mettere bene gli autori con le affiliation distinte per bene? Poi domani sento Alex per capire che vuole fare}
 \end{abstract}
@@
 
-Introduction is here.
+\gi{Adriano: metti un paio di foto per spiegare il problema che consideriamo.}
 
-\section{Related works}\label{sec:relWorks}
+Random Toughs:
+\begin{itemize}
+\item Smart cameras, low-power monitoring of scene. Description of the application scenario. Low frame rate.
+\item Cameras organized in a multimedia network, continuous acquisition and streaming is not feasible
+\item The problem of false alarms, radio module activation
+\item Other tampering attacks like obfuscation (??) which might be due to environmental phenomena such as rain, fog and mist over the camera lenses have to be detected by image analysis methods
+\item Displacement can be perceived by MEMS as well but these device alone are prone to false alarms. Visual inspection is necessary to reduce false alarms
+\item constrained environment: algorithms have to operate with a low computational complexity and memory requirement
+\end{itemize}
 
-Other sections are here.
+\section{Related Works}\label{sec:relWorks}
 
-\section{Problem formulation}\label{sec:probForm}
+\gi{Adriano: metti tutte le reference, ciascuna con un commento di una frase per dire che fa ed una frase (o mezza) per dire i problemi che ha (con particolare riferimento all'ambito low-power) Poi le aggiustiamo in maniera organica}
 
-\section{Proposed solution}\label{sec:propSol}
+\section{Problem Formulation}\label{sec:probForm}
+
+\gi{Adriano: metti le formule di quello della tesi circa displacemente e out of focus. Poi condensiamo il tutto}
+
+\section{Proposed Solution}\label{sec:propSol}
+
+\subsection{Scene Segmentation}\label{subsec:Segmentation}
+
+\subsection{Indicators}\label{subsec:Indicators}
+\gi{Adriano: mettere formule degli indicatori e anche del frame difference qua}
+
+\subsection{Outlier Detection}\label{subsec:MonitoringScheme}
+
+\subsection{Algorithm Summary}\label{subsec:AlgorithmSummary}
+
+\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
 
@@
 
+\subsection{Dataset Description}\label{subsec:Dataset}
+\gi{Adriano: Prova a mettere qua le info}
+
+\subsection{Alternative Approaches}\label{subsec:AlternativeApproach	es}
+\begin{itemize}
+\item Full
+\item Adaptive Region
+\item Voronoi Regions
+\end{itemize}
+
+\subsection{Performance Assessment}
+\gi{Adriano: Dire come vengono calcolate le ROC curves TPR e FPR, le cifre di merito insomma, spiegando bene che parametro varia}
+
+\gi{Adriano: metti entrambe le ROC curves, affiancate e per bene ed alcuni esempi di sequenze}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.7\linewidth]{Immagini/ROCdefocus_cropped.pdf}
+\caption{Defocus}
+\label{fig:ROCdefocus}
+\end{figure}
+
+\subsection{Discussion}
+\gi{Adiano: Aggiungi qua la complessit\'a computazionale}
+
 \section{Conclusion}\label{sec:Conclusion}
-
-Conclusions are here.
+\gi{Adriano: butta in inglese gli ongoing works (come ultima cosa)}
+\cite{aksay2007camera}.
 
 \section*{Acknowledgments}\label{sec:Acknowledgments}
+Authors would like to thank ST for supporting Adriano Gaibotti.
 
-Authors would like to thank YYYYY.
-
+\bibliographystyle{unsrt}
 \bibliography{bibl_tesi}

### 2015-05-14 paper.tex
Inserito abstract e algoritmi

@@
 \begin{abstract}
-\gi{Adriano: riesci a mettere bene gli autori con le affiliation distinte per bene? Poi domani sento Alex per capire che vuole fare}
+One of the most critical issues in smart cameras, used as nodes in wireless multimedia sensor networks (WMSN), is the automatic detection of events that could compromise the correct acquisition of the scene, such as the presence of water on the camera lens, or the displacement of the device, or tampering events.
+In general, techniques used to identify these type of events are called tampering detection algorithms.
+We introduce a solution for the tampering detection that can be used for low-power and embedded smart cameras, in which the acquisition is done with low frame rate, such as one frame every minute.
+
 \end{abstract}
@@
 \subsection{Algorithm Summary}\label{subsec:AlgorithmSummary}
+\begin{algorithm}[tp]
+	% \SetAlgoNoLine
+	\LinesNumbered
+	\SetAlgoNlRelativeSize{0}
+	\SetNlSty{small}{}{.}
+	\textbf{Configuration}:\\
+	\lnl{DEF-Tr0} Extract regions $\{R_k\}, k=1,\dots,K$  \\
+	\lnl{DEF-Tr1} \For{$t=1,\dots,T_{o}$}
+	{	\lnl{DEF-Tr2} Acquire frame $z_t$ \\
+		\lnl{DEF-Tr2a} \For{$k=1,\dots,K$}{
+			\lnl{DEF-Tr3a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
+		}
+		\lnl{DEF-Tr3} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
+	}
+	\lnl{DEF-Tr6a} \For{$k=1,\dots,K$}{
+		\lnl{DEF-Tr7a} Define thresholds $\Gamma_{min}^k$ and $\Gamma_{max}^k$\\
+	}
+	\lnl{DEF-Tr5} Define CDT parameters on $g(t)$ variance\\
+	\textbf{Operational phase}:\\
+	\lnl{DEF-Test1} \For{$t=T_{o},\dots,\infty$}{
+		\lnl{DEF-Test2} Acquire frame $z_t$ \\
+		\lnl{DEF-Tr3b} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
+		\lnl{DEF-Test2a} $n=0$\\
+		\lnl{DEF-Test4a} \For{$k=1,\dots,K$}{
+			\lnl{DEF-Test5a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
+			\lnl{DEF-Test6a} \If{$\frac{\partial g^k}{\partial t}(t) < \Gamma_{min}^k \vee \frac{\partial g^k}{\partial t}(t) > \Gamma_{max}^k $}{
+				\lnl{DEF-Test7a} $n=n+1$\\
+			}
+		}
+		\lnl{DEF-Test81} \If{$n \geq K-1$}{
+			\lnl{DEF-Test91} $z_t$ is a defocused frame\\
+		}
+		\lnl{DEF-Test8} \If{CDT detect a change on $g(t)$ variance}{
+			\lnl{DEF-Test9} $z_t$ is a defocused frame\\
+		}
+	}   
+	\caption{Blur detection algorithm}
+	\label{alg:DEFOCUS}
+\end{algorithm}
 
+\begin{algorithm}[tp]
+	% \SetAlgoNoLine
+	\LinesNumbered
+	\SetAlgoNlRelativeSize{0}
+	\SetNlSty{small}{}{.}
+	\textbf{Configuration}:\\
+	\lnl{DISPL-Tr1} Extract regions $\{R_k\}, k=1,\dots,K$  \\
+	\lnl{DISPL-Tr2} \For{$t=1,\dots,T_{o}$}
+	{	\lnl{DISPL-Tr3} Acquire frame $z_t$ \\
+		\lnl{DISPL-Tr4} \For{$k=1,\dots,K$}{
+			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\frac{\partial l^k}{\partial t}(t)$ for the region $R_k$\\
+		}
+	}
+	\lnl{DISPL-Tr6} \For{$k=1,\dots,K$}{
+		\lnl{DISPL-Tr7} Define thresholds $\Gamma_{min}^k$ and $\Gamma_{max}^k$\\
+	}
+	\textbf{Operational phase}:\\
+	\lnl{DISPL-Test1} \For{$t=T_{o},\dots,\infty$}{
+		\lnl{DISPL-Test2} Acquire frame $z_t$ \\
+		\lnl{DISPL-Test3} $n = 0$ \\
+		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
+			\lnl{DISPL-Test5} Compute $l^k(t)$, $\frac{\partial l^k}{\partial t}(t)$ for the region $R_k$\\
+			\lnl{DISPL-Test6} \If{$\frac{\partial l^k}{\partial t}(t) < \Gamma_{min}^k \vee \frac{\partial l^k}{\partial t}(t) > \Gamma_{max}^k $}{
+				\lnl{DISPL-Test7} $n=n+1$\\
+			}
+		}
+		\lnl{DISPL-Test8} \If{$n\geq K-1$}{
+			\lnl{DISPL-Test9} $z_t$ is a displaced frame\\
+		}
+	}   
+	\caption{Displacement detection algorithm}
+	\label{alg:DISPL}
+\end{algorithm}
 \gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}

### 2015-05-14 paper.tex
Inseriti grafici messo un po di soa e inseriti grafici

@@
 \begin{document}
-\title{Tampering detection for low-power smart camera}
+\title{Tampering detection for low-power smart cameras}
 
-\author{Adriano Gaibotti\inst{1} \and Claudio Marchisio\inst{1} \and Alexandro Sentinelli\inst{1} \and Giacomo Boracchi\inst{2}}
+\author{Adriano Gaibotti\inst{1}\inst{2} \and Claudio Marchisio\inst{1} \and Alexandro Sentinelli\inst{1} \and Giacomo Boracchi\inst{2}}
 
@@
 	Politecnico di Milano, Dipartimento di Elettronica, Informazione e Bioingegneria (DEIB), Via Ponzio 34/5, 20133, Milano (MI), Italy\\
-	\email{giacomo.boracchi@polimi.it}}
+	\email{adriano.gaibotti@mail.polimi.it, giacomo.boracchi@polimi.it}}
 \maketitle
@@
 We introduce a solution for the tampering detection that can be used for low-power and embedded smart cameras, in which the acquisition is done with low frame rate, such as one frame every minute.
+The solution consists in a segmentation of the acquired scene in adaptive regions, and monitoring some features in each region, in order to find a change in their behavior that could be associated to a tampering event.
 
@@
 \gi{Adriano: metti un paio di foto per spiegare il problema che consideriamo.}
+
+\begin{figure}
+\centering
+\subfigure[]{\label{fig:neve} \includegraphics[width=0.4\linewidth]{Immagini/neve}}
+\subfigure[]{\label{fig:pioggia} \includegraphics[width=0.4\linewidth]{Immagini/pioggia}}
+\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
+In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
+\label{fig:tampering}
+\end{figure}
 
@@
 \gi{Adriano: metti tutte le reference, ciascuna con un commento di una frase per dire che fa ed una frase (o mezza) per dire i problemi che ha (con particolare riferimento all'ambito low-power) Poi le aggiustiamo in maniera organica}
+
+\cite{aksay2007camera}: uses background subtraction methods in order to identify defocus and occlusions, doing a comparison in the wavelet domain for defocus detection and histogram comparison  for occlusion detection.
+Displacement is not treated.
+
+\cite{saglam2009real}:  uses background subtraction methods in order to identify defocus, occlusions, and displacements.
+Comparison in the Fourier domain for defocus detection, histogram comparison for occlusion detection, comparison between current background and delayed background for displacement detection.
+
+\cite{alippi2010detecting}: CUSUM CDT on gradient energy content in order to detect blur in frames
+
+\cite{gil2007automatic}:  uses background subtraction methods in order to identify defocus, occlusions, and displacements.
+Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
+
+\cite{ribnick2006real}: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
+
+\cite{kryjak2012fpga}: implementation in a FPGA of a solution based on background modeling, histograms comparisons, edges comparisons.
+
+\cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements.
+Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
+
+\cite{tsesmelis2013tamper}: monitoring of the number of key points extracted by SURF in order to detect defocus events, partition in blocks and HOG descriptors matching for each block in order to detect occlusions.
+These types of solutions requires a lot of computations
 
@@
 	\centering
-\includegraphics[width=0.7\linewidth]{Immagini/ROCdefocus_cropped.pdf}
-\caption{Defocus}
-\label{fig:ROCdefocus}
+	\subfigure[]{\label{fig:ROCdefocus} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus.pdf}}
+	\subfigure[]{\label{fig:ROCdefocusZOOM} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocusZOOM.pdf}}
+	\subfigure[]{\label{fig:ROCdispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement.pdf}}
+	\subfigure[]{\label{fig:ROCdisplZOOM} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacementZOOM.pdf}}
+	\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
+In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
+	\label{fig:ROC}
 \end{figure}
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.7\linewidth]{Immagini/ROCdefocus_cropped.pdf}
+%\caption{Defocus}
+%\label{fig:ROCdefocus}
+%\end{figure}
 
@@
 \gi{Adriano: butta in inglese gli ongoing works (come ultima cosa)}
-\cite{aksay2007camera}.
+.
 

### 2015-05-15 paper.tex
Aggiunto Formule

@@
 In general, techniques used to identify these type of events are called tampering detection algorithms.
-We introduce a solution for the tampering detection that can be used for low-power and embedded smart cameras, in which the acquisition is done with low frame rate, such as one frame every minute.
+We introduce a solution for the tampering detection issue which can be used for low-power and embedded smart cameras, in which the acquisition is done with low frame rate, such as one frame every minute.
 The solution consists in a segmentation of the acquired scene in adaptive regions, and monitoring some features in each region, in order to find a change in their behavior that could be associated to a tampering event.
-
+Experiments show that our solution increase the performances with respect to monitoring the whole scene, with low computational complexity.
+\keywords{tampering, defocus, displacement, segmentation, gradient}
 \end{abstract}
@@
 
+\begin{equation}
+\label{eq:blur_single}
+z(x)=\mathcal{D}[y](x) = \mathcal{B}[y](x) + \eta(x), \qquad x \in \mathcal{X}
+\end{equation}
+
+\begin{equation}
+\label{eq:blur}
+\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(x,s)ds,
+\end{equation}
+
+\begin{equation}
+\label{eq:blur_convolution}
+\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(s-x)ds,
+\end{equation}
+
+\begin{equation}
+\label{eq:blur_multi}
+z_t(x)=\mathcal{D}_t[y_t](x) = \mathcal{B}_t[y_t](x) + \eta(x), \qquad x \in \mathcal{X}.
+\end{equation}
+
+\begin{equation}
+\label{eq:displacement}
+z_t(x)  = \left\{ \begin{array}{rcl}
+y_t(x) + \eta(x) & \mbox{per} & t < T^* \\
+w_t(x) + \eta(x) & \mbox{per} & t \geqslant T^*
+\end{array}\right. ,
+\end{equation}
+
+Per $t<T^*$ abbiamo che i frame vengono acquisiti in condizioni di funzionamento normale:
+\[ z_t(x)=y_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t=1,\dots , T^*-1. \] 
+All'istante di tempo $t = T^*$ avviene un evento di tampering, il quale compromette i frame per $t\geq T^*$.  
+In particolare, nel caso di una sfocatura avremo
+\[ z_t = \mathcal{B}_t[y_t](x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*,\]
+mentre nel caso di uno spostamento della camera avremo
+\[ z_t = w_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*. \]
+
 \section{Proposed Solution}\label{sec:propSol}
@@
 
+\begin{equation}
+\label{eq:energyGradient}
+g(t) = \mathcal{G}[z_t] =\frac{\sum_{\mathcal{X}}\| \nabla z_t(x) \| _2^2 }{|\mathcal{X}|} ,
+\end{equation}
+
+In particolare, per il calcolo delle derivate orizzontali  abbiamo utilizzato il seguente filtro $f_h$:
+\[f_h = f \circledast \left[ \begin{array}{rcl}
+1 & 0 & -1
+\end{array}\right], \] 
+mentre per il calcolo delle derivate verticali abbiamo utilizzato il seguente filtro $f_v$:
+\[f_v = f \circledast \left[ \begin{array}{r}
+1 \\ 0 \\ -1
+\end{array}\right], \]
+dove abbiamo indicato con $\circledast$ l'operatore di convoluzione.
+Il filtro $f$, invece, \`e ottenuto tramite un campionamento della \textit{funzione gaussiana} $h$, con media $0$ e deviazione standard $\sigma$
+\begin{equation}
+\label{eq:gaussian}
+h(i,j)=\frac{1}{2\pi\sigma^2}\exp\left(-\frac{i^2+j^2}{2\sigma^2}\right),
+\end{equation}
+e ponendo il valore massimo di questa funzione nel centro del filtro.
+Con questi filtri \`e possibile calcolare la \textit{norma del gradiente} nel seguente modo:
+\begin{equation}
+\label{eq:normaGradiente}
+\| \nabla z_t(x) \|_2^2=\left(z_t \circledast f_h\right)(x)^2 + \left(z_t \circledast f_v\right)(x)^2.
+\end{equation}
+Una volta calcolata la norma del gradiente \`e possibile farne la media come specificato in \eqref{eq:energyGradient}.
+Il risultato finale \`e un indicatore \textit{scalare} per ciascun frame acquisito, che pu\`o essere monitorato per individuare eventi di sfocature. 
+In particolare ci aspettiamo che l'evento di sfocatura provochi un abbattimento del valore di $g$.
+
+\begin{equation}
+\label{eq:energyLuma}
+l(t) = \mathcal{L}[z_t] =\frac{\sum_{\mathcal{X}} z_t(x) }{|\mathcal{X}|} ,
+\end{equation}
+
+\begin{equation}
+\label{eq:gradientDetr}
+\frac{\partial g}{\partial t}(t) = g(t) - g(t-1),
+\end{equation}
+
+\begin{equation}
+\label{eq:lumaDetr}
+\frac{\partial l}{\partial t}(t) = l(t) - l(t-1).
+\end{equation}
+
 \subsection{Outlier Detection}\label{subsec:MonitoringScheme}
 
+\begin{equation}
+\label{eq:soglieGradiente}
+\begin{array}{lcl}
+\Gamma_{min}^k & = & \hat{\mu}_g^k -\gamma \hat{\sigma}_g^k\\
+\Gamma_{max}^k & = & \hat{\mu}_g^k + \gamma \hat{\sigma}_g^k
+\end{array},
+\end{equation}
+dove $\hat{\mu}_g^k$ indica il valore medio delle osservazioni del training set
+\begin{equation}
+\hat{\mu}_g^k = \frac{\sum_{\tau = 1}^{T_{o}} \frac{\partial g^k}{\partial t}(\tau)}{T_{o}}, \nonumber
+\end{equation}
+$\hat{\sigma}_g^k$ indica la deviazione standard delle osservazioni del training set
+\begin{equation}
+\hat{\sigma}_g^k  = \sqrt{\frac{1}{T_{o}-1}\sum_{\tau=1}^{T_{o}}\left(\frac{\partial g^k}{\partial t}(\tau) - \hat{\mu}_g^k(\tau)\right)^2} \nonumber
+\end{equation}
+e $\gamma>1$ \`e un parametro moltiplicativo ottenuto sperimentalmente.\\
+
+\begin{equation}
+\label{eq:soglieLuma}
+\begin{array}{rcl}
+\Gamma_{min}^k & = & \hat{\mu}_l^k -\gamma \hat{\sigma}_l^k\\
+\Gamma_{max}^k & = & \hat{\mu}_l^k + \gamma \hat{\sigma}_l^k
+\end{array},
+\end{equation}
+dove $\hat{\mu}_l^k$ indica il valore medio delle osservazioni del training set
+\begin{equation}
+\hat{\mu}_l^k = \frac{\sum_{\tau = 1}^{T_{o}} \frac{\partial l^k}{\partial t}(\tau)}{T_{o}}, \nonumber
+\end{equation}
+$\hat{\sigma}_l^k$ indica la deviazione standard delle osservazioni del training set
+\begin{equation}
+\hat{\sigma}_l^k  = \sqrt{\frac{1}{T_{o}-1}\sum_{\tau=1}^{T_{o}}\left(\frac{\partial l^k}{\partial t}(\tau) - \hat{\mu}_l^k(\tau)\right)^2} \nonumber
+\end{equation}
+e $\gamma>1$ \`e un parametro moltiplicativo ottenuto sperimentalmente.\\
+
 \subsection{Algorithm Summary}\label{subsec:AlgorithmSummary}
+
 \begin{algorithm}[tp]

### 2015-05-15 paper.tex
Aggiunto qualcosa...

@@
 
-\author{Adriano Gaibotti\inst{1}\inst{2} \and Claudio Marchisio\inst{1} \and Alexandro Sentinelli\inst{1} \and Giacomo Boracchi\inst{2}}
+\author{Adriano Gaibotti\inst{1}\inst{2} \and Claudio Marchisio\inst{2} \and Alexandro Sentinelli\inst{2} \and Giacomo Boracchi\inst{1}}
 
-\institute{ STMicroelectronics, Advanced System Technology, Via Camillo Olivetti 2, 20864, Agrate Brianza (MB), Italy\\
+\institute{ 
+	Politecnico di Milano, Dipartimento di Elettronica, Informazione e Bioingegneria (DEIB), Via Ponzio 34/5, 20133, Milano (MI), Italy\\
+	\email{adriano.gaibotti@mail.polimi.it, giacomo.boracchi@polimi.it}
+	\and
+	STMicroelectronics, Advanced System Technology, Via Camillo Olivetti 2, 20864, Agrate Brianza (MB), Italy\\
 	\email{\{adriano.gaibotti, claudio.marchisio, alexandro.sentinelli\}@st.com}
-	\and
-	Politecnico di Milano, Dipartimento di Elettronica, Informazione e Bioingegneria (DEIB), Via Ponzio 34/5, 20133, Milano (MI), Italy\\
-	\email{adriano.gaibotti@mail.polimi.it, giacomo.boracchi@polimi.it}}
+}
 \maketitle

### 2015-05-15 paper.tex
Fatto qualcosa...

@@
 \gi{Adriano: Prova a mettere qua le info}
-
-\subsection{Alternative Approaches}\label{subsec:AlternativeApproach	es}
+The proposed method have been tested on a dataset of frame sequences in which there were tampering events.
+A part of this dataset was taken using a Raspberry Pi with its camera module, and the tampering was introduced by moving the device or putting water on the camera.
+The other part of this dataset is made of frame sequences taken from web-cams available on the Internet. 
+In these sequences there were a few events of displacements or defocus, but to have a larger set of tampering events we
+\subsection{Alternative Approaches}\label{subsec:AlternativeApproaches}
 \begin{itemize}

### 2015-05-15 paper.tex
Pushato

@@
 \begin{equation}
-\label{eq:energyGradient}
-g(t) = \mathcal{G}[z_t] =\frac{\sum_{\mathcal{X}}\| \nabla z_t(x) \| _2^2 }{|\mathcal{X}|} ,
+\label{eq:gradientRegions}
+\begin{array}{ccc}
+g^k(t)&  = & \mathcal{G}^k[z_t] = \frac{\sum_{\mathcal{R}_k}\| \nabla z_t(x) \| _2^2 }{|{\mathcal{R}_k}|}\\
+\partial g^k(t) & =& g^k(t)-g^k(t-1) 
+\end{array},
 \end{equation}
@@
 In particolare ci aspettiamo che l'evento di sfocatura provochi un abbattimento del valore di $g$.
+
+\begin{equation}
+\label{eq:lumaRegions}
+\begin{array}{ccc}
+l^k(t)&  = & \mathcal{L}^k[z_t] = \frac{\sum_{\mathcal{R}_k} z_t(x) }{|{\mathcal{R}_k}|}\\
+\partial l^k(t) & =& l^k(t)-l^k(t-1) 
+\end{array},
+\end{equation}
 
@@
 
-\begin{algorithm}[tp]
-	% \SetAlgoNoLine
-	\LinesNumbered
-	\SetAlgoNlRelativeSize{0}
-	\SetNlSty{small}{}{.}
-	\textbf{Configuration}:\\
-	\lnl{DEF-Tr0} Extract regions $\{R_k\}, k=1,\dots,K$  \\
-	\lnl{DEF-Tr1} \For{$t=1,\dots,T_{o}$}
-	{	\lnl{DEF-Tr2} Acquire frame $z_t$ \\
-		\lnl{DEF-Tr2a} \For{$k=1,\dots,K$}{
-			\lnl{DEF-Tr3a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
-		}
-		\lnl{DEF-Tr3} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
-	}
-	\lnl{DEF-Tr6a} \For{$k=1,\dots,K$}{
-		\lnl{DEF-Tr7a} Define thresholds $\Gamma_{min}^k$ and $\Gamma_{max}^k$\\
-	}
-	\lnl{DEF-Tr5} Define CDT parameters on $g(t)$ variance\\
-	\textbf{Operational phase}:\\
-	\lnl{DEF-Test1} \For{$t=T_{o},\dots,\infty$}{
-		\lnl{DEF-Test2} Acquire frame $z_t$ \\
-		\lnl{DEF-Tr3b} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
-		\lnl{DEF-Test2a} $n=0$\\
-		\lnl{DEF-Test4a} \For{$k=1,\dots,K$}{
-			\lnl{DEF-Test5a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
-			\lnl{DEF-Test6a} \If{$\frac{\partial g^k}{\partial t}(t) < \Gamma_{min}^k \vee \frac{\partial g^k}{\partial t}(t) > \Gamma_{max}^k $}{
-				\lnl{DEF-Test7a} $n=n+1$\\
-			}
-		}
-		\lnl{DEF-Test81} \If{$n \geq K-1$}{
-			\lnl{DEF-Test91} $z_t$ is a defocused frame\\
-		}
-		\lnl{DEF-Test8} \If{CDT detect a change on $g(t)$ variance}{
-			\lnl{DEF-Test9} $z_t$ is a defocused frame\\
-		}
-	}   
-	\caption{Blur detection algorithm}
-	\label{alg:DEFOCUS}
-\end{algorithm}
+%\begin{algorithm}[tp]
+%	% \SetAlgoNoLine
+%	\LinesNumbered
+%	\SetAlgoNlRelativeSize{0}
+%	\SetNlSty{small}{}{.}
+%	\textbf{Configuration}:\\
+%	\lnl{DEF-Tr0} Extract regions $\{R_k\}, k=1,\dots,K$  \\
+%	\lnl{DEF-Tr1} \For{$t=1,\dots,T_{o}$}
+%	{	\lnl{DEF-Tr2} Acquire frame $z_t$ \\
+%		\lnl{DEF-Tr2a} \For{$k=1,\dots,K$}{
+%			\lnl{DEF-Tr3a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
+%		}
+%		\lnl{DEF-Tr3} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
+%	}
+%	\lnl{DEF-Tr6a} \For{$k=1,\dots,K$}{
+%		\lnl{DEF-Tr7a} Define thresholds $\Gamma_{min}^k$ and $\Gamma_{max}^k$\\
+%	}
+%	\lnl{DEF-Tr5} Define CDT parameters on $g(t)$ variance\\
+%	\textbf{Operational phase}:\\
+%	\lnl{DEF-Test1} \For{$t=T_{o},\dots,\infty$}{
+%		\lnl{DEF-Test2} Acquire frame $z_t$ \\
+%		\lnl{DEF-Tr3b} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
+%		\lnl{DEF-Test2a} $n=0$\\
+%		\lnl{DEF-Test4a} \For{$k=1,\dots,K$}{
+%			\lnl{DEF-Test5a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
+%			\lnl{DEF-Test6a} \If{$\frac{\partial g^k}{\partial t}(t) < \Gamma_{min}^k \vee \frac{\partial g^k}{\partial t}(t) > \Gamma_{max}^k $}{
+%				\lnl{DEF-Test7a} $n=n+1$\\
+%			}
+%		}
+%		\lnl{DEF-Test81} \If{$n \geq K-1$}{
+%			\lnl{DEF-Test91} $z_t$ is a defocused frame\\
+%		}
+%		\lnl{DEF-Test8} \If{CDT detect a change on $g(t)$ variance}{
+%			\lnl{DEF-Test9} $z_t$ is a defocused frame\\
+%		}
+%	}   
+%	\caption{Blur detection algorithm}
+%	\label{alg:DEFOCUS}
+%\end{algorithm}
 
@@
 	\SetNlSty{small}{}{.}
+	\textbf{Input}: $\gamma$, $T_{o}$, $\{\mathcal{R}_k\}, k=1,\dots,K$ \\
 	\textbf{Configuration}:\\
-	\lnl{DISPL-Tr1} Extract regions $\{R_k\}, k=1,\dots,K$  \\
+%	\lnl{DISPL-Tr1} Extract regions $\{\mathcal{R}_k\}, k=1,\dots,K$  \\
 	\lnl{DISPL-Tr2} \For{$t=1,\dots,T_{o}$}
-	{	\lnl{DISPL-Tr3} Acquire frame $z_t$ \\
+	{	\lnl{DISPL-Tr3} Get frame $z_t$ \\
 		\lnl{DISPL-Tr4} \For{$k=1,\dots,K$}{
-			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\frac{\partial l^k}{\partial t}(t)$ for the region $R_k$\\
+			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $\mathcal{R}_k$\\
 		}
@@
 	\lnl{DISPL-Tr6} \For{$k=1,\dots,K$}{
-		\lnl{DISPL-Tr7} Define thresholds $\Gamma_{min}^k$ and $\Gamma_{max}^k$\\
+		\lnl{DISPL-Tr7} Compute $\sigma_l^k$\\
 	}
@@
 	\lnl{DISPL-Test1} \For{$t=T_{o},\dots,\infty$}{
-		\lnl{DISPL-Test2} Acquire frame $z_t$ \\
-		\lnl{DISPL-Test3} $n = 0$ \\
+		\lnl{DISPL-Test2} Get frame $z_t$ \\
+		\lnl{DISPL-Test3} $n_l = 0$ \\
 		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
-			\lnl{DISPL-Test5} Compute $l^k(t)$, $\frac{\partial l^k}{\partial t}(t)$ for the region $R_k$\\
-			\lnl{DISPL-Test6} \If{$\frac{\partial l^k}{\partial t}(t) < \Gamma_{min}^k \vee \frac{\partial l^k}{\partial t}(t) > \Gamma_{max}^k $}{
-				\lnl{DISPL-Test7} $n=n+1$\\
+			\lnl{DISPL-Test5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $\mathcal{R}_k$\\
+			\lnl{DISPL-Test6} \If{$\partial l^k(t) < -\gamma \sigma_l^k \vee \partial l^k(t) > \gamma \sigma_l^k $}{
+				\lnl{DISPL-Test7} $n_l=n_l+1$\\
 			}
 		}
-		\lnl{DISPL-Test8} \If{$n\geq K-1$}{
-			\lnl{DISPL-Test9} $z_t$ is a displaced frame\\
+		\lnl{DISPL-Test8} \If{$n_l\geq K-1$}{
+			\lnl{DISPL-Test9} $z_t$ is a tampered frame\\
 		}
 	}   
-	\caption{Displacement detection algorithm}
+	\caption{Tampering detection algorithm}
 	\label{alg:DISPL}
@@
 	\centering
-	\subfigure[]{\label{fig:ROCdefocus} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus.pdf}}
+	\subfigure[]{\label{fig:ROCdisplZOOM} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacementZOOM.pdf}}
+%	\subfigure[]{\label{fig:ROCdefocus} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus.pdf}}
 	\subfigure[]{\label{fig:ROCdefocusZOOM} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocusZOOM.pdf}}
-	\subfigure[]{\label{fig:ROCdispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement.pdf}}
-	\subfigure[]{\label{fig:ROCdisplZOOM} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacementZOOM.pdf}}
+%	\subfigure[]{\label{fig:ROCdispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement.pdf}}
 	\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.

### 2015-05-18 paper.tex
abstract

@@
 \begin{document}
-\title{Tampering detection for low-power smart cameras}
+\title{Tampering Detection For Low-Power Smart Cameras}
 
@@
 \begin{abstract}
-One of the most critical issues in smart cameras, used as nodes in wireless multimedia sensor networks (WMSN), is the automatic detection of events that could compromise the correct acquisition of the scene, such as the presence of water on the camera lens, or the displacement of the device, or tampering events.
-In general, techniques used to identify these type of events are called tampering detection algorithms.
-We introduce a solution for the tampering detection issue which can be used for low-power and embedded smart cameras, in which the acquisition is done with low frame rate, such as one frame every minute.
-The solution consists in a segmentation of the acquired scene in adaptive regions, and monitoring some features in each region, in order to find a change in their behavior that could be associated to a tampering event.
-Experiments show that our solution increase the performances with respect to monitoring the whole scene, with low computational complexity.
-\keywords{tampering, defocus, displacement, segmentation, gradient}
+A desirable feature for smart cameras is the ability to autonomously detect any tampering event that would prevent a clean view over the monitored scene.
+No matter whether tampering is due to atmospheric phenomena (e.g., few water drops over the camera lens) or to malicious events (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures.
+Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at video-like frame-rates, nor use sophisticated image-analysis algorithms.
+
+We here introduce a tampering-detection algorithm that has been specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme.
+Any frame yielding anomalous values of the indicators are detected as a tampering events.
+Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection performance for camera displacements.
+Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one per minute, and at a very small computational complexity.
+%and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
+
+\keywords{tampering detection, defocus, displacement detection}
 \end{abstract}
@@
 
-\gi{Adriano: metti un paio di foto per spiegare il problema che consideriamo.}
+One of the most critical issues in smart cameras, used as nodes in wireless multimedia sensor networks (WMSN), is the automatic detection of events that could
+
+In general, techniques used to identify these type of events are called tampering detection algorithms.
+
+A desirable feature for smart cameras is the ability to autonomously detect any tampering event that could impair a clean view over the monitored scene.
+No matter whether tampering events are due to atmospheric phenomena (e.g., water drops over the camera lens) or malicious attacks (e.g.
+displacement of the device), these have to be promptly detected and reported.
 
@@
 
-\gi{Adriano: metti le formule di quello della tesi circa displacemente e out of focus. Poi condensiamo il tutto}
+Let us denote by $z_t$ the frame that acquired at time $t$, which we model as
+\begin{equation}
+\label{eq:observationModel}
+z_t(x)=\mathcal{D}_t[y_t](x) \forall x \in X
+\end{equation}
+where $\mathcal{D}$ denotes the degradation operator, which transforms the original frame $y_t$ in our observation; $X \subset \mathbb{Z}$ denotes the regular pixels grid and $x$ is a vector indicating the pixel coordinates.
+When there is no tampering, the frame
+\begin{equation}
+\label{eq:no tampering}
+\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x) \forall x \in X
+\end{equation}
+where $\eta_t$ is a random variable accounting for all the noise sources (thermal, quantization, photon-counting).
 
+However, when the camera 
 \begin{equation}
-\label{eq:blur_single}
-z(x)=\mathcal{D}[y](x) = \mathcal{B}[y](x) + \eta(x), \qquad x \in \mathcal{X}
+\label{eq:defocus}
+\mathcal{D}_t[y_t](x) = \mathcal{B}_t[y_t](x) + \eta_t(x) \forall x \in X
 \end{equation}

### 2015-05-18 paper.tex
Aggiunte formule feature vector e immagini

--- 2015-05-15 paper.tex
@@
 \subsection{Scene Segmentation}\label{subsec:Segmentation}
-
+Consideriamo la sequenza $\{z_t\}$ di frame acquisiti dalla camera, con $t=1,\dots,T_{c}$.
+Per ciascun pixel $x\in\mathcal{X}$ calcoliamo un vettore $\textbf{d}(x)$ di $5$ elementi
+\begin{equation}
+\label{eq:featureVector}
+\textbf{d}(x)=\left[r(x);c(x);\mu_{\nabla}(x);\sigma_{\nabla}(x);\bar{z}(x)\right], \textbf{d}(x) \in \mathbb{R}^5
+\end{equation}
+dove:
+\begin{itemize}
+	\item $r(x)$ rappresenta il numero di riga del pixel $x$.
+	\item $c(x)$ rappresenta il numero di colonna del pixel $x$.
+	\item $\mu_{\nabla}(x)$ rappresenta il valore del gradiente nel pixel $x$ mediato nel tempo:
+	\begin{equation}
+	\label{eq:segmentazioneGrad}
+	\mu_{\nabla}(x) = \frac{\sum_{t=1}^{T_c}(\|\nabla z_t\|_2^2 \circledast f)(x)}{T_c},
+	\end{equation}
+	dove abbiamo indicato con $\|\nabla z_t\|_2^2$ la norma del gradiente per l'immagine $z_t$, definita in \eqref{eq:normaGradiente}, e con $f$ il filtro gaussiano discreto derivato dal campionamento di \eqref{eq:gaussian}.
+	\item $\sigma_{\nabla}(x)$ rappresenta la deviazione standard nel tempo del gradiente nel pixel $x$:
+	\begin{equation}
+	\label{eq:segmentazioneVar}
+	\sigma_{\nabla}(x)=\sqrt{\frac{1}{T_c - 1}\sum_{t=1}^{T_c}\left(\left(\|\nabla z_t\|_2^2 \circledast f\right)(x)-\mu_{\nabla}(x)\right)^2}.
+	\end{equation}
+	\item $\bar{z}(x)$ rappresenta il valore della luma del pixel $x$ mediato nel tempo:
+	\begin{equation}
+	\label{eq:segmentazioneLuma}
+	\bar{z}(x)=\frac{\sum_{t=1}^{T_c}( z_t \circledast f)(x)}{T_c}.
+	\end{equation}
+\end{itemize}
 \subsection{Indicators}\label{subsec:Indicators}
@@
 \section{Experiments}\label{sec:experiments}
+\begin{figure}
+\centering
+\includegraphics[width=1\linewidth]{Immagini/buenosAires}
+\caption{}
+\label{fig:buenosAires}
+\end{figure}
+\begin{figure}
+\centering
+\includegraphics[width=1\linewidth]{Immagini/buenosAiresDef}
+\caption{}
+\label{fig:buenosAiresDef}
+\end{figure}
+\begin{figure}
+\centering
+\includegraphics[width=1\linewidth]{Immagini/fulvioTestiDispl}
+\caption{}
+\label{fig:fulvioTestiDispl}
+\end{figure}
 

### 2015-05-18 paper.tex
draft Intro

--- 2015-05-18 paper.tex
@@
 \begin{document}
-\title{Tampering Detection For Low-Power Smart Cameras}
+\title{Tampering Detection In Low-Power Smart Cameras}
 
@@
 \begin{abstract}
-A desirable feature for smart cameras is the ability to autonomously detect any tampering event that would prevent a clean view over the monitored scene.
-No matter whether tampering is due to atmospheric phenomena (e.g., few water drops over the camera lens) or to malicious events (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures.
+A desirable feature for smart cameras is the ability to autonomously detect any tampering event that would prevent a clear view over the monitored scene.
+No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious events (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures.
 Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at video-like frame-rates, nor use sophisticated image-analysis algorithms.
@@
 \section{Introduction}\label{sec:introduction}
+\gi{Adriano: decidi che email tenere, non tutte e due :)}
 
-One of the most critical issues in smart cameras, used as nodes in wireless multimedia sensor networks (WMSN), is the automatic detection of events that could
+%We here address the problem of detecting tampering events/attacks in camera devices, and in particular those preventing the proper acquisition of the monitored scene.
 
-In general, techniques used to identify these type of events are called tampering detection algorithms.
+% What is the problem?
+When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures (as in Fig.
+\ref{fig:pioggia}), or in partial occlusions of the scene (as in Fig.
+\ref{fig:neve}).
+Similarly, other intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures (surely for monitoring purposes).
+We refer to these events/attacks as tampering.
 
-A desirable feature for smart cameras is the ability to autonomously detect any tampering event that could impair a clean view over the monitored scene.
-No matter whether tampering events are due to atmospheric phenomena (e.g., water drops over the camera lens) or malicious attacks (e.g.
-displacement of the device), these have to be promptly detected and reported.
+In some cases, tampering is easy to detect, e.g., when the camera integrity is affected and the device goes out-of-order.
+However, in some other situations, the device might not be affected itself, even though tampering prevents the correct interpretation of the scene or of small important details (e.g., the identification of licence plates).
 
-\begin{figure}
+% Why is it interesting and important?
+The early detection of tampering events/attacks is clearly an essential feature for surveillance systems \gi{REFERENCES}, where cameras are expected to autonomously detect any tampering event/attack and promptly report alerts.
+Surveillance cameras are typically connected to the power supply and acquire and process images at normal frame-rates (e.g.
+around 30fps) \gi{REFERENCES?}.
+
+\gi{Qui ST pu\'o aggiungere qualche esempio di applicazione qui o menzionare dispositivi di riferimento (con tanto di link a datasheet).
+Magari aggiungendo che le batterie recenti permettono operativit\'a di due anni a qeusti regimi.}
+We here consider low-power cameras, which are battery powered and that have to operate at very low frame-rates (e.g.
+one frame every minute or more) and that do not have much computational power or memory availability.
+As an example, consider wireless multimedia sensor networks (WMSN) \cite{akyildiz2007survey} where nodes are smart cameras performing very simple operations, wirelessy connected to other devices such that they can acquire and transmit images at regular time interval or upon requests.
+In these WMSN, notes are often not connected to the power supply and have to operate battery-powered, possibly implementing energy harvesting mechanisms \gi{REFERENCES, magari riprescando quelle del paper di TIM}.
+Even though these devices are not employed in critical surveillance applications, low-power smart cameras are becoming popular thanks to their low price which makes them suited for wide monitoring campaign.
+Tampering detection is therefore very important also in low-power smart cameras, because tampered frames are not informative and as such should not be transmitted over the network, also because operations like radio-module activation are quite energy demanding.
+Thus, low-power smart cameras have to promptly report tampering attack without raising many false alerts since these lead unnecessary data-transmissions and to energy consumption.
+
+% Why is it hard? (E.g., why do naive approaches fail?)
+Tampering detection in low-power smart cameras is much more challenging than in conventional surveillance cameras.
+Beside computational aspects -- such as the number of operations per pixels allowed and the memory availability -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), and the acquired sequence does not evolve smoothly.
+This prevents the use of learned background models and the analysis of foreground variations \gi{REFERENCE su BKGR SUBTRACTION}.
+When dynamic environments are acquired at low frame rates, like the example depicted in Fig.
+\ref{fig:buenosAires}, two consecutive frames might depict very different scenes because of changes in the depicted content and in the light conditions: smart cameras have to correctly distinguish between these \emph{normal} changes and changes due to camera displacement or blur/defocus.
+
+% We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose.
+Tampering events can be due to malicious attacks, as for instance a displacement of the camera or spraying some dirty liquid that would prevent the acquisition of (part of) the monitored scene or the interpretation of the content (e.g., the identification of licence plates).
+
+%What are the key components of my approach and results? Also include any specific limitations.
+We here propose an algorithm to detect both camera defocus and of displacement, as stated in Section \ref{sec:probForm}.
+The algorithm relies on simple indicators that are computed with a low computational complexity (Section \ref{subsec:Indicators}) and leverages an outlier detection technique to detect frames yielding anomalous indicator values as related to a tampering event (Section \ref{subsec:MonitoringScheme}).
+In particular, we perform a preliminarily partitioning of the scene into nonoverlapping regions, and separately compute and monitore indicators over different regions.
+Scene partitioning is defined by clustering feature vectors gathering the indicators in different pixels, thus performing a very coarse segmentation of the image (Section \ref{subsec:Segmentation}).
+Regions are computed during the initial configuration of the algorithm and as such do not implies any computational overhead w.r.t.
+operating on the whole image.
+Our experiments in Section \ref{sec:experiments} show that regions substantially improve the detection of camera displacements, while are not beneficial for detecting blur/defocus.
+Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.
+
+\begin{figure}[t!]
 \centering
+\subfigure[]{\label{fig:pioggia} \includegraphics[width=0.4\linewidth]{Immagini/pioggia}}
 \subfigure[]{\label{fig:neve} \includegraphics[width=0.4\linewidth]{Immagini/neve}}
-\subfigure[]{\label{fig:pioggia} \includegraphics[width=0.4\linewidth]{Immagini/pioggia}}
 \caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
@@
 
-Random Toughs:
-\begin{itemize}
-\item Smart cameras, low-power monitoring of scene. Description of the application scenario. Low frame rate.
-\item Cameras organized in a multimedia network, continuous acquisition and streaming is not feasible
-\item The problem of false alarms, radio module activation
-\item Other tampering attacks like obfuscation (??) which might be due to environmental phenomena such as rain, fog and mist over the camera lenses have to be detected by image analysis methods
-\item Displacement can be perceived by MEMS as well but these device alone are prone to false alarms. Visual inspection is necessary to reduce false alarms
-\item constrained environment: algorithms have to operate with a low computational complexity and memory requirement
-\end{itemize}
-
-\section{Related Works}\label{sec:relWorks}
-
-\gi{Adriano: metti tutte le reference, ciascuna con un commento di una frase per dire che fa ed una frase (o mezza) per dire i problemi che ha (con particolare riferimento all'ambito low-power) Poi le aggiustiamo in maniera organica}
-
-\cite{aksay2007camera}: uses background subtraction methods in order to identify defocus and occlusions, doing a comparison in the wavelet domain for defocus detection and histogram comparison  for occlusion detection.
-Displacement is not treated.
-
-\cite{saglam2009real}:  uses background subtraction methods in order to identify defocus, occlusions, and displacements.
-Comparison in the Fourier domain for defocus detection, histogram comparison for occlusion detection, comparison between current background and delayed background for displacement detection.
-
-\cite{alippi2010detecting}: CUSUM CDT on gradient energy content in order to detect blur in frames
-
-\cite{gil2007automatic}:  uses background subtraction methods in order to identify defocus, occlusions, and displacements.
-Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
-
+\subsection{Related Works}\label{subsec:relWorks}
+% Why hasn't it been solved before? (Or, what's wrong with previous proposed solutions? How does mine differ?)
+The literature concerning tampering detection is mostly focused on video surveillance applications and operates at frame-rates around (15- 30 fps) \gi{Adriano, \'e vero?}.
+Background models are typically leveraged to identify defocus and occlusions; in particular  ~\cite{aksay2007camera} performs detect defocus by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered.
+A background-subtraction technique is employed in~\cite{saglam2009real} to identify defocus, occlusions, and displacements.
+\gi{FIXME Comparison in the Fourier domain for defocus detection, histogram comparison for occlusion detection, comparison between current background and delayed background for displacement detection}.
+Background subtraction is also used in \cite{gil2007automatic} to detect defocus, occlusions, and displacements.
+\gi{FIXME Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection}.
+In contrast, no background models are used in \cite{alippi2010detecting}, where a sequential monitoring scheme based on a change-detection test is employed to detect changes in the average gradient energy of each frame to detect defocus due to external disturbances on the camera lens.
 \cite{ribnick2006real}: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
-
 \cite{kryjak2012fpga}: implementation in a FPGA of a solution based on background modeling, histograms comparisons, edges comparisons.
-
 \cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements.
 Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
-
 \cite{tsesmelis2013tamper}: monitoring of the number of key points extracted by SURF in order to detect defocus events, partition in blocks and HOG descriptors matching for each block in order to detect occlusions.
@@
 
+Random Toughs:
+\begin{itemize}
+\item The problem of false alarms, radio module activation
+\item Other tampering attacks like obfuscation (??) which might be due to environmental phenomena such as rain, fog and mist over the camera lenses have to be detected by image analysis methods
+\item Displacement can be perceived by MEMS as well but these device alone are prone to false alarms. Visual inspection is necessary to reduce false alarms
+
+\end{itemize}
+
 \section*{Acknowledgments}\label{sec:Acknowledgments}

### 2015-05-18 paper.tex
problem formulation --draft

@@
 
-Let us denote by $z_t$ the frame that acquired at time $t$, which we model as
+Let us denote by $z_t$ the frame acquired at time $t$, which we model as
 \begin{equation}
 \label{eq:observationModel}
-z_t(x)=\mathcal{D}_t[y_t](x) \forall x \in X
+z_t(x)=\mathcal{D}_t[y_t](x) \quad \forall x \in X
 \end{equation}
-where $\mathcal{D}$ denotes the degradation operator, which transforms the original frame $y_t$ in our observation; $X \subset \mathbb{Z}$ denotes the regular pixels grid and $x$ is a vector indicating the pixel coordinates.
-When there is no tampering, the frame
+where $\mathcal{D}_t$ denotes the degradation operator that transforms the original image $y_t$ in the frame $z_t$ which we acquire; $X \subset \mathbb{Z}$ denotes the regular pixels grid and $x$ is a vector indicating the pixel coordinates.
+As far as there is no tampering attacks/events,
 \begin{equation}
-\label{eq:no tampering}
-\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x) \forall x \in X
+\label{eq:no_tampering}
+\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x) \quad \forall x \in X
 \end{equation}
 where $\eta_t$ is a random variable accounting for all the noise sources (thermal, quantization, photon-counting).
+In normal conditions, all the images $y_t$ (thus also the frames $z_t$) might be different in content but are acquired from the same viewpoint and the same camera orientation.
 
-However, when the camera 
+When at time $\tau^*$ an external disturbance over the camera lens introduces a defocus, the image $y_t$ is blurred by spatially variant operator:
 \begin{equation}
-\label{eq:defocus}
-\mathcal{D}_t[y_t](x) = \mathcal{B}_t[y_t](x) + \eta_t(x) \forall x \in X
+\label{eq:model_defocus}
+\mathcal{D}_t[y_t](x) = \int_{\mathcal{X}}y(s)h_t(x,s)ds\, + \eta_t(x) \quad \forall x \in X, t \leq \tau^*
 \end{equation}
+%where  $\mathcal{B}_t$ corresponds to a linear operator
+%\begin{equation}
+%\label{eq:blur}
+%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(x,s)ds\,,
+%\end{equation}
+where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.
 
+%\begin{equation}
+%\label{eq:blur_convolution}
+%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(s-x)ds,
+%\end{equation}
+%
+%\begin{equation}
+%\label{eq:blur_multi}
+%z_t(x)=\mathcal{D}_t[y_t](x) = \mathcal{B}_t[y_t](x) + \eta(x), \qquad x \in \mathcal{X}.
+%\end{equation}
+A camera displacement at frame $\tau^*$ is modeled as 
 \begin{equation}
-\label{eq:blur}
-\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(x,s)ds,
-\end{equation}
-
-\begin{equation}
-\label{eq:blur_convolution}
-\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(s-x)ds,
-\end{equation}
-
-\begin{equation}
-\label{eq:blur_multi}
-z_t(x)=\mathcal{D}_t[y_t](x) = \mathcal{B}_t[y_t](x) + \eta(x), \qquad x \in \mathcal{X}.
-\end{equation}
-
-\begin{equation}
-\label{eq:displacement}
+\label{eq:model_displacement}
 z_t(x)  = \left\{ \begin{array}{rcl}
@@
 \end{equation}
+where $w_t$ is acquired from a different viewpoint and/or camera orientation than $y_t$.
 
-Per $t<T^*$ abbiamo che i frame vengono acquisiti in condizioni di funzionamento normale:
-\[ z_t(x)=y_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t=1,\dots , T^*-1. \] 
-All'istante di tempo $t = T^*$ avviene un evento di tampering, il quale compromette i frame per $t\geq T^*$.  
-In particolare, nel caso di una sfocatura avremo
-\[ z_t = \mathcal{B}_t[y_t](x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*,\]
-mentre nel caso di uno spostamento della camera avremo
-\[ z_t = w_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*. \]
+The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ and detect frames at time $\tau^*$ that correspond to a tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement}.
+
+%Per $t<T^*$ abbiamo che i frame vengono acquisiti in condizioni di funzionamento normale:
+%\[ z_t(x)=y_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t=1,\dots , T^*-1. \] 
+%All'istante di tempo $t = T^*$ avviene un evento di tampering, il quale compromette i frame per $t\geq T^*$.  
+%In particolare, nel caso di una sfocatura avremo
+%\[ z_t = \mathcal{B}_t[y_t](x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*,\]
+%mentre nel caso di uno spostamento della camera avremo
+%\[ z_t = w_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*. \]
 

### 2015-05-19 paper.tex
experiments

@@
 A part of this dataset was taken using a Raspberry Pi with its camera module, and the tampering was introduced by moving the device or putting water on the camera.
-The other part of this dataset is made of frame sequences taken from web-cams available on the Internet. 
-In these sequences there were a few events of displacements or defocus, but to have a larger set of tampering events we
+The other part of this dataset is made of frame sequences taken from webcams available on the Internet. 
+In these sequences there were a few events of displacements or defocus, but to have a larger set of tampering events we have simulated them.
+In particular we have used gaussian filtering for simulating defocus events, and concatenation of similar sequences for simulating displacement events.\\
+
 \subsection{Alternative Approaches}\label{subsec:AlternativeApproaches}

### 2015-05-19 paper.tex
ripassato abstract e sez 1,2. Manca da sistemare related works

--- 2015-05-18 paper.tex
@@
 \begin{abstract}
-A desirable feature for smart cameras is the ability to autonomously detect any tampering event that would prevent a clear view over the monitored scene.
-No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious events (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures.
+A desirable feature for smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene.
+No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attack (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures.
 Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at video-like frame-rates, nor use sophisticated image-analysis algorithms.
@@
 We here introduce a tampering-detection algorithm that has been specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme.
-Any frame yielding anomalous values of the indicators are detected as a tampering events.
-Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection performance for camera displacements.
-Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one per minute, and at a very small computational complexity.
+Any frame yielding anomalous indicator is detected as a tampering attempt.
+Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements.
+Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, and at a very small computational complexity.
 %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
@@
 % What is the problem?
-When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures (as in Fig.
-\ref{fig:pioggia}), or in partial occlusions of the scene (as in Fig.
-\ref{fig:neve}).
-Similarly, other intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures (surely for monitoring purposes).
+When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as in Fig.
+\ref{fig:pioggia}, or in partial occlusions of the scene, as in Fig.
+\ref{fig:neve}.
+Similarly, other intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be surely useless for monitoring purposes.
 We refer to these events/attacks as tampering.
-
 In some cases, tampering is easy to detect, e.g., when the camera integrity is affected and the device goes out-of-order.
-However, in some other situations, the device might not be affected itself, even though tampering prevents the correct interpretation of the scene or of small important details (e.g., the identification of licence plates).
+However, in many other situations, namely when the device is not physically damaged but the correct interpretation of the scene or of small important details is prevented (e.g., the identification of licence plates), tampering detection is not straightforward and requires image-analysis algorithms.
 
 % Why is it interesting and important?
-The early detection of tampering events/attacks is clearly an essential feature for surveillance systems \gi{REFERENCES}, where cameras are expected to autonomously detect any tampering event/attack and promptly report alerts.
+The early detection of tampering events/attacks is clearly essential in surveillance systems \gi{Adriano, add REFERENCES}, where cameras are expected to autonomously detect any tampering event/attack and promptly report alerts.
 Surveillance cameras are typically connected to the power supply and acquire and process images at normal frame-rates (e.g.
-around 30fps) \gi{REFERENCES?}.
+around few frames per second).
+In these conditions, several algorithms have been presented in the literature \gi{Adriano add REFERENCES}.
 
 \gi{Qui ST pu\'o aggiungere qualche esempio di applicazione qui o menzionare dispositivi di riferimento (con tanto di link a datasheet).
-Magari aggiungendo che le batterie recenti permettono operativit\'a di due anni a qeusti regimi.}
+Magari aggiungendo che le batterie recenti permettono operativit\'a di due anni a questi regimi.}
 We here consider low-power cameras, which are battery powered and that have to operate at very low frame-rates (e.g.
-one frame every minute or more) and that do not have much computational power or memory availability.
-As an example, consider wireless multimedia sensor networks (WMSN) \cite{akyildiz2007survey} where nodes are smart cameras performing very simple operations, wirelessy connected to other devices such that they can acquire and transmit images at regular time interval or upon requests.
+one frame every minute or more) and which have a constrained computational power or memory availability.
+As an example, consider wireless multimedia sensor networks (WMSN) \cite{akyildiz2007survey} where nodes are wirelessy connected smart-cameras, that can acquire and transmit images at regular time interval or upon requests.
 In these WMSN, notes are often not connected to the power supply and have to operate battery-powered, possibly implementing energy harvesting mechanisms \gi{REFERENCES, magari riprescando quelle del paper di TIM}.
-Even though these devices are not employed in critical surveillance applications, low-power smart cameras are becoming popular thanks to their low price which makes them suited for wide monitoring campaign.
-Tampering detection is therefore very important also in low-power smart cameras, because tampered frames are not informative and as such should not be transmitted over the network, also because operations like radio-module activation are quite energy demanding.
-Thus, low-power smart cameras have to promptly report tampering attack without raising many false alerts since these lead unnecessary data-transmissions and to energy consumption.
+Even though these devices are not employed in critical surveillance applications, low-power smart cameras are becoming popular in distributed monitoring of wide environments due to their low price \gi{possiamo mettere qualche REFERENCE}.
+Tampering detection is very important in low-power smart cameras, also considering that tampered frames should not be transmitted over the network, saving energy-demanding operations like the activation of the radio module.
 
@@
 When dynamic environments are acquired at low frame rates, like the example depicted in Fig.
-\ref{fig:buenosAires}, two consecutive frames might depict very different scenes because of changes in the depicted content and in the light conditions: smart cameras have to correctly distinguish between these \emph{normal} changes and changes due to camera displacement or blur/defocus.
+\ref{fig:buenosAires}, two consecutive frames might be very different because of changes in the scene and in the light conditions: smart cameras have to correctly distinguish between these \emph{normal} changes and changes due to camera displacement or blur/defocus.
+Moreover, il low-power smart cameras, tampering has to be reliably detected because false alerts would lead to unnecessary data-transmissions thus increasing the energy consumption.
 
@@
 We here propose an algorithm to detect both camera defocus and of displacement, as stated in Section \ref{sec:probForm}.
-The algorithm relies on simple indicators that are computed with a low computational complexity (Section \ref{subsec:Indicators}) and leverages an outlier detection technique to detect frames yielding anomalous indicator values as related to a tampering event (Section \ref{subsec:MonitoringScheme}).
-In particular, we perform a preliminarily partitioning of the scene into nonoverlapping regions, and separately compute and monitore indicators over different regions.
+The algorithm relies on simple indicators that are computed with a low computational complexity (Section \ref{subsec:Indicators}) and leverages an outlier detection technique to detect frames yielding anomalous indicators as related to a tampering event (Section \ref{subsec:MonitoringScheme}).
+In particular, we perform a preliminarily partitioning of the scene into nonoverlapping regions, and separately compute and monitor indicators over different regions.
 Scene partitioning is defined by clustering feature vectors gathering the indicators in different pixels, thus performing a very coarse segmentation of the image (Section \ref{subsec:Segmentation}).
@@
 operating on the whole image.
-Our experiments in Section \ref{sec:experiments} show that regions substantially improve the detection of camera displacements, while are not beneficial for detecting blur/defocus.
+Our experiments in Section \ref{sec:experiments} show that regions substantially improve the detection of camera displacements,  while when monitoring indicators meant to detect blur/defocus, it is more convenient to analyze the whole scene.
 Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.
@@
 % Why hasn't it been solved before? (Or, what's wrong with previous proposed solutions? How does mine differ?)
-The literature concerning tampering detection is mostly focused on video surveillance applications and operates at frame-rates around (15- 30 fps) \gi{Adriano, \'e vero?}.
+The literature concerning tampering detection is mostly focused on video surveillance applications and operates at few frames per second \gi{Adriano, \'e vero?}.
 Background models are typically leveraged to identify defocus and occlusions; in particular  ~\cite{aksay2007camera} performs detect defocus by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered.
@@
 
-Let us denote by $z_t$ the frame acquired at time $t$, which we model as
+Let $z_t$ be the frame acquired at time $t$, which we model as
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{D}_t$ denotes the degradation operator that transforms the original image $y_t$ in the frame $z_t$ which we acquire; $X \subset \mathbb{Z}$ denotes the regular pixels grid and $x$ is a vector indicating the pixel coordinates.
+where $\mathcal{D}_t$ denotes the degradation operator that transforms the original image $y_t$ in the frame $z_t$; $X \subset \mathbb{Z}^2$ denotes the regular pixels grid and $x\in \mathbb{Z}^2$ indicates the pixel coordinates.
 As far as there is no tampering attacks/events,
@@
 \end{equation}
-where $\eta_t$ is a random variable accounting for all the noise sources (thermal, quantization, photon-counting).
-In normal conditions, all the images $y_t$ (thus also the frames $z_t$) might be different in content but are acquired from the same viewpoint and the same camera orientation.
+where $\eta_t$ is a random variable accounting for different noise sources (e.g., thermal, quantization, photon-counting).
+In normal conditions, all the images $y_t$ (thus also the frames $z_t$) might show different content but are acquired from the same viewpoint and the same camera orientation.
 
-When at time $\tau^*$ an external disturbance over the camera lens introduces a defocus, the image $y_t$ is blurred by spatially variant operator:
+When at time $\tau^*$ an external disturbance introduces \emph{blur/defocus}, the image $y_t$ is degraded by a spatially variant blur operator, and $z_t$ becomes
 \begin{equation}
 \label{eq:model_defocus}
-\mathcal{D}_t[y_t](x) = \int_{\mathcal{X}}y(s)h_t(x,s)ds\, + \eta_t(x) \quad \forall x \in X, t \leq \tau^*
+\mathcal{D}_t[y_t](x) = \int_{\mathcal{X}}y(s)h_t(x,s)ds\, + \eta_t(x) \quad \forall x \in X, t \geq \tau^*
 \end{equation}
@@
 where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.
-
 %\begin{equation}
@@
 %\end{equation}
-A camera displacement at frame $\tau^*$ is modeled as 
+A \emph{camera displacement} at frame $\tau^*$ is instead modeled as 
 \begin{equation}
@@
 \end{equation}
-where $w_t$ is acquired from a different viewpoint and/or camera orientation than $y_t$.
+where $w_t$ relates to a different viewpoint and/or camera orientation than $y_t$.
 
-The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ and detect frames at time $\tau^*$ that correspond to a tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement}.
+The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect time $\tau^*$ when any tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.
 

### 2015-05-19 paper.tex
Sistemato experiments (ancora da finire...)

--- 2015-05-19 paper.tex
@@
 
-\author{Adriano Gaibotti\inst{1}\inst{2} \and Claudio Marchisio\inst{2} \and Alexandro Sentinelli\inst{2} \and Giacomo Boracchi\inst{1}}
+\author{Adriano Gaibotti\inst{1} \and Claudio Marchisio\inst{1} \and Alexandro Sentinelli\inst{1} \and Giacomo Boracchi\inst{2}}
 
 \institute{ 
-	Politecnico di Milano, Dipartimento di Elettronica, Informazione e Bioingegneria (DEIB), Via Ponzio 34/5, 20133, Milano (MI), Italy\\
-	\email{adriano.gaibotti@mail.polimi.it, giacomo.boracchi@polimi.it}
-	\and
 	STMicroelectronics, Advanced System Technology, Via Camillo Olivetti 2, 20864, Agrate Brianza (MB), Italy\\
 	\email{\{adriano.gaibotti, claudio.marchisio, alexandro.sentinelli\}@st.com}
+	\and
+	Politecnico di Milano, Dipartimento di Elettronica, Informazione e Bioingegneria (DEIB), Via Ponzio 34/5, 20133, Milano (MI), Italy\\
+	\email{giacomo.boracchi@polimi.it}
 }
@@
 \section{Introduction}\label{sec:introduction}
-\gi{Adriano: decidi che email tenere, non tutte e due :)}
+%\gi{Adriano: decidi che email tenere, non tutte e due :)}
 
@@
 	\SetNlSty{small}{}{.}
-	\textbf{Input}: $\gamma$, $T_{o}$, $\{\mathcal{R}_k\}, k=1,\dots,K$ \\
+	\textbf{Input}: $\gamma$, $T_{o}$, $\{R_k\}, k=1,\dots,K$ \\
 	\textbf{Configuration}:\\
@@
 		\lnl{DISPL-Tr4} \For{$k=1,\dots,K$}{
-			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $\mathcal{R}_k$\\
+			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
 		}
@@
 		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
-			\lnl{DISPL-Test5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $\mathcal{R}_k$\\
+			\lnl{DISPL-Test5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
 			\lnl{DISPL-Test6} \If{$\partial l^k(t) < -\gamma \sigma_l^k \vee \partial l^k(t) > \gamma \sigma_l^k $}{
@@
 \end{algorithm}
-\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
+%\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
 
 \section{Experiments}\label{sec:experiments}
+The proposed algorithm has been implemented in MATLAB, and has been tested on two datasets.
+The first one refers to frame sequences taken from webcams recording some parts of cities (as frames illustrated in Figure \ref{fig:buenosAires}), where we have introduced some synthetically generated tampering events:
+defocus has been simulated using gaussian filters, while displacement has been created with concatenation of similar frame sequences.
 \begin{figure}
@@
 	\includegraphics[width=1\linewidth]{Immagini/buenosAires}
-\caption{}
+	\caption{Example of frame sequence taken from webcam. Frames have been acquired every $30$ seconds.}
 	\label{fig:buenosAires}
@@
 \centering
-\includegraphics[width=1\linewidth]{Immagini/buenosAiresDef}
-\caption{}
-\label{fig:buenosAiresDef}
+\subfigure[]{\label{fig:buenosAiresDef} \includegraphics[width=1\linewidth]{Immagini/buenosAiresDef}}
+\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=1\linewidth]{Immagini/buenosAiresDispl}}
+\caption{Examples of synthetically generated tampering events.
+\textbf{(a)} Defocus generated with gaussian filters on frames \textit{d} and \textit{e}.
+\textbf{(b)} Displacement created doing concatenation of similar sequences on frames \textit{d} and \textit{e}.}
+\label{fig:buenosAiresTamp}
 \end{figure}
+In some cases these sequences contained real tampering events, as in Figure \ref{fig:fulvioTestiDispl}.\\
 \begin{figure}
@@
 \includegraphics[width=1\linewidth]{Immagini/fulvioTestiDispl}
-\caption{}
+\caption{Example of sequence taken from webcam in which there is a real displacement (frame \textit{d})}
 \label{fig:fulvioTestiDispl}
 \end{figure}
-
-\subsection{Dataset Description}\label{subsec:Dataset}
-\gi{Adriano: Prova a mettere qua le info}
-The proposed method have been tested on a dataset of frame sequences in which there were tampering events.
-A part of this dataset was taken using a Raspberry Pi with its camera module, and the tampering was introduced by moving the device or putting water on the camera.
-The other part of this dataset is made of frame sequences taken from webcams available on the Internet. 
-In these sequences there were a few events of displacements or defocus, but to have a larger set of tampering events we have simulated them.
-In particular we have used gaussian filtering for simulating defocus events, and concatenation of similar sequences for simulating displacement events.\\
+The second dataset was taken using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or putting water on the camera.\\
+Four figures of merit have been suggested to assess the performance of the proposed algorithm:\\
+\begin{description}
+	\item[TP]  True positives. It measures the number of tampering events correctly detected by the algorithm.\\ 
+	\item[TN]  True negatives. It measures the number of frames without tampering that are not detected by the algorithm. \\ 
+	\item[FP]  False positives. It measures the number of tampering events erroneously detected by the algorithm.\\ 
+	\item[FN]  False negatives. It measures the number of tampering events not detected by the algorithm. \\ 
+\end{description} 
+These indicators are computed varying the parameter $\gamma$, which defines the thresholds for the one-shot monitoring. 
+This permits us to generate \textit{ROC curves}, where on the x-axis there is:
+\[1-\text{SPECIFICITY}_\gamma = 1-\frac{\text{TN}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}=\frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma},\]
+while on th y-axis there is:
+\[\text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
+The construction of these curves permits us to
 

### 2015-05-19 paper.tex

--- 2015-05-19 paper.tex
@@
 \includegraphics[width=1\linewidth]{Immagini/fulvioTestiDispl}
-\caption{}
+\caption{Example of sequence taken from webcam in which there is a real displacement (frame \textit{d})}
 \label{fig:fulvioTestiDispl}
 \end{figure}
+The second dataset was taken using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or putting water on the camera.\\
+Four figures of merit have been suggested to assess the performance of the proposed algorithm:\\
+\begin{description}
+	\item[TP]  True positives. It measures the number of tampering events correctly detected by the algorithm.\\ 
+	\item[TN]  True negatives. It measures the number of frames without tampering that are not detected by the algorithm. \\ 
+	\item[FP]  False positives. It measures the number of tampering events erroneously detected by the algorithm.\\ 
+	\item[FN]  False negatives. It measures the number of tampering events not detected by the algorithm. \\ 
+\end{description} 
+These indicators are computed varying the parameter $\gamma$, which defines the thresholds for the one-shot monitoring. 
+This permits us to generate \textit{ROC curves}, where on the x-axis there is:
+\[1-\text{SPECIFICITY}_\gamma = 1-\frac{\text{TN}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}=\frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma},\]
+while on th y-axis there is:
+\[\text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
+The construction of these curves permits us to
 
-\subsection{Dataset Description}\label{subsec:Dataset}
-\gi{Adriano: Prova a mettere qua le info}
-The proposed method have been tested on a dataset of frame sequences in which there were tampering events.
-A part of this dataset was taken using a Raspberry Pi with its camera module, and the tampering was introduced by moving the device or putting water on the camera.
-The other part of this dataset is made of frame sequences taken from web-cams available on the Internet. 
-In these sequences there were a few events of displacements or defocus, but to have a larger set of tampering events we
 \subsection{Alternative Approaches}\label{subsec:AlternativeApproaches}

### 2015-05-19 paper.tex
Aggiunte ROC curve
Esteso un po...

@@
 \begin{array}{ccc}
-g^k(t)&  = & \mathcal{G}^k[z_t] = \frac{\sum_{\mathcal{R}_k}\| \nabla z_t(x) \| _2^2 }{|{\mathcal{R}_k}|}\\
+g^k(t)&  = & \mathcal{G}^k[z_t] = \frac{\sum_{R_k}\| \nabla z_t(x) \| _2^2 }{|{R_k}|}\\
 \partial g^k(t) & =& g^k(t)-g^k(t-1) 
@@
 \begin{array}{ccc}
-l^k(t)&  = & \mathcal{L}^k[z_t] = \frac{\sum_{\mathcal{R}_k} z_t(x) }{|{\mathcal{R}_k}|}\\
+l^k(t)&  = & \mathcal{L}^k[z_t] = \frac{\sum_{R_k} z_t(x) }{|{R_k}|}\\
 \partial l^k(t) & =& l^k(t)-l^k(t-1) 
@@
 \[\text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
-The construction of these curves permits us to
+The construction of these curves permits us to make a comparison with respect to other methods.
+In particular the alternative approaches that we have considered working:
+\begin{itemize}
+\item considering the whole scene for the features computation;
+\item considering adaptive region as described in Section \ref{sec:propSol} for the feature computation;
+\item considering voronoi regions for the feature computation.
+\end{itemize}
 
 \subsection{Alternative Approaches}\label{subsec:AlternativeApproaches}
-\begin{itemize}
-\item Full
-\item Adaptive Region
-\item Voronoi Regions
-\end{itemize}
 
@@
 
+% \begin{figure}
+	% \centering
+	% \subfigure[]{\label{fig:ROCdisplZOOM} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacementZOOM.pdf}}
+% %	\subfigure[]{\label{fig:ROCdefocus} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus.pdf}}
+	% \subfigure[]{\label{fig:ROCdefocusZOOM} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocusZOOM.pdf}}
+% %	\subfigure[]{\label{fig:ROCdispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement.pdf}}
+	% \caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
+In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
+	% \label{fig:ROC}
+% \end{figure}
+
 \begin{figure}
 	\centering
-	\subfigure[]{\label{fig:ROCdisplZOOM} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacementZOOM.pdf}}
-%	\subfigure[]{\label{fig:ROCdefocus} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus.pdf}}
-	\subfigure[]{\label{fig:ROCdefocusZOOM} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocusZOOM.pdf}}
-%	\subfigure[]{\label{fig:ROCdispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement.pdf}}
+	\includegraphics[width=0.7\linewidth]{Immagini/ROCdisplacement_luma}
 	\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
 In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
-	\label{fig:ROC}
+	\label{fig:ROCdispl_luma}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{Immagini/ROCdisplacement_fd}
+	\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
+In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
+	\label{fig:ROCdispl_fd}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{Immagini/ROCdefocus}
+	\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
+In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
+	\label{fig:ROCdispl_luma}
 \end{figure}

### 2015-05-19 paper.tex
inserite note Claudio

--- 2015-05-19 paper.tex
@@
 
-\author{Adriano Gaibotti\inst{1} \and Claudio Marchisio\inst{1} \and Alexandro Sentinelli\inst{1} \and Giacomo Boracchi\inst{2}}
+\author{Adriano Gaibotti\inst{1}\inst{2} \and Claudio Marchisio\inst{2} \and Alexandro Sentinelli\inst{2} \and Giacomo Boracchi\inst{1}}
 
 \institute{ 
+	Politecnico di Milano, Dipartimento di Elettronica, Informazione e Bioingegneria (DEIB), Via Ponzio 34/5, 20133, Milano (MI), Italy\\
+	\email{adriano.gaibotti@mail.polimi.it, giacomo.boracchi@polimi.it}
+	\and
 	STMicroelectronics, Advanced System Technology, Via Camillo Olivetti 2, 20864, Agrate Brianza (MB), Italy\\
 	\email{\{adriano.gaibotti, claudio.marchisio, alexandro.sentinelli\}@st.com}
-	\and
-	Politecnico di Milano, Dipartimento di Elettronica, Informazione e Bioingegneria (DEIB), Via Ponzio 34/5, 20133, Milano (MI), Italy\\
-	\email{giacomo.boracchi@polimi.it}
 }
@@
 \section{Introduction}\label{sec:introduction}
-%\gi{Adriano: decidi che email tenere, non tutte e due :)}
+\gi{Adriano: decidi che email tenere, non tutte e due :)}
 
@@
 In some cases, tampering is easy to detect, e.g., when the camera integrity is affected and the device goes out-of-order.
-However, in many other situations, namely when the device is not physically damaged but the correct interpretation of the scene or of small important details is prevented (e.g., the identification of licence plates), tampering detection is not straightforward and requires image-analysis algorithms.
+However, in many other situations, namely when the device is not physically damaged,	 but the correct interpretation of the scene or of small important details is prevented (e.g., the identification of licence plates), tampering detection is not straightforward and requires image-analysis algorithms.
 
@@
 \gi{Qui ST pu\'o aggiungere qualche esempio di applicazione qui o menzionare dispositivi di riferimento (con tanto di link a datasheet).
-Magari aggiungendo che le batterie recenti permettono operativit\'a di due anni a questi regimi.}
-We here consider low-power cameras, which are battery powered and that have to operate at very low frame-rates (e.g.
+Magari aggiungendo che le batterie recenti permettono operativit\'a di due anni a questi regimi.}\cl{Verifico cosa possiamo dire del SecSoc.}
+This work expressly targets low-power and ultra-low-power smart cameras.
+Such devices typically operate at very low frame-rates (e.g.
+possibly less than one frame every minute), with constrained computational power and memory, and are battery powered (typically lasting one or two years).
+% We here consider low-power cameras, which are battery powered and that have to operate at very low frame-rates (e.g.
 one frame every minute or more) and which have a constrained computational power or memory availability.
-As an example, consider wireless multimedia sensor networks (WMSN) \cite{akyildiz2007survey} where nodes are wirelessy connected smart-cameras, that can acquire and transmit images at regular time interval or upon requests.
-In these WMSN, notes are often not connected to the power supply and have to operate battery-powered, possibly implementing energy harvesting mechanisms \gi{REFERENCES, magari riprescando quelle del paper di TIM}.
-Even though these devices are not employed in critical surveillance applications, low-power smart cameras are becoming popular in distributed monitoring of wide environments due to their low price \gi{possiamo mettere qualche REFERENCE}.
-Tampering detection is very important in low-power smart cameras, also considering that tampered frames should not be transmitted over the network, saving energy-demanding operations like the activation of the radio module.
+As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where nodes are wirelessy connected smart-cameras, that can acquire and transmit images at regular time interval or upon requests.
+Often, in WMSN, the units are not connected to the power supply and have to operate with batteries, possibly implementing energy harvesting mechanisms \gi{REFERENCES, magari riprescando quelle del paper di TIM}.
+Even though these devices are not employed in critical surveillance applications, low-power smart cameras are becoming popular in distributed monitoring of wide environments due to their low cost and maintenance requirements \gi{possiamo mettere qualche REFERENCE}.
+Tampering detection is thus very important in low-power smart cameras, also considering that tampered frames should be identified to avoid unnecessary energy-demanding operations like the local processing or radio activation for transmission over the network.
 
@@
 Tampering detection in low-power smart cameras is much more challenging than in conventional surveillance cameras.
-Beside computational aspects -- such as the number of operations per pixels allowed and the memory availability -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), and the acquired sequence does not evolve smoothly.
+Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), and the acquired sequence does not evolve smoothly.
 This prevents the use of learned background models and the analysis of foreground variations \gi{REFERENCE su BKGR SUBTRACTION}.
@@
 \ref{fig:buenosAires}, two consecutive frames might be very different because of changes in the scene and in the light conditions: smart cameras have to correctly distinguish between these \emph{normal} changes and changes due to camera displacement or blur/defocus.
-Moreover, il low-power smart cameras, tampering has to be reliably detected because false alerts would lead to unnecessary data-transmissions thus increasing the energy consumption.
+Moreover, in low-power smart cameras, tampering has to be reliably detected because false alerts would lead to unnecessary energy waste.
 
@@
 %What are the key components of my approach and results? Also include any specific limitations.
-We here propose an algorithm to detect both camera defocus and of displacement, as stated in Section \ref{sec:probForm}.
+This work presents an algorithm to detect both camera defocus and of displacement, as stated in Section \ref{sec:probForm}.
 The algorithm relies on simple indicators that are computed with a low computational complexity (Section \ref{subsec:Indicators}) and leverages an outlier detection technique to detect frames yielding anomalous indicators as related to a tampering event (Section \ref{subsec:MonitoringScheme}).
@@
 operating on the whole image.
-Our experiments in Section \ref{sec:experiments} show that regions substantially improve the detection of camera displacements,  while when monitoring indicators meant to detect blur/defocus, it is more convenient to analyze the whole scene.
+Our experiments in Section \ref{sec:experiments} show that regions substantially improve the detection of camera displacements, while when monitoring indicators meant to detect blur/defocus, it is more convenient to consider the whole scene at once.
 Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.
@@
 \caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
-In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
+In Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens, while in Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens.}
 \label{fig:tampering}
@@
 	\SetNlSty{small}{}{.}
-	\textbf{Input}: $\gamma$, $T_{o}$, $\{R_k\}, k=1,\dots,K$ \\
+	\textbf{Input}: $\gamma$, $T_{o}$, $\{\mathcal{R}_k\}, k=1,\dots,K$ \\
 	\textbf{Configuration}:\\
@@
 		\lnl{DISPL-Tr4} \For{$k=1,\dots,K$}{
-			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
+			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $\mathcal{R}_k$\\
 		}
@@
 		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
-			\lnl{DISPL-Test5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
+			\lnl{DISPL-Test5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $\mathcal{R}_k$\\
 			\lnl{DISPL-Test6} \If{$\partial l^k(t) < -\gamma \sigma_l^k \vee \partial l^k(t) > \gamma \sigma_l^k $}{
@@
 \end{algorithm}
-%\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
+\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
 
 \section{Experiments}\label{sec:experiments}
-The proposed algorithm has been implemented in MATLAB, and has been tested on two datasets.
-The first one refers to frame sequences taken from webcams recording some parts of cities (as frames illustrated in Figure \ref{fig:buenosAires}), where we have introduced some synthetically generated tampering events:
-defocus has been simulated using gaussian filters, while displacement has been created with concatenation of similar frame sequences.
 \begin{figure}
@@
 \includegraphics[width=1\linewidth]{Immagini/buenosAires}
-	\caption{Example of frame sequence taken from webcam. Frames have been acquired every $30$ seconds.}
+\caption{}
 \label{fig:buenosAires}
@@
 \centering
-\subfigure[]{\label{fig:buenosAiresDef} \includegraphics[width=1\linewidth]{Immagini/buenosAiresDef}}
-\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=1\linewidth]{Immagini/buenosAiresDispl}}
-\caption{Examples of synthetically generated tampering events.
-\textbf{(a)} Defocus generated with gaussian filters on frames \textit{d} and \textit{e}.
-\textbf{(b)} Displacement created doing concatenation of similar sequences on frames \textit{d} and \textit{e}.}
-\label{fig:buenosAiresTamp}
+\includegraphics[width=1\linewidth]{Immagini/buenosAiresDef}
+\caption{}
+\label{fig:buenosAiresDef}
 \end{figure}
-In some cases these sequences contained real tampering events, as in Figure \ref{fig:fulvioTestiDispl}.\\
 \begin{figure}
@@
 \includegraphics[width=1\linewidth]{Immagini/fulvioTestiDispl}
-\caption{Example of sequence taken from webcam in which there is a real displacement (frame \textit{d})}
+\caption{}
 \label{fig:fulvioTestiDispl}
 \end{figure}
-The second dataset was taken using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or putting water on the camera.\\
-Four figures of merit have been suggested to assess the performance of the proposed algorithm:\\
-\begin{description}
-	\item[TP]  True positives. It measures the number of tampering events correctly detected by the algorithm.\\ 
-	\item[TN]  True negatives. It measures the number of frames without tampering that are not detected by the algorithm. \\ 
-	\item[FP]  False positives. It measures the number of tampering events erroneously detected by the algorithm.\\ 
-	\item[FN]  False negatives. It measures the number of tampering events not detected by the algorithm. \\ 
-\end{description} 
-These indicators are computed varying the parameter $\gamma$, which defines the thresholds for the one-shot monitoring. 
-This permits us to generate \textit{ROC curves}, where on the x-axis there is:
-\[1-\text{SPECIFICITY}_\gamma = 1-\frac{\text{TN}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}=\frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma},\]
-while on th y-axis there is:
-\[\text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
-The construction of these curves permits us to
+
+\subsection{Dataset Description}\label{subsec:Dataset}
+\gi{Adriano: Prova a mettere qua le info}
+The proposed method have been tested on a dataset of frame sequences in which there were tampering events.
+A part of this dataset was taken using a Raspberry Pi with its camera module, and the tampering was introduced by moving the device or putting water on the camera.
+The other part of this dataset is made of frame sequences taken from webcams available on the Internet. 
+In these sequences there were a few events of displacements or defocus, but to have a larger set of tampering events we have simulated them.
+In particular we have used gaussian filtering for simulating defocus events, and concatenation of similar sequences for simulating displacement events.\\
 

### 2015-05-20 paper.tex
aggiunti nuove ROC

@@
 \begin{abstract}
-A desirable feature for smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene.
+A desirable feature for smart cameras is the ability to autonomously detect any tampering event/attack that would prevents a clear view over the monitored scene.
 No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attack (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures.
@@
 % Why is it interesting and important?
-The early detection of tampering events/attacks is clearly essential in surveillance systems \gi{Adriano, add REFERENCES}, where cameras are expected to autonomously detect any tampering event/attack and promptly report alerts.
+The early detection of tampering events/attacks is clearly essential in surveillance systems \gi{Adriano, add REFERENCES} \cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering event/attack and promptly report alerts.
 Surveillance cameras are typically connected to the power supply and acquire and process images at normal frame-rates (e.g.
@@
 % Why is it hard? (E.g., why do naive approaches fail?)
-Tampering detection in low-power smart cameras is much more challenging than in conventional surveillance cameras.
+Tampering detection in low-power smart cameras is much more challenging than in conventional surveillance cameras \cite{perrig2004security}.
 Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), and the acquired sequence does not evolve smoothly.
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{Immagini/ROCdefocus}
+	\includegraphics[width=0.7\linewidth]{Immagini/ROCdefocus1}
+	\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
+In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
+	\label{fig:ROCdispl_luma}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{Immagini/ROCdefocus_fd}
 	\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.

### 2015-05-21 paper.tex
Sistemate immagini

@@
 defocus has been simulated using gaussian filters, while displacement has been created with concatenation of similar frame sequences.
-\begin{figure}
-	\centering
-	\includegraphics[width=1\linewidth]{Immagini/buenosAires}
-	\caption{Example of frame sequence taken from webcam. Frames have been acquired every $30$ seconds.}
-	\label{fig:buenosAires}
-\end{figure}
-\begin{figure}
-\centering
-\subfigure[]{\label{fig:buenosAiresDef} \includegraphics[width=1\linewidth]{Immagini/buenosAiresDef}}
-\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=1\linewidth]{Immagini/buenosAiresDispl}}
-\caption{Examples of synthetically generated tampering events.
-\textbf{(a)} Defocus generated with gaussian filters on frames \textit{d} and \textit{e}.
-\textbf{(b)} Displacement created doing concatenation of similar sequences on frames \textit{d} and \textit{e}.}
-\label{fig:buenosAiresTamp}
-\end{figure}
+
 In some cases these sequences contained real tampering events, as in Figure \ref{fig:fulvioTestiDispl}.\\
@@
 \centering
-\includegraphics[width=1\linewidth]{Immagini/fulvioTestiDispl}
-\caption{Example of sequence taken from webcam in which there is a real displacement (frame \textit{d})}
-\label{fig:fulvioTestiDispl}
+\includegraphics[width=1\linewidth]{Immagini/sequenze}
+\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences; \textbf{(d)} real displacement event on 4-th frame.}
+\label{fig:sequences}
 \end{figure}
@@
 \end{itemize}
-
-\subsection{Alternative Approaches}\label{subsec:AlternativeApproaches}
-
-\subsection{Performance Assessment}
-\gi{Adriano: Dire come vengono calcolate le ROC curves TPR e FPR, le cifre di merito insomma, spiegando bene che parametro varia}
-
-\gi{Adriano: metti entrambe le ROC curves, affiancate e per bene ed alcuni esempi di sequenze}
-
-% \begin{figure}
-	% \centering
-	% \subfigure[]{\label{fig:ROCdisplZOOM} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacementZOOM.pdf}}
-% %	\subfigure[]{\label{fig:ROCdefocus} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus.pdf}}
-	% \subfigure[]{\label{fig:ROCdefocusZOOM} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocusZOOM.pdf}}
-% %	\subfigure[]{\label{fig:ROCdispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement.pdf}}
-	% \caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
-In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
-	% \label{fig:ROC}
-% \end{figure}
+%
+%% \subsection{Alternative Approaches}\label{subsec:AlternativeApproaches}
+%
+%
+%\subsection{Performance Assessment}
+%\gi{Adriano: Dire come vengono calcolate le ROC curves TPR e FPR, le cifre di merito insomma, spiegando bene che parametro varia}
+%
+%\gi{Adriano: metti entrambe le ROC curves, affiancate e per bene ed alcuni esempi di sequenze}
+Experimental results are shown in Figures \ref{fig:ROCdispl_luma}, \ref{fig:ROCdispl_fd}, \ref{fig:ROCdef_grad}, \ref{fig:ROCdef_fd}.
 
@@
 \centering
-	\includegraphics[width=0.7\linewidth]{Immagini/ROCdisplacement_luma}
-	\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
-In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
-	\label{fig:ROCdispl_luma}
+\subfigure[]{\label{fig:buenosAiresDef}\includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement_luma}}
+\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement_fd}}
+\caption{Examples of synthetically generated tampering events.
+\textbf{(a)} Defocus generated with gaussian filters on frames \textit{d} and \textit{e}.
+\textbf{(b)} Displacement created doing concatenation of similar sequences on frames \textit{d} and \textit{e}.}
+\label{fig:buenosAiresTamp}
 \end{figure}
@@
 \centering
-	\includegraphics[width=0.7\linewidth]{Immagini/ROCdisplacement_fd}
-	\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
-In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
-	\label{fig:ROCdispl_fd}
+\subfigure[]{\label{fig:buenosAiresDef}\includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus1}}
+\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus_fd}}
+\caption{Examples of synthetically generated tampering events.
+\textbf{(a)} Defocus generated with gaussian filters on frames \textit{d} and \textit{e}.
+\textbf{(b)} Displacement created doing concatenation of similar sequences on frames \textit{d} and \textit{e}.}
+\label{fig:buenosAiresTamp}
 \end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{Immagini/ROCdefocus1}
-	\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
-In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
-	\label{fig:ROCdispl_luma}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{Immagini/ROCdefocus_fd}
-	\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
-In Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens, while in Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens.}
-	\label{fig:ROCdispl_luma}
-\end{figure}
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.7\linewidth]{Immagini/ROCdefocus_cropped.pdf}
-%\caption{Defocus}
-%\label{fig:ROCdefocus}
-%\end{figure}
 

### 2015-05-21 paper.tex
Sistemato didascalie

@@
 defocus has been simulated using gaussian filters, while displacement has been created with concatenation of similar frame sequences.
-
 In some cases these sequences contained real tampering events, as in Figure \ref{fig:fulvioTestiDispl}.\\
@@
 %\gi{Adriano: metti entrambe le ROC curves, affiancate e per bene ed alcuni esempi di sequenze}
-Experimental results are shown in Figures \ref{fig:ROCdispl_luma}, \ref{fig:ROCdispl_fd}, \ref{fig:ROCdef_grad}, \ref{fig:ROCdef_fd}.
-
 \begin{figure}
 \centering
-\subfigure[]{\label{fig:buenosAiresDef}\includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement_luma}}
-\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement_fd}}
-\caption{Examples of synthetically generated tampering events.
-\textbf{(a)} Defocus generated with gaussian filters on frames \textit{d} and \textit{e}.
-\textbf{(b)} Displacement created doing concatenation of similar sequences on frames \textit{d} and \textit{e}.}
-\label{fig:buenosAiresTamp}
+\subfigure[]{\label{fig:ROCdisplacement_luma}\includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement_luma}}
+\subfigure[]{\label{fig:ROCdisplacement_fd} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement_fd}}
+\caption{ROC curves for displacement detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean luma energy. \textbf{(b)} Analysis of the frame differencing.}
+\label{fig:ROCdisplacement}
 \end{figure}
-
 \begin{figure}
@@
 \subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus_fd}}
-\caption{Examples of synthetically generated tampering events.
-\textbf{(a)} Defocus generated with gaussian filters on frames \textit{d} and \textit{e}.
-\textbf{(b)} Displacement created doing concatenation of similar sequences on frames \textit{d} and \textit{e}.}
-\label{fig:buenosAiresTamp}
+\caption{ROC curves for defocus detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean gradient energy. \textbf{(b)} Analysis of the frame differencing.}
+\label{fig:ROCdisplacement}
+\label{fig:ROCdefocus}
 \end{figure}
-
+Experimental results are shown in Figures \ref{fig:ROCdisplacement} and \ref{fig:ROCdefocus}.
 \subsection{Discussion}
-\gi{Adiano: Aggiungi qua la complessit\'a computazionale}
+\gi{Adriano: Aggiungi qua la complessit\'a computazionale}
 

### 2015-05-21 paper.tex
Aggiunta formula FD continuato experiments e fatto ongoing works TODO integrazione STM

@@
 Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), and the acquired sequence does not evolve smoothly.
-This prevents the use of learned background models and the analysis of foreground variations \gi{REFERENCE su BKGR SUBTRACTION}.
+This prevents the use of learned background models and the analysis of foreground variations \cite{piccardi2004background}.
 When dynamic environments are acquired at low frame rates, like the example depicted in Fig.
@@
 %What are the key components of my approach and results? Also include any specific limitations.
-This work presents an algorithm to detect both camera defocus and of displacement, as stated in Section \ref{sec:probForm}.
+This work presents an algorithm to detect both camera defocus and displacement, as stated in Section \ref{sec:probForm}.
 The algorithm relies on simple indicators that are computed with a low computational complexity (Section \ref{subsec:Indicators}) and leverages an outlier detection technique to detect frames yielding anomalous indicators as related to a tampering event (Section \ref{subsec:MonitoringScheme}).
@@
 \caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
-In Figure \ref{fig:pioggia} there is a blur due to rain drops on the camera lens, while in Figure \ref{fig:neve} there is an occlusion due to some snow on the camera lens.}
+\textbf{(a)} Blur due to rain drops on the camera lens.
+\textbf{(b)} Occlusion due to some snow on the camera lens.}
 \label{fig:tampering}
@@
 
+FRAME DIFFERENCE:
+\begin{equation}
+	\label{eq:frameDiffReg}
+	\varphi_k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
+\end{equation}
+
 \subsection{Outlier Detection}\label{subsec:MonitoringScheme}
@@
 The proposed algorithm has been implemented in MATLAB, and has been tested on two datasets.
-The first one refers to frame sequences taken from webcams recording some parts of cities (as frames illustrated in Figure \ref{fig:buenosAires}), where we have introduced some synthetically generated tampering events:
-defocus has been simulated using gaussian filters, while displacement has been created with concatenation of similar frame sequences.
-In some cases these sequences contained real tampering events, as in Figure \ref{fig:fulvioTestiDispl}.\\
+The first one refers to frame sequences taken from webcams recording some parts of cities (as in Figure \ref{fig:sequences}(a)), where we have introduced some synthetically generated tampering events:
+defocus has been simulated using gaussian filters (\ref{fig:sequences}(b)), while displacement has been created with concatenation of similar frame sequences (Figure \ref{fig:sequences}(c)).
+In some cases these sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\
 \begin{figure}
 \centering
-\includegraphics[width=1\linewidth]{Immagini/sequenze}
+\includegraphics[width=0.9\linewidth]{Immagini/sequenze}
 \caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences; \textbf{(d)} real displacement event on 4-th frame.}
@@
 \item considering the whole scene for the features computation;
-\item considering adaptive region as described in Section \ref{sec:propSol} for the feature computation;
-\item considering voronoi regions for the feature computation.
+\item considering adaptive region, as described in Section \ref{sec:propSol}, for the feature computation;
+\item considering voronoi regions \cite{aurenhammer1991voronoi}, which are easier to compute with respect to our solution but don't consider the scene content, for the feature computation.
 \end{itemize}
@@
 Experimental results are shown in Figures \ref{fig:ROCdisplacement} and \ref{fig:ROCdefocus}.
+As we could see there is an improvement, in the detection of camera displacements, with our approach which separately analizes the behavior of indicators in adaptive regions.
+On the other side, as illustrated in Figure \ref{fig:ROCdefocus}, when monitoring indicators meant to detect blur/defocus, it is more convenient to consider the whole scene at once.
 \subsection{Discussion}
@@
 \item Displacement can be perceived by MEMS as well but these device alone are prone to false alarms. Visual inspection is necessary to reduce false alarms
-
 \end{itemize}
+Ongoing work regards the extension of our solution to other types of tampering, as occlusions or imaging sensor degradations.
+Furthermore, we are investigating strategies to improve the detection performance and reduce the number of \textit{FP}s by combination of our solution with other techniques:
+for example we could use sequential techniques on the features, or integrate the frame analysis with the data extracted from MEMS inertial sensors.
 

### 2015-05-22 paper.tex
sistemato piazzamento immagini (ho risparmiato una pagina!)
aggiunta tabella con AUC

@@
 \usepackage{graphicx}
+\usepackage[table,xcdraw]{xcolor}
 \usepackage[ruled]{algorithm2e}
@@
 Any frame yielding anomalous indicator is detected as a tampering attempt.
-Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements.
+Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image \ad{Giacomo: scene al posto di image??} during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements.
 Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, and at a very small computational complexity.
@@
 
-\begin{equation}
-\label{eq:energyLuma}
-l(t) = \mathcal{L}[z_t] =\frac{\sum_{\mathcal{X}} z_t(x) }{|\mathcal{X}|} ,
-\end{equation}
-
-\begin{equation}
-\label{eq:gradientDetr}
-\frac{\partial g}{\partial t}(t) = g(t) - g(t-1),
-\end{equation}
-
-\begin{equation}
-\label{eq:lumaDetr}
-\frac{\partial l}{\partial t}(t) = l(t) - l(t-1).
-\end{equation}
+%\begin{equation}
+%\label{eq:energyLuma}
+%l(t) = \mathcal{L}[z_t] =\frac{\sum_{\mathcal{X}} z_t(x) }{|\mathcal{X}|} ,
+%\end{equation}  
+%
+%
+%\begin{equation}
+%\label{eq:gradientDetr}
+%\frac{\partial g}{\partial t}(t) = g(t) - g(t-1),
+%\end{equation}
+%
+%
+%\begin{equation}
+%\label{eq:lumaDetr}
+%\frac{\partial l}{\partial t}(t) = l(t) - l(t-1).
+%\end{equation}
 
@@
 	\label{eq:frameDiffReg}
-	\varphi_k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
+	\varphi^k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
 \end{equation}
@@
 
-\begin{algorithm}[tp]
+\begin{algorithm}[t]
 	% \SetAlgoNoLine
@@
 In some cases these sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\
-\begin{figure}
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.9\linewidth]{Immagini/sequenze}
+\includegraphics[width=1\linewidth]{Immagini/sequenze}
 \caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences; \textbf{(d)} real displacement event on 4-th frame.}
@@
 %\gi{Adriano: metti entrambe le ROC curves, affiancate e per bene ed alcuni esempi di sequenze}
-\begin{figure}
+\begin{figure}[htb]
 \centering
@@
 \end{figure}
-\begin{figure}
+\begin{figure}[htb]
 \centering
@@
 \gi{Adriano: Aggiungi qua la complessit\'a computazionale}
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[tbh]
+\centering
+\begin{tabular}{l|l|l|l|l|l|l|}
+\cline{2-7}
+& \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Displacement Detection}}  & \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Defocus Detection}} \\ \cline{2-7} 
+& \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} & \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} \\ \hline
+\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Luma}}     & 0.9994 & 0.9989 & 0.9985  &            &            &             \\ \hline
+\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Gradient}} &  		 &  		  &             & 0.9895 & 0.9996 & 0.9996  \\ \hline
+\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{FD}}         & 0.9974 & 0.9944 & 0.9974  & 0.5526 & 0.5322 & 0.5391  \\ \hline
+\end{tabular}
+\caption{Area under curve of the analized solutions}
+\label{tab:AUC}
+\end{table}
+
+Numero di operazioni richieste:
+Calcolo di g(t):
+\begin{itemize}
+\item Norma gradiente: 196 FLOP per pixel
+\item Media delle norme del gradiente
+\end{itemize}
+Calcolo di l(t):
+\begin{itemize}
+\item Media dei valori di grigio
+\item Detrending: 1 FLOP per frame
+\end{itemize}
+Monitoraggio one-shot: 2 confronti
+%Monitoraggio sequenziale: 
+%Ogni 20 frame
+%Calcolo intervalli di confidenza: ~70 FLOP
+%Confronto tra intervalli di confidenza: 2 confronti
 

### 2015-05-26 paper.tex
messa un po di complessità computazionale
gitignore

@@
 
-Numero di operazioni richieste:
-Calcolo di g(t):
-\begin{itemize}
-\item Norma gradiente: 196 FLOP per pixel
-\item Media delle norme del gradiente
-\end{itemize}
-Calcolo di l(t):
-\begin{itemize}
-\item Media dei valori di grigio
-\item Detrending: 1 FLOP per frame
-\end{itemize}
-Monitoraggio one-shot: 2 confronti
+The algorithm uses low computational techniques: the biggest effort is made by computation of the feature $g(t)$, which needs $34$ operations per pixel, due to the Sobel filter used during the convolution.
+The computational load could decrease using Fast Fourier Transform (FFT) \cite{brigham1988fast} and operating in the frequency domain.
+
+%Numero di operazioni richieste:
+%Calcolo di g(t):
+%\begin{itemize}
+%\item Norma gradiente: 34 FLOP per pixel
+%\item Media delle norme del gradiente
+%\end{itemize}
+%Calcolo di l(t):
+%\begin{itemize}
+%\item Media dei valori di grigio
+%\item Detrending: 1 FLOP per frame
+%\end{itemize}
+%Monitoraggio one-shot: 2 confronti
 %Monitoraggio sequenziale: 

### 2015-05-26 paper.tex

@@
 \begin{abstract}
-A desirable feature for smart cameras is the ability to autonomously detect any tampering event/attack that would prevents a clear view over the monitored scene.
-No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attack (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures.
+A desirable feature for smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene.
+No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attacks (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures.
 Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at video-like frame-rates, nor use sophisticated image-analysis algorithms.
@@
 Any frame yielding anomalous indicator is detected as a tampering attempt.
-Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image \ad{Giacomo: scene al posto di image??} during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements.
+Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements.
 Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, and at a very small computational complexity.
@@
 \ref{fig:neve}.
-Similarly, other intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be surely useless for monitoring purposes.
+Similarly, intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be surely useless for monitoring purposes.
 We refer to these events/attacks as tampering.
 In some cases, tampering is easy to detect, e.g., when the camera integrity is affected and the device goes out-of-order.
-However, in many other situations, namely when the device is not physically damaged,	 but the correct interpretation of the scene or of small important details is prevented (e.g., the identification of licence plates), tampering detection is not straightforward and requires image-analysis algorithms.
+However, in many other situations, namely when the device is not physically damaged, tampering detection is not straightforward, and sometimes image analysis is the only viable option.
 
 % Why is it interesting and important?
-The early detection of tampering events/attacks is clearly essential in surveillance systems \gi{Adriano, add REFERENCES} \cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering event/attack and promptly report alerts.
-Surveillance cameras are typically connected to the power supply and acquire and process images at normal frame-rates (e.g.
-around few frames per second).
-In these conditions, several algorithms have been presented in the literature \gi{Adriano add REFERENCES}.
+Tampering prevents the correct scene interpretation and causes the loss of small, important details such as licence plates.
+Tampering detection is therefore an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly reporting alerts.
+Surveillance cameras are typically connected to the power supply and operates at normal frame-rates (e.g.
+around few frames per second): in these conditions, several tampering-detection algorithms have been presented~\gi{Adriano add REFERENCES}.
 
@@
 Magari aggiungendo che le batterie recenti permettono operativit\'a di due anni a questi regimi.}\cl{Verifico cosa possiamo dire del SecSoc.}
-This work expressly targets low-power and ultra-low-power smart cameras.
-Such devices typically operate at very low frame-rates (e.g.
-possibly less than one frame every minute), with constrained computational power and memory, and are battery powered (typically lasting one or two years).
-% We here consider low-power cameras, which are battery powered and that have to operate at very low frame-rates (e.g.
-one frame every minute or more) and which have a constrained computational power or memory availability.
+%
+In this work we expressly target low-power and ultra-low-power smart cameras, devices that operate at very low frame-rates (e.g.
+possibly less than one frame every minute), that are characterized by a constrained computational power and memory, and that are battery powered (typically lasting one or two years).
 As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where nodes are wirelessy connected smart-cameras, that can acquire and transmit images at regular time interval or upon requests.
-Often, in WMSN, the units are not connected to the power supply and have to operate with batteries, possibly implementing energy harvesting mechanisms \gi{REFERENCES, magari riprescando quelle del paper di TIM}.
-Even though these devices are not employed in critical surveillance applications, low-power smart cameras are becoming popular in distributed monitoring of wide environments due to their low cost and maintenance requirements \gi{possiamo mettere qualche REFERENCE}.
-Tampering detection is thus very important in low-power smart cameras, also considering that tampered frames should be identified to avoid unnecessary energy-demanding operations like the local processing or radio activation for transmission over the network.
+Often, in WMSN, the units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting mechanisms~\gi{REFERENCES, Adriano, guarda se c'e' qualche reference dal nostro paper di TIM}.
+Therefore, the power consumption is a serious concern in these devices.
+
+Even though low-power and ultra-low-power cameras not employed in critical surveillance applications, these devices are becoming popular in distributed systems for monitoring wide environments due to their low cost and maintenance requirements \gi{Adriano, Claudio, possiamo mettere qualche REFERENCE?}.
+Tampering detection is therefore very important in low-power smart cameras, where the algorithm are expected to be particularly prompt and have low false-alarm rates to prevent unnecessary energy-demanding operations like the local processing or radio activation for transmitting corrupted images (or false alarms) over the network.
+Unfortunately, tampering detection in such a resource-constrained scenario, has been much less investigated.
 
 % Why is it hard? (E.g., why do naive approaches fail?)
-Tampering detection in low-power smart cameras is much more challenging than in conventional surveillance cameras \cite{perrig2004security}.
-Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), and the acquired sequence does not evolve smoothly.
+Tampering detection in low-power smart cameras is much more challenging than in conventional surveillance cameras~\cite{perrig2004security}.
+Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), thus the acquired sequence does not evolve smoothly.
 This prevents the use of learned background models and the analysis of foreground variations \cite{piccardi2004background}.
-When dynamic environments are acquired at low frame rates, like the example depicted in Fig.
-\ref{fig:buenosAires}, two consecutive frames might be very different because of changes in the scene and in the light conditions: smart cameras have to correctly distinguish between these \emph{normal} changes and changes due to camera displacement or blur/defocus.
-Moreover, in low-power smart cameras, tampering has to be reliably detected because false alerts would lead to unnecessary energy waste.
+When dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions, resulting in sequences those depicted in Figure~\ref{fig:sequences}: smart cameras have to correctly distinguish between these \emph{normal} changes and changes due to camera tampering.
 
@@
 %What are the key components of my approach and results? Also include any specific limitations.
-This work presents an algorithm to detect both camera defocus and displacement, as stated in Section \ref{sec:probForm}.
-The algorithm relies on simple indicators that are computed with a low computational complexity (Section \ref{subsec:Indicators}) and leverages an outlier detection technique to detect frames yielding anomalous indicators as related to a tampering event (Section \ref{subsec:MonitoringScheme}).
-In particular, we perform a preliminarily partitioning of the scene into nonoverlapping regions, and separately compute and monitor indicators over different regions.
-Scene partitioning is defined by clustering feature vectors gathering the indicators in different pixels, thus performing a very coarse segmentation of the image (Section \ref{subsec:Segmentation}).
-Regions are computed during the initial configuration of the algorithm and as such do not implies any computational overhead w.r.t.
-operating on the whole image.
-Our experiments in Section \ref{sec:experiments} show that regions substantially improve the detection of camera displacements, while when monitoring indicators meant to detect blur/defocus, it is more convenient to consider the whole scene at once.
+We present an algorithm to detect both camera defocus and displacement, as stated in Section \ref{sec:probForm}.
+The algorithm presented in Section~\ref{sec:Algorithm} relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blur.
+These indicators can be easily computed and monitored by an outlier-detection technique to detect the respective tampering.
+We show that displacement-detection performance can be improved by means of a preliminarily partitioning of the scene into nonoverlapping regions, and then separately computing and monitoring the average intensity over different regions (Section \ref{subsec:Segmentation}).
+Remarkably, these regions need to be defined during an initial configuration phase of the smart camera, thus there is no relevant computational overhead w.r.t.
+monitoring the whole image.
+Experiments in Section \ref{sec:experiments} show that operating on image regions can substantially improve the detection of camera displacements.
 Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.
@@
 \subfigure[]{\label{fig:neve} \includegraphics[width=0.4\linewidth]{Immagini/neve}}
-\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena.
-\textbf{(a)} Blur due to rain drops on the camera lens.
-\textbf{(b)} Occlusion due to some snow on the camera lens.}
+\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena. \textbf{(a)} Blur due to rain drops on the camera lens. \textbf{(b)} Occlusion due to snow on the camera lens.}
 \label{fig:tampering}
@@
 These types of solutions requires a lot of computations
-
+%
+%
+%
 \section{Problem Formulation}\label{sec:probForm}
-
+%
 Let $z_t$ be the frame acquired at time $t$, which we model as
@@
 
-\section{Proposed Solution}\label{sec:propSol}
+\section{The Proposed Algorithm}\label{sec:propSol}
+
+The proposed tampering-detection algorithm is illustrated in Algorithm \ref{alg:DISPL}, and relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma} and denoted by $l$) and the average norm of the gradient (denoted by $g$).
+As anticipated in Section \ref{sec:introduction}, during algorithm configuration we perform a preliminary segmentation of the scene, yielding $\{R_k\}, k = 1,\dots,K,$  $R_k \subset \mathcal{X}$, a collection of disjoint subsets ($R_i  \cap R_j = \emptyset, \forall i \neq j$), and we compute and monitor the luma indicator separately on each $R_k$.
+Details about the scene segmentation algorithm used are reported in Section \ref{subsec:Segmentation}.
+
+Luma is computed separately on each of the $K$ image regions in each frame $z_t$
+\begin{equation}\label{eq:lumaRegions}
+l_k(t) =\frac{1}{\#R_k} \sum_{R_k} z_t(x), \ \ k = 1, \dots, K\,,
+\end{equation}
+while the gradient norm is
+\begin{equation}
+\label{eq:normaGradiente}
+g(t) = \sum_X \left (\sqrt{\left(z_t \circledast f_h\right)^2(x) + \left(z_t \circledast f_v\right)^2(x)}\right),
+\end{equation}
+where $f_h$ and $f_h$ are horizontal and vertical derivative filters, respectively, $\#(\cdot)$ indicates the cardinality of a set and $\circledast$ the 2d convolution.
+
+Tampering attacks are analyzed by determining whether indicators computed from each incoming frame yield an outlier. Since these indicator are univariate, we simply
+
+Monitoring these indicators is challenging because they are highly correlated overtime and and it is not possible to 
+%\partial l^k(t) & =& l^k(t)-l^k(t-1)
+
+\begin{algorithm}[t]
+	% \SetAlgoNoLine
+	\LinesNumbered
+	\SetAlgoNlRelativeSize{0}
+	\SetNlSty{small}{}{.}
+	\textbf{Input}: $\gamma$, $T_{o}$, $\{R_k\}, k=1,\dots,K$ \\
+	\textbf{Configuration}:\\
+%	\lnl{DISPL-Tr1} Extract regions $\{R_k\}, k=1,\dots,K$  \\
+	\lnl{DISPL-Tr2} \For{$t=1,\dots,T_{o}$}
+	{	\lnl{DISPL-Tr3} Get frame $z_t$ \\
+		\lnl{DISPL-Tr4} \For{$k=1,\dots,K$}{
+			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
+		}
+	}
+	\lnl{DISPL-Tr6} \For{$k=1,\dots,K$}{
+		\lnl{DISPL-Tr7} Compute $\sigma_l^k$\\
+	}
+	\textbf{Operational phase}:\\
+	\lnl{DISPL-Test1} \For{$t=T_{o},\dots,\infty$}{
+		\lnl{DISPL-Test2} Get frame $z_t$ \\
+		\lnl{DISPL-Test3} $n_l = 0$ \\
+		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
+			\lnl{DISPL-Test5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
+			\lnl{DISPL-Test6} \If{$\partial l^k(t) < -\gamma \sigma_l^k \vee \partial l^k(t) > \gamma \sigma_l^k $}{
+				\lnl{DISPL-Test7} $n_l=n_l+1$\\
+			}
+		}
+		\lnl{DISPL-Test8} \If{$n_l\geq K-1$}{
+			\lnl{DISPL-Test9} $z_t$ is a tampered frame\\
+		}
+	}   
+	\caption{Tampering detection algorithm}
+	\label{alg:DISPL}
+\end{algorithm}
 
@@
 \end{itemize}
+%
+%
+%
 \subsection{Indicators}\label{subsec:Indicators}
-\gi{Adriano: mettere formule degli indicatori e anche del frame difference qua}
 
@@
 Con questi filtri \`e possibile calcolare la \textit{norma del gradiente} nel seguente modo:
-\begin{equation}
-\label{eq:normaGradiente}
-\| \nabla z_t(x) \|_2^2=\left(z_t \circledast f_h\right)(x)^2 + \left(z_t \circledast f_v\right)(x)^2.
-\end{equation}
+
 Una volta calcolata la norma del gradiente \`e possibile farne la media come specificato in \eqref{eq:energyGradient}.
@@
 In particolare ci aspettiamo che l'evento di sfocatura provochi un abbattimento del valore di $g$.
-
-\begin{equation}
-\label{eq:lumaRegions}
-\begin{array}{ccc}
-l^k(t)&  = & \mathcal{L}^k[z_t] = \frac{\sum_{R_k} z_t(x) }{|{R_k}|}\\
-\partial l^k(t) & =& l^k(t)-l^k(t-1) 
-\end{array},
-\end{equation}
 
@@
 %\end{equation}
-
-FRAME DIFFERENCE:
-\begin{equation}
-	\label{eq:frameDiffReg}
-	\varphi^k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
-\end{equation}
 
@@
 
-\begin{algorithm}[t]
-	% \SetAlgoNoLine
-	\LinesNumbered
-	\SetAlgoNlRelativeSize{0}
-	\SetNlSty{small}{}{.}
-	\textbf{Input}: $\gamma$, $T_{o}$, $\{R_k\}, k=1,\dots,K$ \\
-	\textbf{Configuration}:\\
-%	\lnl{DISPL-Tr1} Extract regions $\{\mathcal{R}_k\}, k=1,\dots,K$  \\
-	\lnl{DISPL-Tr2} \For{$t=1,\dots,T_{o}$}
-	{	\lnl{DISPL-Tr3} Get frame $z_t$ \\
-		\lnl{DISPL-Tr4} \For{$k=1,\dots,K$}{
-			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
-		}
-	}
-	\lnl{DISPL-Tr6} \For{$k=1,\dots,K$}{
-		\lnl{DISPL-Tr7} Compute $\sigma_l^k$\\
-	}
-	\textbf{Operational phase}:\\
-	\lnl{DISPL-Test1} \For{$t=T_{o},\dots,\infty$}{
-		\lnl{DISPL-Test2} Get frame $z_t$ \\
-		\lnl{DISPL-Test3} $n_l = 0$ \\
-		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
-			\lnl{DISPL-Test5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
-			\lnl{DISPL-Test6} \If{$\partial l^k(t) < -\gamma \sigma_l^k \vee \partial l^k(t) > \gamma \sigma_l^k $}{
-				\lnl{DISPL-Test7} $n_l=n_l+1$\\
-			}
-		}
-		\lnl{DISPL-Test8} \If{$n_l\geq K-1$}{
-			\lnl{DISPL-Test9} $z_t$ is a tampered frame\\
-		}
-	}   
-	\caption{Tampering detection algorithm}
-	\label{alg:DISPL}
-\end{algorithm}
 %\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
@@
 \end{itemize}
+
+FRAME DIFFERENCE:
+\begin{equation}
+	\label{eq:frameDiffReg}
+	\varphi^k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
+\end{equation}
+
 %
@@
 The algorithm uses low computational techniques: the biggest effort is made by computation of the feature $g(t)$, which needs $34$ operations per pixel, due to the Sobel filter used during the convolution.
-The computational load could decrease using Fast Fourier Transform (FFT) \cite{brigham1988fast} and operating in the frequency domain.
 

### 2015-05-26 paper.tex
Inserite reference su weighted kmeans

@@
 As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where nodes are wirelessy connected smart-cameras, that can acquire and transmit images at regular time interval or upon requests.
-Often, in WMSN, the units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting mechanisms~\gi{REFERENCES, Adriano, guarda se c'e' qualche reference dal nostro paper di TIM}.
+Often, in WMSN, the units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting mechanisms~\gi{REFERENCES, Adriano, guarda se c'e' qualche reference dal nostro paper di TIM} \cite{magno2009adaptive}.
 Therefore, the power consumption is a serious concern in these devices.
@@
 We present an algorithm to detect both camera defocus and displacement, as stated in Section \ref{sec:probForm}.
-The algorithm presented in Section~\ref{sec:Algorithm} relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blur.
+The algorithm presented in Section~\ref{sec:propSol} relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blur.
 These indicators can be easily computed and monitored by an outlier-detection technique to detect the respective tampering.
@@
 The literature concerning tampering detection is mostly focused on video surveillance applications and operates at few frames per second \gi{Adriano, \'e vero?}.
-Background models are typically leveraged to identify defocus and occlusions; in particular  ~\cite{aksay2007camera} performs detect defocus by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered.
+Background models are typically leveraged to identify defocus and occlusions; in particular  ~\cite{aksay2007camera} performs detect defocus \ad{defocus detection?} by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered.
 A background-subtraction technique is employed in~\cite{saglam2009real} to identify defocus, occlusions, and displacements.
@@
 \end{itemize}
-%
-%
-%
+
+When the feature vectors are computed for all the pixels we start the segmentation phase.
+This is made using classical clustering techniques:
+in particular we use a modified version of the classical k-means algorithm \cite{han2006data}. 
+In this implementation, each feature vector belongs to a cluster if its \textit{weighted Minkowski distance} \cite{de2012minkowski} from the cluster's centroid is minimum with respect to the distance from the other centroids.
+The distance is weighted considering the standard deviation of each elements of the feature vectors inside each cluster \cite{kottke1994motion}, in order to consider, during the clustering, also the contribute of the fields which have low variance.
+\ad{Giacomo: prova a vedere se si capisce cosa voglio dire in queste frasi!!!}
+
 \subsection{Indicators}\label{subsec:Indicators}
@@
 The algorithm uses low computational techniques: the biggest effort is made by computation of the feature $g(t)$, which needs $34$ operations per pixel, due to the Sobel filter used during the convolution.
+The computational load could decrease using Fast Fourier Transform (FFT) \cite{brigham1988fast} and operating in the frequency domain.
 

### 2015-05-26 paper.tex
edits in Section 3

--- 2015-05-26 paper.tex
@@
 \newcommand{\cl}[1]{{\textcolor{red}{[\small \textbf{Claudio}: #1]}}}
+
+%\newcommand{\gi}[1]{}
+%\newcommand{\ad}[1]{}
+%\newcommand{\cl}[1]{}
 
@@
 \end{equation}
-%where  $\mathcal{B}_t$ corresponds to a linear operator
-%\begin{equation}
-%\label{eq:blur}
-%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(x,s)ds\,,
-%\end{equation}
 where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.
-%\begin{equation}
-%\label{eq:blur_convolution}
-%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(s-x)ds,
-%\end{equation}
-%
-%\begin{equation}
-%\label{eq:blur_multi}
-%z_t(x)=\mathcal{D}_t[y_t](x) = \mathcal{B}_t[y_t](x) + \eta(x), \qquad x \in \mathcal{X}.
-%\end{equation}
+
 A \emph{camera displacement} at frame $\tau^*$ is instead modeled as 
@@
 The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect time $\tau^*$ when any tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.
-
-%Per $t<T^*$ abbiamo che i frame vengono acquisiti in condizioni di funzionamento normale:
-%\[ z_t(x)=y_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t=1,\dots , T^*-1. \] 
-%All'istante di tempo $t = T^*$ avviene un evento di tampering, il quale compromette i frame per $t\geq T^*$.  
-%In particolare, nel caso di una sfocatura avremo
-%\[ z_t = \mathcal{B}_t[y_t](x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*,\]
-%mentre nel caso di uno spostamento della camera avremo
-%\[ z_t = w_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*. \]
+We require that the initial $T_0$ frames are without tampering and provided for training.
 
@@
 
-The proposed tampering-detection algorithm is illustrated in Algorithm \ref{alg:DISPL}, and relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma} and denoted by $l$) and the average norm of the gradient (denoted by $g$).
-As anticipated in Section \ref{sec:introduction}, during algorithm configuration we perform a preliminary segmentation of the scene, yielding $\{R_k\}, k = 1,\dots,K,$  $R_k \subset \mathcal{X}$, a collection of disjoint subsets ($R_i  \cap R_j = \emptyset, \forall i \neq j$), and we compute and monitor the luma indicator separately on each $R_k$.
-Details about the scene segmentation algorithm used are reported in Section \ref{subsec:Segmentation}.
+Algorithm \ref{alg:DISPL} presents the proposed tampering detection, which relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$).
+The former is meant to detect camera displacement, the latter camera defocus.
+As anticipated in Section \ref{sec:introduction}, during the initial configuration, the scene is segmented in $\{R_k\}, k = 1,\dots,K$ disjoint regions ($R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$).
+Details about the employed segmentation algorithm are in Section \ref{subsec:Segmentation}.
 
-Luma is computed separately on each of the $K$ image regions in each frame $z_t$
+In each frame $z_t$ the luma is separately computed in of the $K$ regions,
 \begin{equation}\label{eq:lumaRegions}
@@
 \end{equation}
-while the gradient norm is
+while the average gradient norm is computed over the whole image
 \begin{equation}
@@
 \end{equation}
-where $f_h$ and $f_h$ are horizontal and vertical derivative filters, respectively, $\#(\cdot)$ indicates the cardinality of a set and $\circledast$ the 2d convolution.
+where $f_h$ and $f_h$ are the horizontal and vertical derivative filters, respectively, $\#(\cdot)$ indicates the cardinality of a set and $\circledast$ the 2d convolution.
 
-Tampering attacks are analyzed by determining whether indicators computed from each incoming frame yield an outlier. Since these indicator are univariate, we simply
+As shown in Figure REFERENCE\_TODO.a), tampering results in outliers in the sequence of indicators.
+Detecting these outliers is not straightforward, since indicators heavily fluctuates in normal conditions because of short-term variations in the depicted scene or illumination changes.
+For outlier-detection purposes~\cite{Chandola2009}, it is more convenient to operate on sequences of independent and identically distributed (i.i.d.) values, where outliers can be detected by density-based methods or simply by confidence regions that is what we considered in our algorithm.
+Unfortunately, trends or deterministic variations in the indicator sequences are not easy to compensate because of unpredictable evolutions of the scene.
 
-Monitoring these indicators is challenging because they are highly correlated overtime and and it is not possible to 
-%\partial l^k(t) & =& l^k(t)-l^k(t-1)
+To obtain indicators that in normal conditions can be treated as i.i.d. realizations of a random variable we detrend~\cite{Gustafsson2000} the indicator sequence by performing temporal derivative
+\begin{equation}\label{eq:detrending}
+ \partial l_k(t) = l_k(t)-l_k(t-1),  \ \ k = 1, \dots, K\,,
+\end{equation}
+and we similarly define $\partial g(t)$.
+Detrended indicators are more suitable to be monitored by a confidence interval since, as shown in Figure REFERENCE\_TODO.b).
+However, a tampering yielding permanent changes in the indicators, as those in Figure REFERENCE\_TODO.a results in a single outlier in the detrended indicator~\eqref{eq:detrending}.
 
@@
 	\SetNlSty{small}{}{.}
-	\textbf{Input}: $\gamma$, $T_{o}$, $\{R_k\}, k=1,\dots,K$ \\
+	\textbf{Input}: $\gamma$, $T_{o}$, $\{R_k\} k=1,\dots,K$ \\
 	\textbf{Configuration}:\\
-%	\lnl{DISPL-Tr1} Extract regions $\{R_k\}, k=1,\dots,K$  \\
-	\lnl{DISPL-Tr2} \For{$t=1,\dots,T_{o}$}
-	{	\lnl{DISPL-Tr3} Get frame $z_t$ \\
-		\lnl{DISPL-Tr4} \For{$k=1,\dots,K$}{
-			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
-		}
-	}
-	\lnl{DISPL-Tr6} \For{$k=1,\dots,K$}{
-		\lnl{DISPL-Tr7} Compute $\sigma_l^k$\\
-	}
+	\lnl{DISPL-Tr1} Compute $\partial l_k(t)$ and $\partial g(t), t = 1, \cdots, T_o$\\ 
+	\lnl{DISPL-Tr7} Compute $\sigma_{l_k}$ and $\sigma_{g}$, the standard deviations of $\partial l_k$ and $\partial g$.\\
+
 	\textbf{Operational phase}:\\
-	\lnl{DISPL-Test1} \For{$t=T_{o},\dots,\infty$}{
-		\lnl{DISPL-Test2} Get frame $z_t$ \\
-		\lnl{DISPL-Test3} $n_l = 0$ \\
+	\lnl{DISPL-Test1} \For{$t=T_{o} + 1,\dots,\infty$}{
+		\lnl{DISPL-Test2} Get frame $z_t$, set $n_l =0$
 		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
-			\lnl{DISPL-Test5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
-			\lnl{DISPL-Test6} \If{$\partial l^k(t) < -\gamma \sigma_l^k \vee \partial l^k(t) > \gamma \sigma_l^k $}{
+			\lnl{DISPL-Test5} Compute $\partial l_k(t)$\\
+			\lnl{DISPL-Test6} \If{$\partial l_k(t) < -\gamma \sigma_{l_k} \vee \partial l_k(t) > \gamma \sigma_{l_k} $}{
 				\lnl{DISPL-Test7} $n_l=n_l+1$\\
@@
 		\lnl{DISPL-Test8} \If{$n_l\geq K-1$}{
-			\lnl{DISPL-Test9} $z_t$ is a tampered frame\\
+			\lnl{DISPL-Test9} tampering (probably displacement or occlusion) has occurred in $z_t$ \\
 		}
+		\lnl{DISPL-Test10} Compute $\partial g(t)$\\
+		\lnl{DISPL-Test11} \If{$\partial g(t) < -\gamma \sigma_{g} \vee \partial l(t) > \gamma \sigma_{g} $}{
+				\lnl{DISPL-Test9} tampering (probably out-of-focus) has occurred in $z_t$ \\
 								}
-	\caption{Tampering detection algorithm}
+
+	}   
+	\caption{The Proposed Tampering-Detection Algorithm}
 	\label{alg:DISPL}
@@
 \end{itemize}
-%
-%
-%
-\subsection{Indicators}\label{subsec:Indicators}
-
-\begin{equation}
-\label{eq:gradientRegions}
-\begin{array}{ccc}
-g^k(t)&  = & \mathcal{G}^k[z_t] = \frac{\sum_{R_k}\| \nabla z_t(x) \| _2^2 }{|{R_k}|}\\
-\partial g^k(t) & =& g^k(t)-g^k(t-1) 
-\end{array},
-\end{equation}
-
-In particolare, per il calcolo delle derivate orizzontali  abbiamo utilizzato il seguente filtro $f_h$:
-\[f_h = f \circledast \left[ \begin{array}{rcl}
-1 & 0 & -1
-\end{array}\right], \] 
-mentre per il calcolo delle derivate verticali abbiamo utilizzato il seguente filtro $f_v$:
-\[f_v = f \circledast \left[ \begin{array}{r}
-1 \\ 0 \\ -1
-\end{array}\right], \]
-dove abbiamo indicato con $\circledast$ l'operatore di convoluzione.
-Il filtro $f$, invece, \`e ottenuto tramite un campionamento della \textit{funzione gaussiana} $h$, con media $0$ e deviazione standard $\sigma$
-\begin{equation}
-\label{eq:gaussian}
-h(i,j)=\frac{1}{2\pi\sigma^2}\exp\left(-\frac{i^2+j^2}{2\sigma^2}\right),
-\end{equation}
-e ponendo il valore massimo di questa funzione nel centro del filtro.
-Con questi filtri \`e possibile calcolare la \textit{norma del gradiente} nel seguente modo:
-
-Una volta calcolata la norma del gradiente \`e possibile farne la media come specificato in \eqref{eq:energyGradient}.
-Il risultato finale \`e un indicatore \textit{scalare} per ciascun frame acquisito, che pu\`o essere monitorato per individuare eventi di sfocature. 
-In particolare ci aspettiamo che l'evento di sfocatura provochi un abbattimento del valore di $g$.
 
@@
 \end{document}
+
+%
+%
+%
+\subsection{Indicators}\label{subsec:Indicators}
+
+%\begin{equation}
+%\label{eq:gradientRegions}
+%\begin{array}{ccc}
+%g^k(t)&  = & \mathcal{G}^k[z_t] = \frac{\sum_{R_k}\| \nabla z_t(x) \| _2^2 }{|{R_k}|}\\
+%\partial g^k(t) & =& g^k(t)-g^k(t-1) 
+%\end{array},
+%\end{equation}
+%
+%In particolare, per il calcolo delle derivate orizzontali  abbiamo utilizzato il seguente filtro $f_h$:
+%\[f_h = f \circledast \left[ \begin{array}{rcl}
+%1 & 0 & -1
+%\end{array}\right], \] 
+%mentre per il calcolo delle derivate verticali abbiamo utilizzato il seguente filtro $f_v$:
+%\[f_v = f \circledast \left[ \begin{array}{r}
+%1 \\ 0 \\ -1
+%\end{array}\right], \]
+%dove abbiamo indicato con $\circledast$ l'operatore di convoluzione.
+%Il filtro $f$, invece, \`e ottenuto tramite un campionamento della \textit{funzione gaussiana} $h$, con media $0$ e deviazione standard $\sigma$
+%\begin{equation}
+%\label{eq:gaussian}
+%h(i,j)=\frac{1}{2\pi\sigma^2}\exp\left(-\frac{i^2+j^2}{2\sigma^2}\right),
+%\end{equation}
+%e ponendo il valore massimo di questa funzione nel centro del filtro.
+%Con questi filtri \`e possibile calcolare la \textit{norma del gradiente} nel seguente modo:
+%
+%Una volta calcolata la norma del gradiente \`e possibile farne la media come specificato in \eqref{eq:energyGradient}.
+%Il risultato finale \`e un indicatore \textit{scalare} per ciascun frame acquisito, che pu\`o essere monitorato per individuare eventi di sfocature. 
+%In particolare ci aspettiamo che l'evento di sfocatura provochi un abbattimento del valore di $g$.

### 2015-05-26 paper.tex
section 3 fino ad outlier detection

@@
 The literature concerning tampering detection is mostly focused on video surveillance applications and operates at few frames per second \gi{Adriano, \'e vero?}.
-Background models are typically leveraged to identify defocus and occlusions; in particular  ~\cite{aksay2007camera} performs detect defocus \ad{defocus detection?} by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered.
+Background models are typically leveraged to identify defocus and occlusions; in particular~\cite{aksay2007camera} performs defocus detection by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered.
 A background-subtraction technique is employed in~\cite{saglam2009real} to identify defocus, occlusions, and displacements.
@@
 
-To obtain indicators that in normal conditions can be treated as i.i.d. realizations of a random variable we detrend~\cite{Gustafsson2000} the indicator sequence by performing temporal derivative
+To obtain indicators that in normal conditions can be treated as i.i.d. realizations of a random variable we, detrend~\cite{Gustafsson2000} the indicator sequence by a temporal derivative
 \begin{equation}\label{eq:detrending}
@@
 However, a tampering yielding permanent changes in the indicators, as those in Figure REFERENCE\_TODO.a results in a single outlier in the detrended indicator~\eqref{eq:detrending}.
+
+We perform outlier detection by defining confidence regions for our indicator in tampering-free frames.
+In practice, since all these are scalar indicators, it is enough to compute the indicator mean ($\overline{\partial l}_k(t), k = 1,\dots,K$ and $\overline{\partial g}(t)$) and standard deviations ($\sigma_{l_k}, k = 1,\dots,K$ and $\sigma_g$) over the tampering-free frames provided for training, namely $z_t\,, t = 1, \dots, T_o$.
+Confidence regions are then defined as
+\begin{equation}\label{eq:confidenceRegions}
+ [\overline{\partial l}_k(t) - \gamma \sigma_{l_k}, \overline{\partial l}_k(t) + \gamma \sigma_{l_k}],  k = 1,\dots,K
+\end{equation}
+where $\gamma >0$ is a tuning parameter.
+A similar region is built for $\partial g$.
+During operations, any indicator falling outside the corresponding confidence region is considered anomalous and tampering is detected.
 
@@
 								}
-
 	}   
@@
 \end{itemize}
-
-%\begin{equation}
-%\label{eq:energyLuma}
-%l(t) = \mathcal{L}[z_t] =\frac{\sum_{\mathcal{X}} z_t(x) }{|\mathcal{X}|} ,
-%\end{equation}  
-%
-%
-%\begin{equation}
-%\label{eq:gradientDetr}
-%\frac{\partial g}{\partial t}(t) = g(t) - g(t-1),
-%\end{equation}
-%
-%
-%\begin{equation}
-%\label{eq:lumaDetr}
-%\frac{\partial l}{\partial t}(t) = l(t) - l(t-1).
-%\end{equation}
-
-\subsection{Outlier Detection}\label{subsec:MonitoringScheme}
-
-\begin{equation}
-\label{eq:soglieGradiente}
-\begin{array}{lcl}
-\Gamma_{min}^k & = & \hat{\mu}_g^k -\gamma \hat{\sigma}_g^k\\
-\Gamma_{max}^k & = & \hat{\mu}_g^k + \gamma \hat{\sigma}_g^k
-\end{array},
-\end{equation}
-dove $\hat{\mu}_g^k$ indica il valore medio delle osservazioni del training set
-\begin{equation}
-\hat{\mu}_g^k = \frac{\sum_{\tau = 1}^{T_{o}} \frac{\partial g^k}{\partial t}(\tau)}{T_{o}}, \nonumber
-\end{equation}
-$\hat{\sigma}_g^k$ indica la deviazione standard delle osservazioni del training set
-\begin{equation}
-\hat{\sigma}_g^k  = \sqrt{\frac{1}{T_{o}-1}\sum_{\tau=1}^{T_{o}}\left(\frac{\partial g^k}{\partial t}(\tau) - \hat{\mu}_g^k(\tau)\right)^2} \nonumber
-\end{equation}
-e $\gamma>1$ \`e un parametro moltiplicativo ottenuto sperimentalmente.\\
-
-\begin{equation}
-\label{eq:soglieLuma}
-\begin{array}{rcl}
-\Gamma_{min}^k & = & \hat{\mu}_l^k -\gamma \hat{\sigma}_l^k\\
-\Gamma_{max}^k & = & \hat{\mu}_l^k + \gamma \hat{\sigma}_l^k
-\end{array},
-\end{equation}
-dove $\hat{\mu}_l^k$ indica il valore medio delle osservazioni del training set
-\begin{equation}
-\hat{\mu}_l^k = \frac{\sum_{\tau = 1}^{T_{o}} \frac{\partial l^k}{\partial t}(\tau)}{T_{o}}, \nonumber
-\end{equation}
-$\hat{\sigma}_l^k$ indica la deviazione standard delle osservazioni del training set
-\begin{equation}
-\hat{\sigma}_l^k  = \sqrt{\frac{1}{T_{o}-1}\sum_{\tau=1}^{T_{o}}\left(\frac{\partial l^k}{\partial t}(\tau) - \hat{\mu}_l^k(\tau)\right)^2} \nonumber
-\end{equation}
-e $\gamma>1$ \`e un parametro moltiplicativo ottenuto sperimentalmente.\\
-
-\subsection{Algorithm Summary}\label{subsec:AlgorithmSummary}
-
-%\begin{algorithm}[tp]
-%	% \SetAlgoNoLine
-%	\LinesNumbered
-%	\SetAlgoNlRelativeSize{0}
-%	\SetNlSty{small}{}{.}
-%	\textbf{Configuration}:\\
-%	\lnl{DEF-Tr0} Extract regions $\{R_k\}, k=1,\dots,K$  \\
-%	\lnl{DEF-Tr1} \For{$t=1,\dots,T_{o}$}
-%	{	\lnl{DEF-Tr2} Acquire frame $z_t$ \\
-%		\lnl{DEF-Tr2a} \For{$k=1,\dots,K$}{
-%			\lnl{DEF-Tr3a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
-%		}
-%		\lnl{DEF-Tr3} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
-%	}
-%	\lnl{DEF-Tr6a} \For{$k=1,\dots,K$}{
-%		\lnl{DEF-Tr7a} Define thresholds $\Gamma_{min}^k$ and $\Gamma_{max}^k$\\
-%	}
-%	\lnl{DEF-Tr5} Define CDT parameters on $g(t)$ variance\\
-%	\textbf{Operational phase}:\\
-%	\lnl{DEF-Test1} \For{$t=T_{o},\dots,\infty$}{
-%		\lnl{DEF-Test2} Acquire frame $z_t$ \\
-%		\lnl{DEF-Tr3b} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
-%		\lnl{DEF-Test2a} $n=0$\\
-%		\lnl{DEF-Test4a} \For{$k=1,\dots,K$}{
-%			\lnl{DEF-Test5a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
-%			\lnl{DEF-Test6a} \If{$\frac{\partial g^k}{\partial t}(t) < \Gamma_{min}^k \vee \frac{\partial g^k}{\partial t}(t) > \Gamma_{max}^k $}{
-%				\lnl{DEF-Test7a} $n=n+1$\\
-%			}
-%		}
-%		\lnl{DEF-Test81} \If{$n \geq K-1$}{
-%			\lnl{DEF-Test91} $z_t$ is a defocused frame\\
-%		}
-%		\lnl{DEF-Test8} \If{CDT detect a change on $g(t)$ variance}{
-%			\lnl{DEF-Test9} $z_t$ is a defocused frame\\
-%		}
-%	}   
-%	\caption{Blur detection algorithm}
-%	\label{alg:DEFOCUS}
-%\end{algorithm}
-
-%\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
 

### 2015-05-26 paper.tex
Rimessa la segmentazione...

@@
 
+When the feature vectors are computed for all the pixels we start the segmentation phase.
+This is made using classical clustering techniques:
+in particular we use a modified version of the classical k-means algorithm \cite{han2006data}. 
+In this implementation, each feature vector belongs to a cluster if its \textit{weighted Minkowski distance} \cite{de2012minkowski} from the cluster's centroid is minimum with respect to the distance from the other centroids.
+The distance is weighted considering the standard deviation of each elements of the feature vectors inside each cluster \cite{kottke1994motion}, in order to consider, during the clustering, also the contribute of the fields which have low variance.
+\ad{Giacomo: prova a vedere se si capisce cosa voglio dire in queste frasi!!!}
+
 \section{Experiments}\label{sec:experiments}

### 2015-05-26 paper.tex
Messo immagine indicatori

--- 2015-05-26 paper.tex
@@
 \newcommand{\cl}[1]{{\textcolor{red}{[\small \textbf{Claudio}: #1]}}}
-
-%\newcommand{\gi}[1]{}
-%\newcommand{\ad}[1]{}
-%\newcommand{\cl}[1]{}
 
@@
 As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where nodes are wirelessy connected smart-cameras, that can acquire and transmit images at regular time interval or upon requests.
-Often, in WMSN, the units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting mechanisms~\gi{REFERENCES, Adriano, guarda se c'e' qualche reference dal nostro paper di TIM} \cite{magno2009adaptive}.
+Often, in WMSN, the units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting mechanisms~\gi{REFERENCES, Adriano, guarda se c'e' qualche reference dal nostro paper di TIM}.
 Therefore, the power consumption is a serious concern in these devices.
@@
 We present an algorithm to detect both camera defocus and displacement, as stated in Section \ref{sec:probForm}.
-The algorithm presented in Section~\ref{sec:propSol} relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blur.
+The algorithm presented in Section~\ref{sec:Algorithm} relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blur.
 These indicators can be easily computed and monitored by an outlier-detection technique to detect the respective tampering.
@@
 The literature concerning tampering detection is mostly focused on video surveillance applications and operates at few frames per second \gi{Adriano, \'e vero?}.
-Background models are typically leveraged to identify defocus and occlusions; in particular  ~\cite{aksay2007camera} performs detect defocus \ad{defocus detection?} by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered.
+Background models are typically leveraged to identify defocus and occlusions; in particular  ~\cite{aksay2007camera} performs detect defocus by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered.
 A background-subtraction technique is employed in~\cite{saglam2009real} to identify defocus, occlusions, and displacements.
@@
 \end{equation}
+%where  $\mathcal{B}_t$ corresponds to a linear operator
+%\begin{equation}
+%\label{eq:blur}
+%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(x,s)ds\,,
+%\end{equation}
 where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.
-
+%\begin{equation}
+%\label{eq:blur_convolution}
+%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(s-x)ds,
+%\end{equation}
+%
+%\begin{equation}
+%\label{eq:blur_multi}
+%z_t(x)=\mathcal{D}_t[y_t](x) = \mathcal{B}_t[y_t](x) + \eta(x), \qquad x \in \mathcal{X}.
+%\end{equation}
 A \emph{camera displacement} at frame $\tau^*$ is instead modeled as 
@@
 The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect time $\tau^*$ when any tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.
-We require that the initial $T_0$ frames are without tampering and provided for training.
+
+%Per $t<T^*$ abbiamo che i frame vengono acquisiti in condizioni di funzionamento normale:
+%\[ z_t(x)=y_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t=1,\dots , T^*-1. \] 
+%All'istante di tempo $t = T^*$ avviene un evento di tampering, il quale compromette i frame per $t\geq T^*$.  
+%In particolare, nel caso di una sfocatura avremo
+%\[ z_t = \mathcal{B}_t[y_t](x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*,\]
+%mentre nel caso di uno spostamento della camera avremo
+%\[ z_t = w_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*. \]
 
@@
 
-Algorithm \ref{alg:DISPL} presents the proposed tampering detection, which relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$).
-The former is meant to detect camera displacement, the latter camera defocus.
-As anticipated in Section \ref{sec:introduction}, during the initial configuration, the scene is segmented in $\{R_k\}, k = 1,\dots,K$ disjoint regions ($R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$).
-Details about the employed segmentation algorithm are in Section \ref{subsec:Segmentation}.
-
-In each frame $z_t$ the luma is separately computed in of the $K$ regions,
-\begin{equation}\label{eq:lumaRegions}
-l_k(t) =\frac{1}{\#R_k} \sum_{R_k} z_t(x), \ \ k = 1, \dots, K\,,
-\end{equation}
-while the average gradient norm is computed over the whole image
-\begin{equation}
-\label{eq:normaGradiente}
-g(t) = \sum_X \left (\sqrt{\left(z_t \circledast f_h\right)^2(x) + \left(z_t \circledast f_v\right)^2(x)}\right),
-\end{equation}
-where $f_h$ and $f_h$ are the horizontal and vertical derivative filters, respectively, $\#(\cdot)$ indicates the cardinality of a set and $\circledast$ the 2d convolution.
-
-As shown in Figure REFERENCE\_TODO.a), tampering results in outliers in the sequence of indicators.
-Detecting these outliers is not straightforward, since indicators heavily fluctuates in normal conditions because of short-term variations in the depicted scene or illumination changes.
-For outlier-detection purposes~\cite{Chandola2009}, it is more convenient to operate on sequences of independent and identically distributed (i.i.d.) values, where outliers can be detected by density-based methods or simply by confidence regions that is what we considered in our algorithm.
-Unfortunately, trends or deterministic variations in the indicator sequences are not easy to compensate because of unpredictable evolutions of the scene.
-
-To obtain indicators that in normal conditions can be treated as i.i.d. realizations of a random variable we detrend~\cite{Gustafsson2000} the indicator sequence by performing temporal derivative
-\begin{equation}\label{eq:detrending}
- \partial l_k(t) = l_k(t)-l_k(t-1),  \ \ k = 1, \dots, K\,,
-\end{equation}
-and we similarly define $\partial g(t)$.
-Detrended indicators are more suitable to be monitored by a confidence interval since, as shown in Figure REFERENCE\_TODO.b).
-However, a tampering yielding permanent changes in the indicators, as those in Figure REFERENCE\_TODO.a results in a single outlier in the detrended indicator~\eqref{eq:detrending}.
+The proposed tampering-detection algorithm relies on two indicators that are very easy to compute: the average intensity (also refer to \emph{luma} and denoted by $l$) and the average norm of the gradient (denoted by $g$).
+To analyze an incoming frame, we compute these indicators at first and then detect possible tampering attacks by determining whether these falls outside a confidence interval defined during an initial training phase.
 
@@
 	\SetNlSty{small}{}{.}
-	\textbf{Input}: $\gamma$, $T_{o}$, $\{R_k\} k=1,\dots,K$ \\
+	\textbf{Input}: $\gamma$, $T_{o}$, $\{R_k\}, k=1,\dots,K$ \\
 	\textbf{Configuration}:\\
-	\lnl{DISPL-Tr1} Compute $\partial l_k(t)$ and $\partial g(t), t = 1, \cdots, T_o$\\ 
-	\lnl{DISPL-Tr7} Compute $\sigma_{l_k}$ and $\sigma_{g}$, the standard deviations of $\partial l_k$ and $\partial g$.\\
-
+%	\lnl{DISPL-Tr1} Extract regions $\{\mathcal{R}_k\}, k=1,\dots,K$  \\
+	\lnl{DISPL-Tr2} \For{$t=1,\dots,T_{o}$}
+	{	\lnl{DISPL-Tr3} Get frame $z_t$ \\
+		\lnl{DISPL-Tr4} \For{$k=1,\dots,K$}{
+			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
+		}
+	}
+	\lnl{DISPL-Tr6} \For{$k=1,\dots,K$}{
+		\lnl{DISPL-Tr7} Compute $\sigma_l^k$\\
+	}
 	\textbf{Operational phase}:\\
-	\lnl{DISPL-Test1} \For{$t=T_{o} + 1,\dots,\infty$}{
-		\lnl{DISPL-Test2} Get frame $z_t$, set $n_l =0$
+	\lnl{DISPL-Test1} \For{$t=T_{o},\dots,\infty$}{
+		\lnl{DISPL-Test2} Get frame $z_t$ \\
+		\lnl{DISPL-Test3} $n_l = 0$ \\
 		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
-			\lnl{DISPL-Test5} Compute $\partial l_k(t)$\\
-			\lnl{DISPL-Test6} \If{$\partial l_k(t) < -\gamma \sigma_{l_k} \vee \partial l_k(t) > \gamma \sigma_{l_k} $}{
+			\lnl{DISPL-Test5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
+			\lnl{DISPL-Test6} \If{$\partial l^k(t) < -\gamma \sigma_l^k \vee \partial l^k(t) > \gamma \sigma_l^k $}{
 				\lnl{DISPL-Test7} $n_l=n_l+1$\\
@@
 		\lnl{DISPL-Test8} \If{$n_l\geq K-1$}{
-			\lnl{DISPL-Test9} tampering (probably displacement or occlusion) has occurred in $z_t$ \\
+			\lnl{DISPL-Test9} $z_t$ is a tampered frame\\
 		}
-		\lnl{DISPL-Test10} Compute $\partial g(t)$\\
-		\lnl{DISPL-Test11} \If{$\partial g(t) < -\gamma \sigma_{g} \vee \partial l(t) > \gamma \sigma_{g} $}{
-				\lnl{DISPL-Test9} tampering (probably out-of-focus) has occurred in $z_t$ \\
 	}   
-
-	}   
-	\caption{The Proposed Tampering-Detection Algorithm}
+	\caption{Tampering detection algorithm}
 	\label{alg:DISPL}
@@
 \end{itemize}
+%
+%
+%
+\subsection{Indicators}\label{subsec:Indicators}
+
+\begin{equation}
+\label{eq:gradientRegions}
+\begin{array}{ccc}
+g^k(t)&  = & \mathcal{G}^k[z_t] = \frac{\sum_{R_k}\| \nabla z_t(x) \| _2^2 }{|{R_k}|}\\
+\partial g^k(t) & =& g^k(t)-g^k(t-1) 
+\end{array},
+\end{equation}
+
+In particolare, per il calcolo delle derivate orizzontali  abbiamo utilizzato il seguente filtro $f_h$:
+\[f_h = f \circledast \left[ \begin{array}{rcl}
+1 & 0 & -1
+\end{array}\right], \] 
+mentre per il calcolo delle derivate verticali abbiamo utilizzato il seguente filtro $f_v$:
+\[f_v = f \circledast \left[ \begin{array}{r}
+1 \\ 0 \\ -1
+\end{array}\right], \]
+dove abbiamo indicato con $\circledast$ l'operatore di convoluzione.
+Il filtro $f$, invece, \`e ottenuto tramite un campionamento della \textit{funzione gaussiana} $h$, con media $0$ e deviazione standard $\sigma$
+\begin{equation}
+\label{eq:gaussian}
+h(i,j)=\frac{1}{2\pi\sigma^2}\exp\left(-\frac{i^2+j^2}{2\sigma^2}\right),
+\end{equation}
+e ponendo il valore massimo di questa funzione nel centro del filtro.
+Con questi filtri \`e possibile calcolare la \textit{norma del gradiente} nel seguente modo:
+\begin{equation}
+\label{eq:normaGradiente}
+\| \nabla z_t(x) \|_2^2=\left(z_t \circledast f_h\right)(x)^2 + \left(z_t \circledast f_v\right)(x)^2.
+\end{equation}
+Una volta calcolata la norma del gradiente \`e possibile farne la media come specificato in \eqref{eq:energyGradient}.
+Il risultato finale \`e un indicatore \textit{scalare} per ciascun frame acquisito, che pu\`o essere monitorato per individuare eventi di sfocature. 
+In particolare ci aspettiamo che l'evento di sfocatura provochi un abbattimento del valore di $g$.
+
+\begin{equation}
+\label{eq:lumaRegions}
+\begin{array}{ccc}
+l^k(t)&  = & \mathcal{L}^k[z_t] = \frac{\sum_{R_k} z_t(x) }{|{R_k}|}\\
+\partial l^k(t) & =& l^k(t)-l^k(t-1) 
+\end{array},
+\end{equation}
 
@@
 %\end{equation}
+
+FRAME DIFFERENCE:
+\begin{equation}
+	\label{eq:frameDiffReg}
+	\varphi^k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
+\end{equation}
 
@@
 \end{itemize}
-
-FRAME DIFFERENCE:
-\begin{equation}
-	\label{eq:frameDiffReg}
-	\varphi^k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
-\end{equation}
-
 %
@@
 \end{document}
-
-%
-%
-%
-\subsection{Indicators}\label{subsec:Indicators}
-
-%\begin{equation}
-%\label{eq:gradientRegions}
-%\begin{array}{ccc}
-%g^k(t)&  = & \mathcal{G}^k[z_t] = \frac{\sum_{R_k}\| \nabla z_t(x) \| _2^2 }{|{R_k}|}\\
-%\partial g^k(t) & =& g^k(t)-g^k(t-1) 
-%\end{array},
-%\end{equation}
-%
-%In particolare, per il calcolo delle derivate orizzontali  abbiamo utilizzato il seguente filtro $f_h$:
-%\[f_h = f \circledast \left[ \begin{array}{rcl}
-%1 & 0 & -1
-%\end{array}\right], \] 
-%mentre per il calcolo delle derivate verticali abbiamo utilizzato il seguente filtro $f_v$:
-%\[f_v = f \circledast \left[ \begin{array}{r}
-%1 \\ 0 \\ -1
-%\end{array}\right], \]
-%dove abbiamo indicato con $\circledast$ l'operatore di convoluzione.
-%Il filtro $f$, invece, \`e ottenuto tramite un campionamento della \textit{funzione gaussiana} $h$, con media $0$ e deviazione standard $\sigma$
-%\begin{equation}
-%\label{eq:gaussian}
-%h(i,j)=\frac{1}{2\pi\sigma^2}\exp\left(-\frac{i^2+j^2}{2\sigma^2}\right),
-%\end{equation}
-%e ponendo il valore massimo di questa funzione nel centro del filtro.
-%Con questi filtri \`e possibile calcolare la \textit{norma del gradiente} nel seguente modo:
-%
-%Una volta calcolata la norma del gradiente \`e possibile farne la media come specificato in \eqref{eq:energyGradient}.
-%Il risultato finale \`e un indicatore \textit{scalare} per ciascun frame acquisito, che pu\`o essere monitorato per individuare eventi di sfocature. 
-%In particolare ci aspettiamo che l'evento di sfocatura provochi un abbattimento del valore di $g$.

### 2015-05-26 paper.tex
...

--- 2015-05-26 paper.tex
@@
 
-As shown in Figure REFERENCE\_TODO.a), tampering results in outliers in the sequence of indicators.
+\begin{figure}[tb]
+\centering
+\label{fig:indicatori} 
+\includegraphics[width=1\linewidth]{Immagini/indicatori}
+\caption{Behavior of the average norm of the gradient \textbf{(a)} and its detrending \textbf{(b)} in a sequence with a defocus event.
+The tampering event come at $t = 600$.
+As illustrated in \textbf{(a)}, even in the case in which there aren't tampering events, there is nonstationarity.
+Detrending is done in order to preserve stationarity in normal situations.
+Considering the detrending \textbf{(b)} persistent changes become instantaneous:
+in particular there is a peak in the instant in which tampering begins.}
+\end{figure}
+
+As shown in Figure \ref{fig:indicatori}(a), tampering results in outliers in the sequence of indicators.
 Detecting these outliers is not straightforward, since indicators heavily fluctuates in normal conditions because of short-term variations in the depicted scene or illumination changes.

### 2015-05-26 paper.tex

--- 2015-05-26 paper.tex
@@
 
+\begin{equation}
+\label{eq:lumaRegions}
+\begin{array}{ccc}
+l^k(t)&  = & \mathcal{L}^k[z_t] = \frac{\sum_{R_k} z_t(x) }{|{R_k}|}\\
+\partial l^k(t) & =& l^k(t)-l^k(t-1) 
+\end{array},
+\end{equation}
+
+%\begin{equation}
+%\label{eq:energyLuma}
+%l(t) = \mathcal{L}[z_t] =\frac{\sum_{\mathcal{X}} z_t(x) }{|\mathcal{X}|} ,
+%\end{equation}  
+%
+%
+%\begin{equation}
+%\label{eq:gradientDetr}
+%\frac{\partial g}{\partial t}(t) = g(t) - g(t-1),
+%\end{equation}
+%
+%
+%\begin{equation}
+%\label{eq:lumaDetr}
+%\frac{\partial l}{\partial t}(t) = l(t) - l(t-1).
+%\end{equation}
+
+FRAME DIFFERENCE:
+\begin{equation}
+	\label{eq:frameDiffReg}
+	\varphi^k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
+\end{equation}
+
+\subsection{Outlier Detection}\label{subsec:MonitoringScheme}
+
+\begin{equation}
+\label{eq:soglieGradiente}
+\begin{array}{lcl}
+\Gamma_{min}^k & = & \hat{\mu}_g^k -\gamma \hat{\sigma}_g^k\\
+\Gamma_{max}^k & = & \hat{\mu}_g^k + \gamma \hat{\sigma}_g^k
+\end{array},
+\end{equation}
+dove $\hat{\mu}_g^k$ indica il valore medio delle osservazioni del training set
+\begin{equation}
+\hat{\mu}_g^k = \frac{\sum_{\tau = 1}^{T_{o}} \frac{\partial g^k}{\partial t}(\tau)}{T_{o}}, \nonumber
+\end{equation}
+$\hat{\sigma}_g^k$ indica la deviazione standard delle osservazioni del training set
+\begin{equation}
+\hat{\sigma}_g^k  = \sqrt{\frac{1}{T_{o}-1}\sum_{\tau=1}^{T_{o}}\left(\frac{\partial g^k}{\partial t}(\tau) - \hat{\mu}_g^k(\tau)\right)^2} \nonumber
+\end{equation}
+e $\gamma>1$ \`e un parametro moltiplicativo ottenuto sperimentalmente.\\
+
+\begin{equation}
+\label{eq:soglieLuma}
+\begin{array}{rcl}
+\Gamma_{min}^k & = & \hat{\mu}_l^k -\gamma \hat{\sigma}_l^k\\
+\Gamma_{max}^k & = & \hat{\mu}_l^k + \gamma \hat{\sigma}_l^k
+\end{array},
+\end{equation}
+dove $\hat{\mu}_l^k$ indica il valore medio delle osservazioni del training set
+\begin{equation}
+\hat{\mu}_l^k = \frac{\sum_{\tau = 1}^{T_{o}} \frac{\partial l^k}{\partial t}(\tau)}{T_{o}}, \nonumber
+\end{equation}
+$\hat{\sigma}_l^k$ indica la deviazione standard delle osservazioni del training set
+\begin{equation}
+\hat{\sigma}_l^k  = \sqrt{\frac{1}{T_{o}-1}\sum_{\tau=1}^{T_{o}}\left(\frac{\partial l^k}{\partial t}(\tau) - \hat{\mu}_l^k(\tau)\right)^2} \nonumber
+\end{equation}
+e $\gamma>1$ \`e un parametro moltiplicativo ottenuto sperimentalmente.\\
+
+\subsection{Algorithm Summary}\label{subsec:AlgorithmSummary}
+
+%\begin{algorithm}[tp]
+%	% \SetAlgoNoLine
+%	\LinesNumbered
+%	\SetAlgoNlRelativeSize{0}
+%	\SetNlSty{small}{}{.}
+%	\textbf{Configuration}:\\
+%	\lnl{DEF-Tr0} Extract regions $\{R_k\}, k=1,\dots,K$  \\
+%	\lnl{DEF-Tr1} \For{$t=1,\dots,T_{o}$}
+%	{	\lnl{DEF-Tr2} Acquire frame $z_t$ \\
+%		\lnl{DEF-Tr2a} \For{$k=1,\dots,K$}{
+%			\lnl{DEF-Tr3a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
+%		}
+%		\lnl{DEF-Tr3} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
+%	}
+%	\lnl{DEF-Tr6a} \For{$k=1,\dots,K$}{
+%		\lnl{DEF-Tr7a} Define thresholds $\Gamma_{min}^k$ and $\Gamma_{max}^k$\\
+%	}
+%	\lnl{DEF-Tr5} Define CDT parameters on $g(t)$ variance\\
+%	\textbf{Operational phase}:\\
+%	\lnl{DEF-Test1} \For{$t=T_{o},\dots,\infty$}{
+%		\lnl{DEF-Test2} Acquire frame $z_t$ \\
+%		\lnl{DEF-Tr3b} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
+%		\lnl{DEF-Test2a} $n=0$\\
+%		\lnl{DEF-Test4a} \For{$k=1,\dots,K$}{
+%			\lnl{DEF-Test5a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
+%			\lnl{DEF-Test6a} \If{$\frac{\partial g^k}{\partial t}(t) < \Gamma_{min}^k \vee \frac{\partial g^k}{\partial t}(t) > \Gamma_{max}^k $}{
+%				\lnl{DEF-Test7a} $n=n+1$\\
+%			}
+%		}
+%		\lnl{DEF-Test81} \If{$n \geq K-1$}{
+%			\lnl{DEF-Test91} $z_t$ is a defocused frame\\
+%		}
+%		\lnl{DEF-Test8} \If{CDT detect a change on $g(t)$ variance}{
+%			\lnl{DEF-Test9} $z_t$ is a defocused frame\\
+%		}
+%	}   
+%	\caption{Blur detection algorithm}
+%	\label{alg:DEFOCUS}
+%\end{algorithm}
+
+%\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
+
 \section{Experiments}\label{sec:experiments}

### 2015-05-26 paper.tex
Aggiunta immagine indicatori

@@
 and we similarly define $\partial g(t)$.
-Detrended indicators are more suitable to be monitored by a confidence interval since, as shown in Figure REFERENCE\_TODO.b).
-However, a tampering yielding permanent changes in the indicators, as those in Figure REFERENCE\_TODO.a results in a single outlier in the detrended indicator~\eqref{eq:detrending}.
+Detrended indicators are more suitable to be monitored by a confidence interval since, as shown in Figure \ref{fig:indicatori}(b).
+However, a tampering yielding permanent changes in the indicators, as those in Figure \ref{fig:indicatori}(a) results in a single outlier in the detrended indicator~\eqref{eq:detrending}.
+
+\begin{figure}[tb]
+\centering
+\label{fig:indicatori} 
+\includegraphics[width=1\linewidth]{Immagini/indicatori}
+\caption{Behavior of the average norm of the gradient \textbf{(a)} and its detrending \textbf{(b)} in a sequence with a defocus event.
+The tampering event come at $t = 600$.
+As illustrated in \textbf{(a)}, even in the case in which there aren't tampering events, there is nonstationarity.
+Detrending is done in order to preserve stationarity in normal situations.
+Considering the detrending \textbf{(b)} persistent changes become instantaneous:
+in particular there is a peak in the instant in which tampering begins.}
+\end{figure}
 

### 2015-05-27 paper.tex
Messa immagine indicatori nel paper
sistemata reference a figura 2

@@
 
-As shown in Figure REFERENCE\_TODO.a), tampering results in outliers in the sequence of indicators.
+\begin{figure}[tb]
+\centering
+\includegraphics[width=1\linewidth]{Immagini/indicatori}
+\caption{Behavior of the average norm of the gradient \textbf{(a)} and its detrending \textbf{(b)} in a sequence with a defocus event.
+The tampering event come at $t = 600$.
+As illustrated in \textbf{(a)}, even in the case in which there aren't tampering events, there is nonstationarity.
+Detrending is done in order to preserve stationarity in normal situations.
+Considering the detrending \textbf{(b)} persistent changes become instantaneous:
+in particular there is a peak in the instant in which tampering begins.}
+\label{fig:indicatori} 
+\end{figure}
+
+As shown in Figure \ref{fig:indicatori}(a), tampering results in outliers in the sequence of indicators.
 Detecting these outliers is not straightforward, since indicators heavily fluctuates in normal conditions because of short-term variations in the depicted scene or illumination changes.
@@
 However, a tampering yielding permanent changes in the indicators, as those in Figure \ref{fig:indicatori}(a) results in a single outlier in the detrended indicator~\eqref{eq:detrending}.
-
-\begin{figure}[tb]
-\centering
-\label{fig:indicatori} 
-\includegraphics[width=1\linewidth]{Immagini/indicatori}
-\caption{Behavior of the average norm of the gradient \textbf{(a)} and its detrending \textbf{(b)} in a sequence with a defocus event.
-The tampering event come at $t = 600$.
-As illustrated in \textbf{(a)}, even in the case in which there aren't tampering events, there is nonstationarity.
-Detrending is done in order to preserve stationarity in normal situations.
-Considering the detrending \textbf{(b)} persistent changes become instantaneous:
-in particular there is a peak in the instant in which tampering begins.}
-\end{figure}
 
@@
 The first one refers to frame sequences taken from webcams recording some parts of cities (as in Figure \ref{fig:sequences}(a)), where we have introduced some synthetically generated tampering events:
-defocus has been simulated using gaussian filters (\ref{fig:sequences}(b)), while displacement has been created with concatenation of similar frame sequences (Figure \ref{fig:sequences}(c)).
+defocus has been simulated using gaussian filters (Figure \ref{fig:sequences}(b)), while displacement has been created with concatenation of similar frame sequences (Figure \ref{fig:sequences}(c)).
 In some cases these sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\

### 2015-05-27 paper.tex
Messe le ROC nuove

@@
 
+\begin{figure}[t!]
+\centering
+\subfigure[]{\label{fig:pioggia} \includegraphics[width=0.4\linewidth]{Immagini/pioggia}}
+\subfigure[]{\label{fig:neve} \includegraphics[width=0.4\linewidth]{Immagini/neve}}
+\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena. \textbf{(a)} Blur due to rain drops on the camera lens. \textbf{(b)} Occlusion due to snow on the camera lens.}
+\label{fig:tampering}
+\end{figure}
+
 % Why is it interesting and important?
@@
 Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.
-
-\begin{figure}[t!]
-\centering
-\subfigure[]{\label{fig:pioggia} \includegraphics[width=0.4\linewidth]{Immagini/pioggia}}
-\subfigure[]{\label{fig:neve} \includegraphics[width=0.4\linewidth]{Immagini/neve}}
-\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena. \textbf{(a)} Blur due to rain drops on the camera lens. \textbf{(b)} Occlusion due to snow on the camera lens.}
-\label{fig:tampering}
-\end{figure}
 
@@
 %\gi{Adriano: metti entrambe le ROC curves, affiancate e per bene ed alcuni esempi di sequenze}
-\begin{figure}[htb]
+\begin{figure}[t]
 \centering
-\subfigure[]{\label{fig:ROCdisplacement_luma}\includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement_luma}}
-\subfigure[]{\label{fig:ROCdisplacement_fd} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement_fd}}
+\subfigure[]{\label{fig:ROCdisplacement}\includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement}}
+\subfigure[]{\label{fig:ROCdefocus} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus}}
 \caption{ROC curves for displacement detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean luma energy. \textbf{(b)} Analysis of the frame differencing.}
-\label{fig:ROCdisplacement}
+\label{fig:ROC}
 \end{figure}
-\begin{figure}[htb]
-\centering
-\subfigure[]{\label{fig:buenosAiresDef}\includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus1}}
-\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus_fd}}
-\caption{ROC curves for defocus detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean gradient energy. \textbf{(b)} Analysis of the frame differencing.}
-\label{fig:ROCdisplacement}
-\label{fig:ROCdefocus}
-\end{figure}
+%\begin{figure}[htb]
+%\centering
+%\subfigure[]{\label{fig:buenosAiresDef}\includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus1}}
+%\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus_fd}}
+%\caption{ROC curves for defocus detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean gradient energy. \textbf{(b)} Analysis of the frame differencing.}
+%\label{fig:ROCdisplacement}
+%\label{fig:ROCdefocus}
+%\end{figure}
 Experimental results are shown in Figures \ref{fig:ROCdisplacement} and \ref{fig:ROCdefocus}.
@@
 The algorithm uses low computational techniques: the biggest effort is made by computation of the feature $g(t)$, which needs $34$ operations per pixel, due to the Sobel filter used during the convolution.
-
+The computational load could be decreased using Fast Fourier Transform (FFT) \cite{brigham1988fast} and operating in the frequency domain.
 %Numero di operazioni richieste:

### 2015-05-27 paper.tex
edits sostanziali a sezione 1 e 3. Mancano da sistemare - Segmentation (impostata) - Related works( da condensare) - experiments (impostati, da rifinire)

@@
 % Why is it interesting and important?
-Tampering prevents the correct scene interpretation and causes the loss of small, important details such as licence plates.
-Tampering detection is therefore an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly reporting alerts.
+Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly reporting alerts.
+In fact, tampering prevents the correct scene interpretation: even a mild blur might result in important details such as licence plates.
 Surveillance cameras are typically connected to the power supply and operates at normal frame-rates (e.g.
-around few frames per second): in these conditions, several tampering-detection algorithms have been presented~\gi{Adriano add REFERENCES}.
+around few frames per second): in these conditions, several tampering-detection algorithms have been presented~\cite{TODO}.
 
-\gi{Qui ST pu\'o aggiungere qualche esempio di applicazione qui o menzionare dispositivi di riferimento (con tanto di link a datasheet).
-Magari aggiungendo che le batterie recenti permettono operativit\'a di due anni a questi regimi.}\cl{Verifico cosa possiamo dire del SecSoc.}
-%
-In this work we expressly target low-power and ultra-low-power smart cameras, devices that operate at very low frame-rates (e.g.
-possibly less than one frame every minute), that are characterized by a constrained computational power and memory, and that are battery powered (typically lasting one or two years).
-As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where nodes are wirelessy connected smart-cameras, that can acquire and transmit images at regular time interval or upon requests.
-Often, in WMSN, the units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting mechanisms~\gi{REFERENCES, Adriano, guarda se c'e' qualche reference dal nostro paper di TIM} \cite{magno2009adaptive}.
-Therefore, the power consumption is a serious concern in these devices.
-
-Even though low-power and ultra-low-power cameras not employed in critical surveillance applications, these devices are becoming popular in distributed systems for monitoring wide environments due to their low cost and maintenance requirements \gi{Adriano, Claudio, possiamo mettere qualche REFERENCE?}.
-Tampering detection is therefore very important in low-power smart cameras, where the algorithm are expected to be particularly prompt and have low false-alarm rates to prevent unnecessary energy-demanding operations like the local processing or radio activation for transmitting corrupted images (or false alarms) over the network.
-Unfortunately, tampering detection in such a resource-constrained scenario, has been much less investigated.
+In this work we expressly target low-power and ultra-low-power smart cameras\gi{una reference qua ci vuole}, battery-powered devices that are characterized by a constrained computational power and memory availability.
+Even though these devices are not meant for critical surveillance applications, they are becoming popular in distributed systems for monitoring wide environments due to their low cost and maintenance requirements.
+%As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where smart-cameras can acquire and transmit images at regular time interval or upon requests.
+% These units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting~\cite{magno2009adaptive}. 
+%In practice, because of energy constrain, low-power devices have to operate at very low frame-rates: possibly less than one frame every minute.
 
 % Why is it hard? (E.g., why do naive approaches fail?)
-Tampering detection in low-power smart cameras is much more challenging than in conventional surveillance cameras~\cite{perrig2004security}.
+Tampering detection in low-power smart cameras has not been much investigated in the literature~\cite{alippi2010detecting} while being more challenging than in conventional surveillance cameras~\cite{perrig2004security}.
 Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), thus the acquired sequence does not evolve smoothly.
 This prevents the use of learned background models and the analysis of foreground variations \cite{piccardi2004background}.
-When dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions, resulting in sequences those depicted in Figure~\ref{fig:sequences}: smart cameras have to correctly distinguish between these \emph{normal} changes and changes due to camera tampering.
-
+In fact, when dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions, yielding sequences like those in Figure~\ref{fig:sequences}.
+Smart cameras have to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames.
+Moreover, the false-alarm rate is a serious concern, a any false alarm results in wasting energy for data processing and network transmissions.
 % We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose.
@@
 %What are the key components of my approach and results? Also include any specific limitations.
+\gi{AGGIUNGERE qualcosa su one-shot monitoring}
 We present an algorithm to detect both camera defocus and displacement, as stated in Section \ref{sec:probForm}.
 The algorithm presented in Section~\ref{sec:propSol} relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blur.
-These indicators can be easily computed and monitored by an outlier-detection technique to detect the respective tampering.
-We show that displacement-detection performance can be improved by means of a preliminarily partitioning of the scene into nonoverlapping regions, and then separately computing and monitoring the average intensity over different regions (Section \ref{subsec:Segmentation}).
-Remarkably, these regions need to be defined during an initial configuration phase of the smart camera, thus there is no relevant computational overhead w.r.t.
+These indicators can be easily computed and monitored by an outlier-detection technique to detect the corresponding tampering.
+We show that displacement-detection performance can be improved by preliminarily partitioning the scene into nonoverlapping regions, and then separately computing and monitoring the average intensity over different regions (Section \ref{subsec:Segmentation}).
+Remarkably, these regions can be defined during an initial configuration phase, thus separately monitoring different regions have a negligible computational overhead w.r.t.
 monitoring the whole image.
 Experiments in Section \ref{sec:experiments} show that operating on image regions can substantially improve the detection of camera displacements.
-Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.
+% Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.
 
@@
 %
-Let $z_t$ be the frame acquired at time $t$, which we model as
+Let $z_t$ be the frame acquired at time $t$
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{D}_t$ denotes the degradation operator that transforms the original image $y_t$ in the frame $z_t$; $X \subset \mathbb{Z}^2$ denotes the regular pixels grid and $x\in \mathbb{Z}^2$ indicates the pixel coordinates.
+where $\mathcal{D}_t$ denotes an operator that transforms the original image $y_t$ in the acquired frame $z_t$; $X \subset \mathbb{Z}^2$ denotes the regular pixels grid and $x\in \mathbb{Z}^2$ indicates the pixel coordinates.
 As far as there is no tampering attacks/events,
@@
 \end{equation}
-where $\eta_t$ is a random variable accounting for different noise sources (e.g., thermal, quantization, photon-counting).
-In normal conditions, all the images $y_t$ (thus also the frames $z_t$) might show different content but are acquired from the same viewpoint and the same camera orientation.
+where $\eta_t$ is a random variable accounting for image noise, %sources (e.g., thermal, quantization, photon-counting). 
+and all the images $y_t$ might show different content but are acquired from the same viewpoint and the same camera orientation.
 
-When at time $\tau^*$ an external disturbance introduces \emph{blur/defocus}, the image $y_t$ is degraded by a spatially variant blur operator, and $z_t$ becomes
+When, at time $\tau^*$, an external disturbance introduces \emph{blur/defocus}, the image $y_t$ is degraded by an unknown blur operator, and $z_t$ becomes
 \begin{equation}
@@
 The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect time $\tau^*$ when any tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.
-We require that the initial $T_0$ frames are without tampering and provided for training.
+We assume that $T_0$ tampering-free frames and provided for training.
 
-\section{The Proposed Algorithm}\label{sec:propSol}
+\section{Tampering Detection}\label{sec:propSol}
 
-Algorithm \ref{alg:DISPL} presents the proposed tampering detection, which relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$).
+Algorithm \ref{alg:DISPL} presents the proposed tampering-detection algorithm, which relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$).
 The former is meant to detect camera displacement, the latter camera defocus.
-As anticipated in Section \ref{sec:introduction}, during the initial configuration, the scene is segmented in $\{R_k\}, k = 1,\dots,K$ disjoint regions ($R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$).
-Details about the employed segmentation algorithm are in Section \ref{subsec:Segmentation}.
+As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$.
+The employed segmentation is detailed in Section \ref{subsec:Segmentation}.
 
-In each frame $z_t$ the luma is separately computed in of the $K$ regions,
+For each $z_t$, the luma is separately computed within each of the $K$ regions:
 \begin{equation}\label{eq:lumaRegions}
@@
 \end{equation}
-where $f_h$ and $f_h$ are the horizontal and vertical derivative filters, respectively, $\#(\cdot)$ indicates the cardinality of a set and $\circledast$ the 2d convolution.
+where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, $\circledast$ denotes the 2d convolution and $\#(\cdot)$ the cardinality of a set.
 
@@
 
-As shown in Figure \ref{fig:indicatori}(a), tampering results in outliers in the sequence of indicators.
-Detecting these outliers is not straightforward, since indicators heavily fluctuates in normal conditions because of short-term variations in the depicted scene or illumination changes.
-For outlier-detection purposes~\cite{Chandola2009}, it is more convenient to operate on sequences of independent and identically distributed (i.i.d.) values, where outliers can be detected by density-based methods or simply by confidence regions that is what we considered in our algorithm.
-Unfortunately, trends or deterministic variations in the indicator sequences are not easy to compensate because of unpredictable evolutions of the scene.
-
-To obtain indicators that in normal conditions can be treated as i.i.d. realizations of a random variable we, detrend~\cite{Gustafsson2000} the indicator sequence by a temporal derivative
+%Tampering changes the sequence of indicators:
+Figure \ref{fig:indicatori}(a) shows how a displacement \gi{TODO} affects the values of $g$.
+First of all, we observe that displaced frames yield \gi{WHAT} in the indicator sequence.
+Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here, since these are meant for independent and identically distributed (i.i.d.) random variables.
+In fact, normal changes in the scene or in the illumination yield indicators that follow an unpredictable (but deterministic) trend, beside some high-frequency fluctuations.
+Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by a temporal derivative (line~\ref{DISPL-Test5})
 \begin{equation}\label{eq:detrending}
@@
 and we similarly define $\partial g(t)$.
-Detrended indicators are more suitable to be monitored by a confidence interval since, as shown in Figure \ref{fig:indicatori}(b).
-However, a tampering yielding permanent changes in the indicators, as those in Figure \ref{fig:indicatori}(a) results in a single outlier in the detrended indicator~\eqref{eq:detrending}.
 
-We perform outlier detection by defining confidence regions for our indicator in tampering-free frames.
-In practice, since all these are scalar indicators, it is enough to compute the indicator mean ($\overline{\partial l}_k(t), k = 1,\dots,K$ and $\overline{\partial g}(t)$) and standard deviations ($\sigma_{l_k}, k = 1,\dots,K$ and $\sigma_g$) over the tampering-free frames provided for training, namely $z_t\,, t = 1, \dots, T_o$.
-Confidence regions are then defined as
+The sequence of detrended indicators can be suitably monitored by defining the following confidence interval: 
 \begin{equation}\label{eq:confidenceRegions}
- [\overline{\partial l}_k(t) - \gamma \sigma_{l_k}, \overline{\partial l}_k(t) + \gamma \sigma_{l_k}],  k = 1,\dots,K
+ [\overline{\partial l}_k(t) - \gamma_l \sigma_{l_k}, \overline{\partial l}_k(t) + \gamma_l \sigma_{l_k}],  k = 1,\dots,K
 \end{equation}
-where $\gamma >0$ is a tuning parameter.
-A similar region is built for $\partial g$.
-During operations, any indicator falling outside the corresponding confidence region is considered anomalous and tampering is detected.
+where $\overline{\partial l}_k(t)$ denotes the mean and $\sigma_{l_k}$ the standard deviation of $\partial l$ over $R_k$, computed from the tampering-free frames provided for training (i.e.
+$z_t\,, t = 1, \dots, T_o$) and $\gamma_l >0$ is a tuning parameter.
+A similar interval is built for $\partial g$ (lines \ref{DISPL-Tr7}).
+
+During operations, any indicator falling outside its confidence region is considered anomalous (line~\ref{DISPL-Test6}).
+To raise a camera-displacement alert we require at least $K-1$ indicators $\partial l_k$ to yield an outlier (line~\ref{DISPL-Test8}).
+This threshold was set to allow a single region not to change, as it typically happens in the sky when the camera is displaced, either horizontally or vertically.
+In contrast, any outlier in $\partial g$ would yield a blurring\gi{probabilmente e' piu' corretto parlare di blurring invece che defocus..
+Adriano, riesci ad allineare?
+Io menzionerei anche le occlusions assieme al displacement, che dici?} alert (line~\ref{DISPL-Test11}).
+
+It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, only the first and last tampered frames correspond to an outlier in the detrended indicators, see Figure~\ref{fig:indicatori}(b).
+Therefore, detrended indicators have to be monitored in a one-shot manner, targeting the detection of the first-tampered frame.
+On the one hand, this one-shot solution can increase the detection promptness, and this is particularly important at the low frame-rates we consider.
+On the other hand, it disregards the fact that tampering might persist over several frames.
+To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, and possibly combined with the one-shot monitoring.
 
@@
 	\SetNlSty{small}{}{.}
-	\textbf{Input}: $\gamma$, $T_{o}$, $\{R_k\} k=1,\dots,K$ \\
+	\textbf{Input}: $\gamma_l, \gamma_g$, $\{z_t, t = 1, \dots, T_{0}\}$, $\{R_k \ k=1,\dots,K\}$ \\
 	\textbf{Configuration}:\\
 	\lnl{DISPL-Tr1} Compute $\partial l_k(t)$ and $\partial g(t), t = 1, \cdots, T_o$\\ 
-	\lnl{DISPL-Tr7} Compute $\sigma_{l_k}$ and $\sigma_{g}$, the standard deviations of $\partial l_k$ and $\partial g$.\\
+	\lnl{DISPL-Tr7} Compute $\partial l_k$, $\partial g$, $\sigma_{l_k}$ and $\sigma_{g}$.\\
 
@@
 	\lnl{DISPL-Test1} \For{$t=T_{o} + 1,\dots,\infty$}{
-		\lnl{DISPL-Test2} Get frame $z_t$, set $n_l =0$
+		\lnl{DISPL-Test2} Get frame $z_t$, set $n_l =0$;\\
 		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
 			\lnl{DISPL-Test5} Compute $\partial l_k(t)$\\
-			\lnl{DISPL-Test6} \If{$\partial l_k(t) < -\gamma \sigma_{l_k} \vee \partial l_k(t) > \gamma \sigma_{l_k} $}{
+			\lnl{DISPL-Test6} \If{$\partial l_k(t) < -\gamma_l \sigma_{l_k} \vee \partial l_k(t) > \gamma_l \sigma_{l_k} $}{
 				\lnl{DISPL-Test7} $n_l=n_l+1$\\
@@
 		\lnl{DISPL-Test10} Compute $\partial g(t)$\\
-		\lnl{DISPL-Test11} \If{$\partial g(t) < -\gamma \sigma_{g} \vee \partial l(t) > \gamma \sigma_{g} $}{
-				\lnl{DISPL-Test9} tampering (probably out-of-focus) has occurred in $z_t$ \\
+		\lnl{DISPL-Test11} \If{$\partial g(t) < -\gamma_g \sigma_{g} \vee \partial l(t) > \gamma_g \sigma_{g} $}{
+				\lnl{DISPL-Test9} tampering (probably blurring) has occurred in $z_t$ \\
 								}
@@
 \subsection{Scene Segmentation}\label{subsec:Segmentation}
-Consideriamo la sequenza $\{z_t\}$ di frame acquisiti dalla camera, con $t=1,\dots,T_{c}$.
-Per ciascun pixel $x\in\mathcal{X}$ calcoliamo un vettore $\textbf{d}(x)$ di $5$ elementi
+Scene has to be preliminarily segmented to define regions  that our algorithm takes as input.
+To this purpose, we need an external training sequence\gi{Dobbiamo dargli un nome} depicting the scene in its normal conditions without any tampering.
+Segmentation is performed by associating to each pixel $x\in\mathcal{X}$ a feature vector $\textbf{f}(x)\in \mathbb{R}^5$:
 \begin{equation}
 \label{eq:featureVector}
-\textbf{d}(x)=\left[r(x);c(x);\mu_{\nabla}(x);\sigma_{\nabla}(x);\bar{z}(x)\right], \textbf{d}(x) \in \mathbb{R}^5
+\textbf{f}(x)=\left[r(x);c(x);\overline{g}(x);\sigma_{g}(x);\bar{l}(x)\right]\,, \ \forall x \in X
 \end{equation}
-dove:
-\begin{itemize}
-	\item $r(x)$ rappresenta il numero di riga del pixel $x$.
-	\item $c(x)$ rappresenta il numero di colonna del pixel $x$.
-	\item $\mu_{\nabla}(x)$ rappresenta il valore del gradiente nel pixel $x$ mediato nel tempo:
-	\begin{equation}
-	\label{eq:segmentazioneGrad}
-	\mu_{\nabla}(x) = \frac{\sum_{t=1}^{T_c}(\|\nabla z_t\|_2^2 \circledast f)(x)}{T_c},
-	\end{equation}
-	dove abbiamo indicato con $\|\nabla z_t\|_2^2$ la norma del gradiente per l'immagine $z_t$, definita in \eqref{eq:normaGradiente}, e con $f$ il filtro gaussiano discreto derivato dal campionamento di \eqref{eq:gaussian}.
-	\item $\sigma_{\nabla}(x)$ rappresenta la deviazione standard nel tempo del gradiente nel pixel $x$:
-	\begin{equation}
-	\label{eq:segmentazioneVar}
-	\sigma_{\nabla}(x)=\sqrt{\frac{1}{T_c - 1}\sum_{t=1}^{T_c}\left(\left(\|\nabla z_t\|_2^2 \circledast f\right)(x)-\mu_{\nabla}(x)\right)^2}.
-	\end{equation}
-	\item $\bar{z}(x)$ rappresenta il valore della luma del pixel $x$ mediato nel tempo:
-	\begin{equation}
-	\label{eq:segmentazioneLuma}
-	\bar{z}(x)=\frac{\sum_{t=1}^{T_c}( z_t \circledast f)(x)}{T_c}.
-	\end{equation}
-\end{itemize}
+where $r(x)$ and $c(x)$ represents the row and columns of $x$, $\overline{g}(x)$ and is the gradient norm (smoothed in a neigbhor of $x$) averaged over the training sequence and $\sigma_g$ is the standard deviation of the gradient norm (smoothed in a neigbhor of $x$), computed overtime, and $\bar{l}(x)$ is the average of the gradient over time \gi{ma la deviazione standard della luma nel tempo?
+cosa faceva?}
 
-When the feature vectors are computed for all the pixels we start the segmentation phase.
-This is made using classical clustering techniques:
-in particular we use a modified version of the classical k-means algorithm \cite{han2006data}. 
-In this implementation, each feature vector belongs to a cluster if its \textit{weighted Minkowski distance} \cite{de2012minkowski} from the cluster's centroid is minimum with respect to the distance from the other centroids.
-The distance is weighted considering the standard deviation of each elements of the feature vectors inside each cluster \cite{kottke1994motion}, in order to consider, during the clustering, also the contribute of the fields which have low variance.
+Segmentation consists in clustering the feature vectors computed over the whole image by means a weighted version of the k-means algorithm~\cite{han2006data}.
+In particular, since the components of the feature vector might span very different ranges, the distance between a feature vector and the center of a cluster is measured as the \textit{weighted Minkowski distance}~\cite{de2012minkowski}\gi{..
+ho tolto questo, non e' uguale allo standard k-means il resto?
+from the cluster's centroid is minimum with respect to the distance from the other centroids}, which takes into account the standard deviation of each component for each cluster.
+
+inside each cluster \cite{kottke1994motion}, in order to consider, during the clustering, also the contribute of the fields which have low variance.
 \ad{Giacomo: prova a vedere se si capisce cosa voglio dire in queste frasi!!!}
 
-\begin{equation}
-\label{eq:lumaRegions}
-\begin{array}{ccc}
-l^k(t)&  = & \mathcal{L}^k[z_t] = \frac{\sum_{R_k} z_t(x) }{|{R_k}|}\\
-\partial l^k(t) & =& l^k(t)-l^k(t-1) 
-\end{array},
-\end{equation}
+The initial number of clusters is instead chosen by \gi{COME? e poi non ricordo se hai trovato il numero di clusters facendo il k-means con i pesi}
 
-%\begin{equation}
-%\label{eq:energyLuma}
-%l(t) = \mathcal{L}[z_t] =\frac{\sum_{\mathcal{X}} z_t(x) }{|\mathcal{X}|} ,
-%\end{equation}  
-%
-%
-%\begin{equation}
-%\label{eq:gradientDetr}
-%\frac{\partial g}{\partial t}(t) = g(t) - g(t-1),
-%\end{equation}
-%
-%
-%\begin{equation}
-%\label{eq:lumaDetr}
-%\frac{\partial l}{\partial t}(t) = l(t) - l(t-1).
-%\end{equation}
+Finally, some canonical morphological image-operations are executed to remove boundaries between different regions, and eventually too small regions.
+This defines the regions $\{R_k\,, \ k=1,\dots,K\}$.
+
+\subsubsection{Computational Complexity}
+
+The algorithm uses low computational techniques: the biggest effort is made by computation of the feature $g(t)$, which needs $34$ operations per pixel, due to the Sobel filter used during the convolution.
+The computational load could be decreased using Fast Fourier Transform (FFT) \cite{brigham1988fast} and operating in the frequency domain.
+
+Segmentation has not been considered because...
 
@@
 \end{equation}
-
-\subsection{Outlier Detection}\label{subsec:MonitoringScheme}
-
-\begin{equation}
-\label{eq:soglieGradiente}
-\begin{array}{lcl}
-\Gamma_{min}^k & = & \hat{\mu}_g^k -\gamma \hat{\sigma}_g^k\\
-\Gamma_{max}^k & = & \hat{\mu}_g^k + \gamma \hat{\sigma}_g^k
-\end{array},
-\end{equation}
-dove $\hat{\mu}_g^k$ indica il valore medio delle osservazioni del training set
-\begin{equation}
-\hat{\mu}_g^k = \frac{\sum_{\tau = 1}^{T_{o}} \frac{\partial g^k}{\partial t}(\tau)}{T_{o}}, \nonumber
-\end{equation}
-$\hat{\sigma}_g^k$ indica la deviazione standard delle osservazioni del training set
-\begin{equation}
-\hat{\sigma}_g^k  = \sqrt{\frac{1}{T_{o}-1}\sum_{\tau=1}^{T_{o}}\left(\frac{\partial g^k}{\partial t}(\tau) - \hat{\mu}_g^k(\tau)\right)^2} \nonumber
-\end{equation}
-e $\gamma>1$ \`e un parametro moltiplicativo ottenuto sperimentalmente.\\
-
-\begin{equation}
-\label{eq:soglieLuma}
-\begin{array}{rcl}
-\Gamma_{min}^k & = & \hat{\mu}_l^k -\gamma \hat{\sigma}_l^k\\
-\Gamma_{max}^k & = & \hat{\mu}_l^k + \gamma \hat{\sigma}_l^k
-\end{array},
-\end{equation}
-dove $\hat{\mu}_l^k$ indica il valore medio delle osservazioni del training set
-\begin{equation}
-\hat{\mu}_l^k = \frac{\sum_{\tau = 1}^{T_{o}} \frac{\partial l^k}{\partial t}(\tau)}{T_{o}}, \nonumber
-\end{equation}
-$\hat{\sigma}_l^k$ indica la deviazione standard delle osservazioni del training set
-\begin{equation}
-\hat{\sigma}_l^k  = \sqrt{\frac{1}{T_{o}-1}\sum_{\tau=1}^{T_{o}}\left(\frac{\partial l^k}{\partial t}(\tau) - \hat{\mu}_l^k(\tau)\right)^2} \nonumber
-\end{equation}
-e $\gamma>1$ \`e un parametro moltiplicativo ottenuto sperimentalmente.\\
-
-\subsection{Algorithm Summary}\label{subsec:AlgorithmSummary}
 
@@
 \section{Experiments}\label{sec:experiments}
-The proposed algorithm has been implemented in MATLAB, and has been tested on two datasets.
-The first one refers to frame sequences taken from webcams recording some parts of cities (as in Figure \ref{fig:sequences}(a)), where we have introduced some synthetically generated tampering events:
-defocus has been simulated using gaussian filters (Figure \ref{fig:sequences}(b)), while displacement has been created with concatenation of similar frame sequences (Figure \ref{fig:sequences}(c)).
-In some cases these sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\
+\gi{Per ultimo: The proposed algorithm has been implemented in MATLAB, }
+
+Experiments are meant to show the effectiveness of the proposed algorithm on real-world sequences, and in particular on assessing the advantages of using segmentation for detecting camera displacements.
+
+We have acquired two datasets to test our method.
+The first dataset contains \gi{how many?} sequences taken from webcams recording urban areas (as in Figure \ref{fig:sequences}(a)), where we have synthetically introduced blurring by means of a gaussian filters (Figure \ref{fig:sequences}(b)), and displacement by concatenating frames from sequences depicting similar environments (Figure \ref{fig:sequences}(c)).
+In some cases, sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\
 \begin{figure}[t]
@@
 \end{figure}
-The second dataset was taken using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or putting water on the camera.\\
-Four figures of merit have been suggested to assess the performance of the proposed algorithm:\\
+The second dataset was recorded using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or spraying water on the camera len.\gi{una footnote dicendo sequences can be released upon request?}
+
+Typical figures of merit for one-shot monitoring methods have been used to assess the performance of the proposed algorithm.
+In particular we compute the ROC curves varying $\gamma_l$ and $\gamma_g$ in a range of values.
+ROC curve reports :
+\[1-\text{SPECIFICITY}_\gamma = 1-\frac{\text{TN}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}=\frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma},\]
+while on th y-axis there is:
+\[\text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
+The construction of these curves permits us to make a comparison with respect to other methods.\gi{io le metterei solo in una equazione entrambe, e per 1-specificity solo la terza formula}
+
 \begin{description}
@@
 \end{description} 
-These indicators are computed varying the parameter $\gamma$, which defines the thresholds for the one-shot monitoring. 
-This permits us to generate \textit{ROC curves}, where on the x-axis there is:
-\[1-\text{SPECIFICITY}_\gamma = 1-\frac{\text{TN}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}=\frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma},\]
-while on th y-axis there is:
-\[\text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
-The construction of these curves permits us to make a comparison with respect to other methods.
+\gi{questo lo metterei solo nel testo sotto le formule sopra}
+
+\gi{io inizierei dai metodi con cui confrontarci}
 In particular the alternative approaches that we have considered working:
@@
 On the other side, as illustrated in Figure \ref{fig:ROCdefocus}, when monitoring indicators meant to detect blur/defocus, it is more convenient to consider the whole scene at once.
-\subsection{Discussion}
-\gi{Adriano: Aggiungi qua la complessit\'a computazionale}
-% Please add the following required packages to your document preamble:
-% \usepackage[table,xcdraw]{xcolor}
-% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+
 \begin{table}[tbh]
@@
 
-The algorithm uses low computational techniques: the biggest effort is made by computation of the feature $g(t)$, which needs $34$ operations per pixel, due to the Sobel filter used during the convolution.
-The computational load could be decreased using Fast Fourier Transform (FFT) \cite{brigham1988fast} and operating in the frequency domain.
 %Numero di operazioni richieste:

### 2015-05-28 paper.tex
aggiunta frase nell'intro

@@
 
-We here introduce a tampering-detection algorithm that has been specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme.
+We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme.
 Any frame yielding anomalous indicator is detected as a tampering attempt.
 Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements.
-Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, and at a very small computational complexity.
+Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity.
 %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
@@
 %What are the key components of my approach and results? Also include any specific limitations.
-\gi{AGGIUNGERE qualcosa su one-shot monitoring}
-We present an algorithm to detect both camera defocus and displacement, as stated in Section \ref{sec:probForm}.
-The algorithm presented in Section~\ref{sec:propSol} relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blur.
-These indicators can be easily computed and monitored by an outlier-detection technique to detect the corresponding tampering.
-We show that displacement-detection performance can be improved by preliminarily partitioning the scene into nonoverlapping regions, and then separately computing and monitoring the average intensity over different regions (Section \ref{subsec:Segmentation}).
-Remarkably, these regions can be defined during an initial configuration phase, thus separately monitoring different regions have a negligible computational overhead w.r.t.
+We here address the detection of camera blurring and displacement, which are presented in Section \ref{sec:probForm}.
+The proposed algorithm, detailed in Section~\ref{sec:propSol}, relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blurring.
+These indicators can be easily computed and monitored by an outlier-detection technique in a low-power smart camera, thus our solution is meant as a very prompt trigger to be possibly combined with other sequential monitoring techniques.
+In particular, we show that separately monitoring the indicators over different regions of the scene can substantially improve the displacement-detection performance.
+Remarkably, these regions can be preliminarily defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t.
 monitoring the whole image.
-Experiments in Section \ref{sec:experiments} show that operating on image regions can substantially improve the detection of camera displacements.
+Experiments in Section \ref{sec:experiments} show that leveraging image regions can substantially improve the displacement-detection performance.
 % Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.

### 2015-05-28 paper.tex
Aggiunti plot nuovi

@@
 \includegraphics[width=1\linewidth]{Immagini/indicatori}
-\caption{Behavior of the average norm of the gradient \textbf{(a)} and its detrending \textbf{(b)} in a sequence with a defocus event.
-The tampering event come at $t = 600$.
-As illustrated in \textbf{(a)}, even in the case in which there aren't tampering events, there is nonstationarity.
-Detrending is done in order to preserve stationarity in normal situations.
-Considering the detrending \textbf{(b)} persistent changes become instantaneous:
-in particular there is a peak in the instant in which tampering begins.}
+\caption{Example of displacement and behavior of the average value of the luma.
+The event comes at frame $157$ and the corresponding value of $l(t)$ is depicted with a red star.
+\textbf{(a)} Considering the average value of the luma in the whole scene.
+\textbf{(b)} and \textbf{(c)} Considering the average value of the luma in a specific region.}
 \label{fig:indicatori} 

### 2015-05-28 paper.tex
tolti i related

--- 2015-05-28 paper.tex
@@
 \begin{abstract}
-A desirable feature for smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene.
+A desirable feature in smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene.
 No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attacks (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures.
@@
 %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
-
-\keywords{tampering detection, defocus, displacement detection}
+\keywords{tampering detection, blurring detection, displacement detection}
 \end{abstract}
@@
 \section{Introduction}\label{sec:introduction}
-%\gi{Adriano: decidi che email tenere, non tutte e due :)}
-
-%We here address the problem of detecting tampering events/attacks in camera devices, and in particular those preventing the proper acquisition of the monitored scene.
-
 % What is the problem?
@@
 \ref{fig:neve}.
-Similarly, intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be surely useless for monitoring purposes.
+Similarly, intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be useless for monitoring purposes.
 We refer to these events/attacks as tampering.
@@
 % Why is it interesting and important?
-Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly reporting alerts.
-In fact, tampering prevents the correct scene interpretation: even a mild blur might result in important details such as licence plates.
+Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly report alerts.
+In fact, tampering prevents the correct scene interpretation: even a mild blur might hinder the identification of important details such as licence plates.
 Surveillance cameras are typically connected to the power supply and operates at normal frame-rates (e.g.
-around few frames per second): in these conditions, several tampering-detection algorithms have been presented~\cite{TODO}.
+around few frames per second): in these conditions, several tampering-detection algorithms have been presented.
 
-In this work we expressly target low-power and ultra-low-power smart cameras\gi{una reference qua ci vuole}, battery-powered devices that are characterized by a constrained computational power and memory availability.
+In particular, camera displacement and occlusions are typically detected by background models or frame comparison, while blurring is instead detected by monitoring the high-frequency component of images, as this is expected to drop.
+In ~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain.
+Two background models computed over different time intervals are used in~\cite{saglam2009real}.
+In~\cite{gil2007automatic}, tampering detection is performed by combining background subtraction together with edge detection and normalized cross-correlation.
+Background models and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while number of SURF~\cite{SURF} keypoints is used to detect blurring.
+A frame buffer rather than an explicit background is used~\cite{ribnick2006real}.
+All these solutions, including~\cite{ribnick2006real,harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.
+
+In this work we expressly target low-power and ultra-low-power smart cameras\gi{Claudio, riesci a mettere una reference qua?}, battery-powered devices that are characterized by a constrained computational power and memory availability.
 Even though these devices are not meant for critical surveillance applications, they are becoming popular in distributed systems for monitoring wide environments due to their low cost and maintenance requirements.
@@
 % Why is it hard? (E.g., why do naive approaches fail?)
-Tampering detection in low-power smart cameras has not been much investigated in the literature~\cite{alippi2010detecting} while being more challenging than in conventional surveillance cameras~\cite{perrig2004security}.
+Tampering detection in low-power smart cameras has not been much investigated in the literature~\cite{alippi2010detecting}, while being more challenging than in conventional surveillance cameras~\cite{perrig2004security}.
 Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), thus the acquired sequence does not evolve smoothly.
@@
 Smart cameras have to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames.
-Moreover, the false-alarm rate is a serious concern, a any false alarm results in wasting energy for data processing and network transmissions.
+Moreover, false alarms represent a serious problem in low-power smart cameras as these result in useless data processing and network transmissions, thus energy waste.
 % We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose.
@@
 %What are the key components of my approach and results? Also include any specific limitations.
-We here address the detection of camera blurring and displacement, which are presented in Section \ref{sec:probForm}.
-The proposed algorithm, detailed in Section~\ref{sec:propSol}, relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blurring.
-These indicators can be easily computed and monitored by an outlier-detection technique in a low-power smart camera, thus our solution is meant as a very prompt trigger to be possibly combined with other sequential monitoring techniques.
-In particular, we show that separately monitoring the indicators over different regions of the scene can substantially improve the displacement-detection performance.
+We address the detection of camera blurring and displacement, which are presented in Section \ref{sec:probForm}.
+The proposed algorithm, detailed in Section~\ref{sec:propSol}, relies on two indicators: the average image intensity is used to detect displacement, and the average gradient norm to detect blurring.
+These indicators can be easily computed and monitored by an outlier-detection technique in a low-power smart camera.
+In particular, we show that separately monitoring the indicators over different regions of the image can substantially improve the displacement-detection performance.
 Remarkably, these regions can be preliminarily defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t.
 monitoring the whole image.
+Thus, our solution represents as a prompt trigger that can be possibly combined with other sequential monitoring techniques.
 Experiments in Section \ref{sec:experiments} show that leveraging image regions can substantially improve the displacement-detection performance.
@@
 
-\subsection{Related Works}\label{subsec:relWorks}
-% Why hasn't it been solved before? (Or, what's wrong with previous proposed solutions? How does mine differ?)
-The literature concerning tampering detection is mostly focused on video surveillance applications and operates at few frames per second \gi{Adriano, \'e vero?}.
-Background models are typically leveraged to identify defocus and occlusions; in particular~\cite{aksay2007camera} performs defocus detection by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered.
-A background-subtraction technique is employed in~\cite{saglam2009real} to identify defocus, occlusions, and displacements.
-\gi{FIXME Comparison in the Fourier domain for defocus detection, histogram comparison for occlusion detection, comparison between current background and delayed background for displacement detection}.
-Background subtraction is also used in \cite{gil2007automatic} to detect defocus, occlusions, and displacements.
-\gi{FIXME Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection}.
-In contrast, no background models are used in \cite{alippi2010detecting}, where a sequential monitoring scheme based on a change-detection test is employed to detect changes in the average gradient energy of each frame to detect defocus due to external disturbances on the camera lens.
-\cite{ribnick2006real}: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
-\cite{kryjak2012fpga}: implementation in a FPGA of a solution based on background modeling, histograms comparisons, edges comparisons.
-\cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements.
+\begin{figure}[t]
+\centering
+\includegraphics[width=1\linewidth]{Immagini/sequenze}
+\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences; \textbf{(d)} real displacement event on 4-th frame.}
+\label{fig:sequences}
+\end{figure}
+
+%A low-complexity and sequential monitoring scheme over the average gradient norm is used in~\cite{alippi2010detecting} to detect blurring due to external disturbances on the camera lens.
+
+%\subsection{Related Works}\label{subsec:relWorks}
+%% The literature concerning tampering detection is mostly focused on video surveillance applications and process at few frames per second.
+% 
+%%
+%: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
+%\cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements.
 Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
-\cite{tsesmelis2013tamper}: monitoring of the number of key points extracted by SURF in order to detect defocus events, partition in blocks and HOG descriptors matching for each block in order to detect occlusions.
-These types of solutions requires a lot of computations
 %
@@
 In some cases, sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\
-\begin{figure}[t]
-\centering
-\includegraphics[width=1\linewidth]{Immagini/sequenze}
-\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences; \textbf{(d)} real displacement event on 4-th frame.}
-\label{fig:sequences}
-\end{figure}
+
 The second dataset was recorded using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or spraying water on the camera len.\gi{una footnote dicendo sequences can be released upon request?}
@@
 \section{Conclusion}\label{sec:Conclusion}
-\gi{Adriano: butta in inglese gli ongoing works (come ultima cosa)}
-.
+A low-complexity and sequential monitoring scheme over the average gradient norm is used in~\cite{alippi2010detecting} to detect blurring due to external disturbances on the camera lens.
 

### 2015-05-28 paper.tex
segmantation da completare

@@
 where $\mathcal{D}_t$ denotes an operator that transforms the original image $y_t$ in the acquired frame $z_t$; $X \subset \mathbb{Z}^2$ denotes the regular pixels grid and $x\in \mathbb{Z}^2$ indicates the pixel coordinates.
-As far as there is no tampering attacks/events,
+As far as there are no tampering attacks/events,
 \begin{equation}
@@
 \end{equation}
-where $\eta_t$ is a random variable accounting for image noise, %sources (e.g., thermal, quantization, photon-counting). 
-and all the images $y_t$ might show different content but are acquired from the same viewpoint and the same camera orientation.
+where $\eta_t$ is a random variable accounting for image noise and $y_t$ are acquired from the same viewpoint and camera orientation, even though $y_t$ might show different content because of scene changes.
 
-When, at time $\tau^*$, an external disturbance introduces \emph{blur/defocus}, the image $y_t$ is degraded by an unknown blur operator, and $z_t$ becomes
+When, at time $\tau^*$, an external disturbance introduces \emph{blurring}, the image $y_t$ is degraded by an unknown blur operator, and $z_t$ becomes
 \begin{equation}
@@
 Algorithm \ref{alg:DISPL} presents the proposed tampering-detection algorithm, which relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$).
-The former is meant to detect camera displacement, the latter camera defocus.
+The former is meant to detect camera displacement, the latter camera defocus (because blur cases a drop in the high-frequency content).
 As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$.
@@
 %Tampering changes the sequence of indicators:
-Figure \ref{fig:indicatori}(a) shows how a displacement \gi{TODO} affects the values of $g$.
-First of all, we observe that displaced frames yield \gi{WHAT} in the indicator sequence.
-Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here, since these are meant for independent and identically distributed (i.i.d.) random variables.
-In fact, normal changes in the scene or in the illumination yield indicators that follow an unpredictable (but deterministic) trend, beside some high-frequency fluctuations.
+Figure~\ref{fig:indicatori}(a) shows how a camera displacement affects the image intensities.
+First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) and is even more clearly visible in the sequences $l_k(t)$ of two regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)).
+Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here.
+In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables and here, normal changes in the scene or in the illumination, yield indicators that follow an unpredictable (but deterministic) trend.
 Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by a temporal derivative (line~\ref{DISPL-Test5})
@@
 
-The sequence of detrended indicators can be suitably monitored by defining the following confidence interval: 
+The sequence of detrended indicators are reported in Figure~\ref{fig:indicatori}, and can be suitably monitored by defining the following confidence interval:
 \begin{equation}\label{eq:confidenceRegions}
- [\overline{\partial l}_k(t) - \gamma_l \sigma_{l_k}, \overline{\partial l}_k(t) + \gamma_l \sigma_{l_k}],  k = 1,\dots,K
+ [\overline{\partial l}_k(t) - \gamma_l \sigma_{l_k}, \overline{\partial l}_k(t) + \gamma_l \sigma_{l_k}], \ \  k = 1,\dots,K\,,
 \end{equation}
@@
 This threshold was set to allow a single region not to change, as it typically happens in the sky when the camera is displaced, either horizontally or vertically.
-In contrast, any outlier in $\partial g$ would yield a blurring\gi{probabilmente e' piu' corretto parlare di blurring invece che defocus..
-Adriano, riesci ad allineare?
-Io menzionerei anche le occlusions assieme al displacement, che dici?} alert (line~\ref{DISPL-Test11}).
+In contrast, any outlier in $\partial g$ would yield a blurring alert (line~\ref{DISPL-Test11}).
 
-It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, only the first and last tampered frames correspond to an outlier in the detrended indicators, see Figure~\ref{fig:indicatori}(b).
-Therefore, detrended indicators have to be monitored in a one-shot manner, targeting the detection of the first-tampered frame.
-On the one hand, this one-shot solution can increase the detection promptness, and this is particularly important at the low frame-rates we consider.
+It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, only the first and last tampered frames correspond to an outlier in the detrended indicators, see the detrended indicators in Figure~\ref{fig:indicatori}.
+Therefore, detrended indicators have to be monitored in a one-shot manner, targeting the detection of the first tampered-frame.
+On the one hand, this one-shot solution can provide prompt detections, and this is particularly important at the low frame-rates we consider.
 On the other hand, it disregards the fact that tampering might persist over several frames.
-To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, and possibly combined with the one-shot monitoring.
+To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, and possibly combined with the proposed one-shot monitoring.
 
@@
 \subsection{Scene Segmentation}\label{subsec:Segmentation}
-Scene has to be preliminarily segmented to define regions  that our algorithm takes as input.
-To this purpose, we need an external training sequence\gi{Dobbiamo dargli un nome} depicting the scene in its normal conditions without any tampering.
-Segmentation is performed by associating to each pixel $x\in\mathcal{X}$ a feature vector $\textbf{f}(x)\in \mathbb{R}^5$:
+Scene has to be preliminarily segmented to define regions that our algorithm takes as input (see Algorithm \ref{alg:DISPL}).
+To this purpose, we use part of the training frames before $T_0$ to compute a feature vector $\textbf{f}(x)\in \mathbb{R}^5$ for each pixel $x\in\mathcal{X}$
 \begin{equation}
 \label{eq:featureVector}
-\textbf{f}(x)=\left[r(x);c(x);\overline{g}(x);\sigma_{g}(x);\bar{l}(x)\right]\,, \ \forall x \in X
+\textbf{f}(x)=\left[r(x);c(x);\bar{l}(x);\sigma_{l}(x);\overline{g}(x);\sigma_{g}(x)\right]\,, \ \forall x \in X
 \end{equation}
-where $r(x)$ and $c(x)$ represents the row and columns of $x$, $\overline{g}(x)$ and is the gradient norm (smoothed in a neigbhor of $x$) averaged over the training sequence and $\sigma_g$ is the standard deviation of the gradient norm (smoothed in a neigbhor of $x$), computed overtime, and $\bar{l}(x)$ is the average of the gradient over time \gi{ma la deviazione standard della luma nel tempo?
+where $r(x)$ and $c(x)$ represent the row and columns of $x$, respectively,
+
+$\overline{g}(x)$ and is the gradient norm (smoothed in a neigbhor of $x$) averaged over the training sequence and $\sigma_g$ is the standard deviation of the gradient norm (smoothed in a neigbhor of $x$), computed overtime, and $\bar{l}(x)$ is the average of the gradient over time \gi{ma la deviazione standard della luma nel tempo?
 cosa faceva?}
 
-Segmentation consists in clustering the feature vectors computed over the whole image by means a weighted version of the k-means algorithm~\cite{han2006data}.
-In particular, since the components of the feature vector might span very different ranges, the distance between a feature vector and the center of a cluster is measured as the \textit{weighted Minkowski distance}~\cite{de2012minkowski}\gi{..
-ho tolto questo, non e' uguale allo standard k-means il resto?
-from the cluster's centroid is minimum with respect to the distance from the other centroids}, which takes into account the standard deviation of each component for each cluster.
+Segmentation consists in clustering the feature vectors~\eqref{eq:featureVector} over the whole image by means of a weighted version of the k-means algorithm~\cite{kottke1994motion}.
+In particular, since the components of the feature vector span very different ranges, the distance between a feature vector and any cluster's centroid, is weighted by the standard deviation of each component for each cluster.
 
-inside each cluster \cite{kottke1994motion}, in order to consider, during the clustering, also the contribute of the fields which have low variance.
-\ad{Giacomo: prova a vedere se si capisce cosa voglio dire in queste frasi!!!}
+The initial number of clusters is instead chosen by \gi{aggiungere reeference e dire che è una procedura iterativa in cui abbiamo inserito il kmeans peasato.}
 
-The initial number of clusters is instead chosen by \gi{COME? e poi non ricordo se hai trovato il numero di clusters facendo il k-means con i pesi}
-
-Finally, some canonical morphological image-operations are executed to remove boundaries between different regions, and eventually too small regions.
+Finally, morphological image-processing operations are executed to remove boundaries between different regions and eventually regions that are too small.
 This defines the regions $\{R_k\,, \ k=1,\dots,K\}$.

### 2015-05-28 paper.tex
Modifiche alle immagini...

--- 2015-05-28 paper.tex
@@
 In~\cite{gil2007automatic}, tampering detection is performed by combining background subtraction together with edge detection and normalized cross-correlation.
-Background models and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while number of SURF~\cite{SURF} keypoints is used to detect blurring.
+Background models and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while number of SURF~\cite{bay2006surf} keypoints is used to detect blurring.
 A frame buffer rather than an explicit background is used~\cite{ribnick2006real}.
@@
 The first dataset contains \gi{how many?} sequences taken from webcams recording urban areas (as in Figure \ref{fig:sequences}(a)), where we have synthetically introduced blurring by means of a gaussian filters (Figure \ref{fig:sequences}(b)), and displacement by concatenating frames from sequences depicting similar environments (Figure \ref{fig:sequences}(c)).
-In some cases, sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\
+In some cases, sequences contained real tampering events, as in Figure \ref{fig:indicatori}.\\
 
-The second dataset was recorded using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or spraying water on the camera len.\gi{una footnote dicendo sequences can be released upon request?}
+The second dataset was recorded using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or spraying water on the camera lens.\gi{una footnote dicendo sequences can be released upon request?}
 

### 2015-05-28 paper.tex
Inserito criterio di calinski abracadabra

@@
 
-The initial number of clusters is instead chosen by \gi{aggiungere reeference e dire che è una procedura iterativa in cui abbiamo inserito il kmeans peasato.}
+The number of clusters is instead chosen by the Calinski-Harabasz criterion~\cite{calinski1974dendrite}, which considers different k-means instances with different number of clusters, and chooses the best one using a cost function.
+\ad{which is an iterative procedure where we put weighted k-means}.\gi{aggiungere reeference e dire che è una procedura iterativa in cui abbiamo inserito il kmeans peasato.}
 

### 2015-05-28 paper.tex
minor edits

--- 2015-05-28 paper.tex
@@
 In~\cite{gil2007automatic}, tampering detection is performed by combining background subtraction together with edge detection and normalized cross-correlation.
-Background models and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while number of SURF~\cite{bay2006surf} keypoints is used to detect blurring.
+Background models and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while number of SURF~\cite{SURF} keypoints is used to detect blurring.
 A frame buffer rather than an explicit background is used~\cite{ribnick2006real}.
@@
 
-Segmentation consists in clustering the feature vectors~\eqref{eq:featureVector} over the whole image by means of a weighted version of the k-means algorithm~\cite{kottke1994motion}.
-In particular, since the components of the feature vector span very different ranges, the distance between a feature vector and any cluster's centroid, is weighted by the standard deviation of each component for each cluster.
+Segmentation consists in clustering the feature vectors~\eqref{eq:featureVector} over the whole image by a weighted k-means~\cite{kottke1994motion}.
+In particular, since the components of the feature vector span very different ranges, when computing the distance between a feature vector and any cluster's centroid, we scale the distance on each component by its standard deviation.
 
@@
 The first dataset contains \gi{how many?} sequences taken from webcams recording urban areas (as in Figure \ref{fig:sequences}(a)), where we have synthetically introduced blurring by means of a gaussian filters (Figure \ref{fig:sequences}(b)), and displacement by concatenating frames from sequences depicting similar environments (Figure \ref{fig:sequences}(c)).
-In some cases, sequences contained real tampering events, as in Figure \ref{fig:indicatori}.\\
+In some cases, sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\
 
-The second dataset was recorded using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or spraying water on the camera lens.\gi{una footnote dicendo sequences can be released upon request?}
+The second dataset was recorded using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or spraying water on the camera len.\gi{una footnote dicendo sequences can be released upon request?}
 

### 2015-05-28 paper.tex
Computational Complexity

@@
 Figure~\ref{fig:indicatori}(a) shows how a camera displacement affects the image intensities.
-First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) and is even more clearly visible in the sequences $l_k(t)$ of two regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)).
+First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) which is more clearly visible in the sequences $l_k(t)$ of two displayed regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)).
 Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here.
-In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables and here, normal changes in the scene or in the illumination, yield indicators that follow an unpredictable (but deterministic) trend.
+In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables, while here normal changes in the scene or in the illumination, yield indicators that follow an unpredictable (but deterministic) trend.
 Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by a temporal derivative (line~\ref{DISPL-Test5})
@@
 \label{eq:featureVector}
-\textbf{f}(x)=\left[r(x);c(x);\bar{l}(x);\sigma_{l}(x);\overline{g}(x);\sigma_{g}(x)\right]\,, \ \forall x \in X
+\textbf{f}(x)=\left[r(x);c(x);\bar{l}(x);\sigma_{l}(x);\overline{g}(x);\sigma_{g}(x)\right]\,, \ \forall x \in X\,.
 \end{equation}
-where $r(x)$ and $c(x)$ represent the row and columns of $x$, respectively,
-
-$\overline{g}(x)$ and is the gradient norm (smoothed in a neigbhor of $x$) averaged over the training sequence and $\sigma_g$ is the standard deviation of the gradient norm (smoothed in a neigbhor of $x$), computed overtime, and $\bar{l}(x)$ is the average of the gradient over time \gi{ma la deviazione standard della luma nel tempo?
-cosa faceva?}
+In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and columns of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the average and standard deviation of the intensity at $x$ computed over time and $\overline{g}(x)$ and $\sigma_g(x)$ are the average norm of the gradient and its standard deviation, computed overtime.
+These feature vectors are meant to cluster image regions according to their spatial and temporal behavior during the training frames.
 
 Segmentation consists in clustering the feature vectors~\eqref{eq:featureVector} over the whole image by a weighted k-means~\cite{kottke1994motion}.
-In particular, since the components of the feature vector span very different ranges, when computing the distance between a feature vector and any cluster's centroid, we scale the distance on each component by its standard deviation.
+In particular, since the components of the feature vector span very different ranges, we compute the distance between a feature vector and a cluster centroid by scaling the Euclidean distance of a weight inversely proportional to the standard deviation over the cluster.
+\gi{Adriano, controlla bene}
+The number of clusters is defined by the Calinski-Harabasz criterion~\cite{calinski1974dendrite}, which initializes the weighted k-means using different number of clusters, and then chooses the best one according a cost function\gi{Adriano controllare come lo chiama}.
 
-The number of clusters is instead chosen by the Calinski-Harabasz criterion~\cite{calinski1974dendrite}, which considers different k-means instances with different number of clusters, and chooses the best one using a cost function.
-\ad{which is an iterative procedure where we put weighted k-means}.\gi{aggiungere reeference e dire che è una procedura iterativa in cui abbiamo inserito il kmeans peasato.}
-
-Finally, morphological image-processing operations are executed to remove boundaries between different regions and eventually regions that are too small.
+Finally, morphological image-processing operations are executed to remove boundaries between different regions, and eventually regions that are too small.
 This defines the regions $\{R_k\,, \ k=1,\dots,K\}$.
@@
 
-The algorithm uses low computational techniques: the biggest effort is made by computation of the feature $g(t)$, which needs $34$ operations per pixel, due to the Sobel filter used during the convolution.
-The computational load could be decreased using Fast Fourier Transform (FFT) \cite{brigham1988fast} and operating in the frequency domain.
+The most computationally demanding operations in the algorithm consists in computing of the indicators $g(t)$.
+Assuming that derivatives in~\eqref{eq:normaGradiente} are computed by the Sobel filters, this requires $34$ operations per pixels (and execution times however can be substantially decreased whether hardware implementations of the FFT transform are available).
+Computing $l(t)$ requires 1 operation per pixel and outlier detection requires two comparisons per frame.
+Using separate regions has a negligible impact over the overall computational complexity, as well as on the monitoring activity.
+Segmentation has not been considered because it is executed only for training the algorithm, and can be performed on an external device, connected during configuration.
 
-Segmentation has not been considered because...
-
-FRAME DIFFERENCE:
-\begin{equation}
-	\label{eq:frameDiffReg}
-	\varphi^k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
-\end{equation}
+%FRAME DIFFERENCE:
+%\begin{equation}
+%	\label{eq:frameDiffReg}
+%	\varphi^k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
+%\end{equation}
 
@@
 \section{Experiments}\label{sec:experiments}
-\gi{Per ultimo: The proposed algorithm has been implemented in MATLAB, }
 
@@
 
+\gi{Per ultimo: The proposed algorithm has been implemented in MATLAB, }
+
 \section{Conclusion}\label{sec:Conclusion}

### 2015-05-28 paper.tex
minor edits

@@
 We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme.
-Any frame yielding anomalous indicator is detected as a tampering attempt.
+Any frame yielding anomalous indicator is detected as a tampered frame.
 Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements.
@@
 % What is the problem?
-When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as in Fig.
-\ref{fig:pioggia}, or in partial occlusions of the scene, as in Fig.
-\ref{fig:neve}.
+When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as shown in Figure \ref{fig:pioggia}, or in partial occlusions of the scene, as in Fig \ref{fig:neve}.
 Similarly, intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be useless for monitoring purposes.
@@
 Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly report alerts.
-In fact, tampering prevents the correct scene interpretation: even a mild blur might hinder the identification of important details such as licence plates.
+In fact, tampering prevents the correct interpretation of the scenex: even a mild blur might hinder the identification of important details such as licence plates.
 Surveillance cameras are typically connected to the power supply and operates at normal frame-rates (e.g.
@@
 
-In particular, camera displacement and occlusions are typically detected by background models or frame comparison, while blurring is instead detected by monitoring the high-frequency component of images, as this is expected to drop.
-In ~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain.
-Two background models computed over different time intervals are used in~\cite{saglam2009real}.
+In particular, camera displacement and occlusions are typically detected by estimating the background or computing frame difference, while blurring is instead detected by monitoring the high-frequency components of images, as these are expected to drop.
+In~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain.
+Two background models, estimated over different time intervals, are used in~\cite{saglam2009real}.
 In~\cite{gil2007automatic}, tampering detection is performed by combining background subtraction together with edge detection and normalized cross-correlation.
-Background models and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while number of SURF~\cite{bay2006surf} keypoints is used to detect blurring.
+Background estimates and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while the number of SURF~\cite{bay2006surf} keypoints is used to detect blurring.
 A frame buffer rather than an explicit background is used~\cite{ribnick2006real}.
-All these solutions, including~\cite{ribnick2006real,harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.
+All these solutions, including~\cite{harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.
 
@@
 % Why is it hard? (E.g., why do naive approaches fail?)
-Tampering detection in low-power smart cameras has not been much investigated in the literature~\cite{alippi2010detecting}, while being more challenging than in conventional surveillance cameras~\cite{perrig2004security}.
+Tampering detection in low-power smart cameras has not been much investigated in the literature~\cite{alippi2010detecting}, also because it is more challenging than in conventional surveillance cameras~\cite{perrig2004security}.
 Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), thus the acquired sequence does not evolve smoothly.
-This prevents the use of learned background models and the analysis of foreground variations \cite{piccardi2004background}.
+This prevents the use of learned background models and the analysis of foreground variations.
 In fact, when dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions, yielding sequences like those in Figure~\ref{fig:sequences}.
 Smart cameras have to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames.
-Moreover, false alarms represent a serious problem in low-power smart cameras as these result in useless data processing and network transmissions, thus energy waste.
+Moreover, false alarms represent a serious problem in low-power smart cameras since they result in energy waste because of useless data processing and network transmissions.
 % We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose.
@@
 Remarkably, these regions can be preliminarily defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t.
-monitoring the whole image.
-Thus, our solution represents as a prompt trigger that can be possibly combined with other sequential monitoring techniques.
+monitoring the whole image (Section \ref{subsec:coputationalComplexity}).
+Thus, our solution represents a prompt trigger that can be possibly combined with other sequential monitoring techniques.
 Experiments in Section \ref{sec:experiments} show that leveraging image regions can substantially improve the displacement-detection performance.
@@
 
-The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect time $\tau^*$ when any tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.
+The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect the time instant $\tau^*$ when tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.
 We assume that $T_0$ tampering-free frames and provided for training.
@@
 Algorithm \ref{alg:DISPL} presents the proposed tampering-detection algorithm, which relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$).
-The former is meant to detect camera displacement, the latter camera defocus (because blur cases a drop in the high-frequency content).
+The former is meant to detect camera displacement, the latter blurring over camera lens (because blur cases a drop in the high-frequency content).
 As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$.
@@
 
-\subsubsection{Computational Complexity}
+\subsection{Computational Complexity}\label{subsec:coputationalComplexity}
 
@@
 As we could see there is an improvement, in the detection of camera displacements, with our approach which separately analizes the behavior of indicators in adaptive regions.
-On the other side, as illustrated in Figure \ref{fig:ROCdefocus}, when monitoring indicators meant to detect blur/defocus, it is more convenient to consider the whole scene at once.
+On the other side, as illustrated in Figure \ref{fig:ROCdefocus}, when monitoring indicators meant to detect blurring, it is more convenient to consider the whole scene at once.
 
@@
 \cline{2-7}
-& \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Displacement Detection}}  & \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Defocus Detection}} \\ \cline{2-7} 
+& \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Displacement Detection}}  & \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Blurring Detection}} \\ \cline{2-7} 
 & \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} & \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} \\ \hline

### 2015-05-28 paper.tex
stesura completa

@@
 \begin{equation}\label{eq:lumaRegions}
-l_k(t) =\frac{1}{\#R_k} \sum_{R_k} z_t(x), \ \ k = 1, \dots, K\,,
+l_k(t) =\frac{1}{\#R_k} \sum_{x \in R_k} z_t(x), \ \ k = 1, \dots, K\,,
 \end{equation}
@@
 \label{eq:normaGradiente}
-g(t) = \sum_X \left (\sqrt{\left(z_t \circledast f_h\right)^2(x) + \left(z_t \circledast f_v\right)^2(x)}\right),
+g(t) = \sum_{x \in X} \left (\sqrt{\left(z_t \circledast f_h\right)^2(x) + \left(z_t \circledast f_v\right)^2(x)}\right),
 \end{equation}
@@
 Figure~\ref{fig:indicatori}(a) shows how a camera displacement affects the image intensities.
-First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) which is more clearly visible in the sequences $l_k(t)$ of two displayed regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)).
+First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) which is more clearly visible in the sequences $l_k(t)$ corresponding to the two displayed regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)).
 Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here.
-In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables, while here normal changes in the scene or in the illumination, yield indicators that follow an unpredictable (but deterministic) trend.
-Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by a temporal derivative (line~\ref{DISPL-Test5})
+In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables, while here changes in the scene or in the illumination in non-tampered frames, yield indicators that follow an unpredictable trend.
+Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by temporal derivative (line~\ref{DISPL-Test5})
 \begin{equation}\label{eq:detrending}
@@
 
-The sequence of detrended indicators are reported in Figure~\ref{fig:indicatori}, and can be suitably monitored by defining the following confidence interval:
+The sequences of detrended indicators are reported in Figure~\ref{fig:indicatori}, and can be suitably monitored by defining the following confidence intervals:
 \begin{equation}\label{eq:confidenceRegions}
@@
 $z_t\,, t = 1, \dots, T_o$) and $\gamma_l >0$ is a tuning parameter.
-A similar interval is built for $\partial g$ (lines \ref{DISPL-Tr7}).
+A similar interval is built for $\partial g$ (line~\ref{DISPL-Tr7}).
 
@@
 On the one hand, this one-shot solution can provide prompt detections, and this is particularly important at the low frame-rates we consider.
-On the other hand, it disregards the fact that tampering might persist over several frames.
+On the other hand, one-shot monitoring disregards the fact that tampering might persist over several frames.
 To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, and possibly combined with the proposed one-shot monitoring.
@@
 	\SetNlSty{small}{}{.}
-	\textbf{Input}: $\gamma_l, \gamma_g$, $\{z_t, t = 1, \dots, T_{0}\}$, $\{R_k \ k=1,\dots,K\}$ \\
+	\textbf{Input}: $\gamma_l, \gamma_g$, $\{z_t, t = 1, \dots, T_{0}\}$, $\{R_k, \ k=1,\dots,K\}$ \\
 	\textbf{Configuration}:\\
 	\lnl{DISPL-Tr1} Compute $\partial l_k(t)$ and $\partial g(t), t = 1, \cdots, T_o$\\ 
-	\lnl{DISPL-Tr7} Compute $\partial l_k$, $\partial g$, $\sigma_{l_k}$ and $\sigma_{g}$.\\
+	\lnl{DISPL-Tr7} Compute $\overline{\partial l}_k$, $\overline{\partial g}$, $\sigma_{l_k}$ and $\sigma_{g}$.\\
 
@@
 \end{equation}
-In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and columns of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the average and standard deviation of the intensity at $x$ computed over time and $\overline{g}(x)$ and $\sigma_g(x)$ are the average norm of the gradient and its standard deviation, computed overtime.
-These feature vectors are meant to cluster image regions according to their spatial and temporal behavior during the training frames.
+In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and column of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the average and standard deviation of the intensity at $x$ computed over time, and $\overline{g}(x)$ and $\sigma_g(x)$ are the average norm of the gradient and its standard deviation, computed overtime.
+These feature vectors are meant to cluster image regions according to their spatial appearance and temporal behavior over training frames.
 
@@
 Finally, morphological image-processing operations are executed to remove boundaries between different regions, and eventually regions that are too small.
-This defines the regions $\{R_k\,, \ k=1,\dots,K\}$.
+This defines the regions $\{R_k\,, \ k=1,\dots,K\}$, and two examples are reported in Figure~\ref{fig:indicatori}.
 
@@
 The most computationally demanding operations in the algorithm consists in computing of the indicators $g(t)$.
-Assuming that derivatives in~\eqref{eq:normaGradiente} are computed by the Sobel filters, this requires $34$ operations per pixels (and execution times however can be substantially decreased whether hardware implementations of the FFT transform are available).
-Computing $l(t)$ requires 1 operation per pixel and outlier detection requires two comparisons per frame.
-Using separate regions has a negligible impact over the overall computational complexity, as well as on the monitoring activity.
-Segmentation has not been considered because it is executed only for training the algorithm, and can be performed on an external device, connected during configuration.
+Assuming that derivatives in~\eqref{eq:normaGradiente} are computed by the Sobel filters, this requires $34$ operations per pixels\footnote{Execution times can be however  decreased whether hardware implementations of the FFT transform are available}.
+Computing $l(t)$ requires 1 operation per pixel and the outlier detection over the indicators require two comparisons per frame.
+Thus, monitoring regions has a negligible impact over the overall computational complexity.
+Segmentation has not been here considered because it is executed only when training the algorithm, and can be performed on an external device connected to the smart camera during configuration.
 
@@
 
-Experiments are meant to show the effectiveness of the proposed algorithm on real-world sequences, and in particular on assessing the advantages of using segmentation for detecting camera displacements.
+Experiments are meant to show the effectiveness of the proposed algorithm on real-world sequences, and to specifically assess the advantage of using segmentation for detecting camera displacements. 
+\subsubsection{Considered Methods}
+As far as camera-displacement is concerned, Algorithm \eqref{alg:DISPL} is compared against three variants: 1) monitoring the luma on the whole image (\emph{Luma-Total}), 2) monitoring the total energy of the difference between two adjacent frames (\emph{Frame Difference}), and 3) monitoring the luma on $K$ regions obtained by clustering $[r(x), c(x)]$, which yields a Voronoi partitioning of the scene (\emph{Luma-Voronoi}).
+Frame difference is considered as a baseline technique for displacement detection, while comparison against \emph{Luma-Voronoi} shows the advantages of setting regions depending of the image content.
+In the experiments on blurring detection we compare Algorithm \ref{alg:DISPL} against monitoring the gradient separately over image regions, as done for $\partial l_k$.
 
-We have acquired two datasets to test our method.
-The first dataset contains \gi{how many?} sequences taken from webcams recording urban areas (as in Figure \ref{fig:sequences}(a)), where we have synthetically introduced blurring by means of a gaussian filters (Figure \ref{fig:sequences}(b)), and displacement by concatenating frames from sequences depicting similar environments (Figure \ref{fig:sequences}(c)).
-In some cases, sequences contained real tampering events, as in Figure \ref{fig:indicatori}.\\
+\subsubsection{Datasets}
+We acquired two datasets: the first dataset contains \gi{Adriano: how many?} sequences of \gi{Adriano: how many?} frames each that have been recorded from webcam monitoring urban areas (e.g., see Figure \ref{fig:sequences}(a)).
+The second dataset was recorded using a Raspberry Pi Model B+ \gi{Adriano: refrence} mounting a camera module.
+In the first dataset tampering was synthetically generated by blurring frames (see, Figure \ref{fig:sequences}(b)) or concatenating frames from different sequences depicting similar environments (see Figure \ref{fig:sequences}(c)).
+Few sequences of the first dataset contained real tampering events \gi{Adriano: quanti?
+solo displacement?
+niente blur considerato qui?} (Figure \ref{fig:indicatori}).
+In the second dataset, tampering was introduced by moving the device or spraying water on the camera lens.
+%\gi{una footnote dicendo sequences can be released upon request?}
 
-The second dataset was recorded using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or spraying water on the camera lens.\gi{una footnote dicendo sequences can be released upon request?}
+\subsubsection{Performance Assessment}
+The typical figure of merit for assessing detection performance in one-shot monitoring methods are the ROC curves: in our case, these can be computed by varying $\gamma_l$ and $\gamma_g$ in a range of values.
+In particular, each point in the ROC curve depicts a pair of values (1-SPECIFICITY, RECALL) defined as
+\[1-\text{SPECIFICITY}_\gamma = \frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}, \ \text{ and } \ \text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}\]
+where TP, FP, TN, FN represent the number of true positives (tampered frames correctly detected), of false positives (tampering-free frames detected), of true negatives (tampering-free frames not detected) and of false negatives (number of missed tampering frames), for a given value of threshold $\gamma_l$ (or $\gamma_g$).
 
-Typical figures of merit for one-shot monitoring methods have been used to assess the performance of the proposed algorithm.
-In particular we compute the ROC curves varying $\gamma_l$ and $\gamma_g$ in a range of values.
-ROC curve reports :
-\[1-\text{SPECIFICITY}_\gamma = 1-\frac{\text{TN}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}=\frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma},\]
-while on th y-axis there is:
-\[\text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
-The construction of these curves permits us to make a comparison with respect to other methods.\gi{io le metterei solo in una equazione entrambe, e per 1-specificity solo la terza formula}
+\subsubsection{Discussion}
+ROC curves and the AUC in Figure~\ref{fig:ROCdisplacement} caption indicate that camera displacements can be effectively detected.
+However, the best performance are achieved by Algorithm \ref{alg:DISPL} the (\emph{Frame-Difference} curve has not been reported because it was below these).
+This result indicate that separately monitoring the luma over image regions is an effective strategy and that the regions have to be carefully designed.
+Please note that, for the visualization sake, these ROC curves have been reported in a limited range of x and y axis.
+While the improvement might seem negligible we observe that Algorithm~\ref{alg:DISPL} achieves \gi{Adriano, quanto?} false positive rates (i.e., 1-SPECIFICITY) than..
 
-\begin{description}
-	\item[TP]  True positives. It measures the number of tampering events correctly detected by the algorithm.\\ 
-	\item[TN]  True negatives. It measures the number of frames without tampering that are not detected by the algorithm. \\ 
-	\item[FP]  False positives. It measures the number of tampering events erroneously detected by the algorithm.\\ 
-	\item[FN]  False negatives. It measures the number of tampering events not detected by the algorithm. \\ 
-\end{description} 
-\gi{questo lo metterei solo nel testo sotto le formule sopra}
+In contrast, Figure~\ref{fig:ROCdefocus} confirms that when monitoring $g$ it is more convenient to consider the whole scene at once.
+This is probably due to the fact that typically, regions $\{R_k\}$ do not include strong edges, which are the most informative to detect blurring.
 
-\gi{io inizierei dai metodi con cui confrontarci}
-In particular the alternative approaches that we have considered working:
-\begin{itemize}
-\item considering the whole scene for the features computation;
-\item considering adaptive region, as described in Section \ref{sec:propSol}, for the feature computation;
-\item considering voronoi regions \cite{aurenhammer1991voronoi}, which are easier to compute with respect to our solution but don't consider the scene content, for the feature computation.
-\end{itemize}
-
-FRAME DIFFERENCE:
-\begin{equation}
-	\label{eq:frameDiffReg}
-	\varphi^k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
-\end{equation}
-
-%
-%% \subsection{Alternative Approaches}\label{subsec:AlternativeApproaches}
-%
-%
-%\subsection{Performance Assessment}
-%\gi{Adriano: Dire come vengono calcolate le ROC curves TPR e FPR, le cifre di merito insomma, spiegando bene che parametro varia}
-%
-%\gi{Adriano: metti entrambe le ROC curves, affiancate e per bene ed alcuni esempi di sequenze}
 \begin{figure}[t]
@@
 %\end{figure}
-Experimental results are shown in Figures \ref{fig:ROCdisplacement} and \ref{fig:ROCdefocus}.
-As we could see there is an improvement, in the detection of camera displacements, with our approach which separately analizes the behavior of indicators in adaptive regions.
-On the other side, as illustrated in Figure \ref{fig:ROCdefocus}, when monitoring indicators meant to detect blurring, it is more convenient to consider the whole scene at once.
 
-\begin{table}[tbh]
-\centering
-\begin{tabular}{l|l|l|l|l|l|l|}
-\cline{2-7}
-& \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Displacement Detection}}  & \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Blurring Detection}} \\ \cline{2-7} 
-& \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} & \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} \\ \hline
-\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Luma}}     & 0.9994 & 0.9989 & 0.9985  &            &            &             \\ \hline
-\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Gradient}} &  		 &  		  &             & 0.9895 & 0.9996 & 0.9996  \\ \hline
-\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{FD}}         & 0.9974 & 0.9944 & 0.9974  & 0.5526 & 0.5322 & 0.5391  \\ \hline
-\end{tabular}
-\caption{Area under curve of the analized solutions}
-\label{tab:AUC}
-\end{table}
+%
+%\begin{table}[tbh]
+%\centering
+%\begin{tabular}{l|l|l|l|l|l|l|}
+%\cline{2-7}
+%& \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Displacement Detection}}  & \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Blurring Detection}} \\ \cline{2-7} 
+%& \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} & \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} \\ \hline
+%\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Luma}}     & 0.9994 & 0.9989 & 0.9985  &            &            &             \\ \hline
+%\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Gradient}} &  		 &  		  &             & 0.9895 & 0.9996 & 0.9996  \\ \hline
+%\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{FD}}         & 0.9974 & 0.9944 & 0.9974  & 0.5526 & 0.5322 & 0.5391  \\ \hline
+%\end{tabular}
+%\caption{Area under curve of the analized solutions}
+%\label{tab:AUC}
+%\end{table}
 
@@
 
-\gi{Per ultimo: The proposed algorithm has been implemented in MATLAB, }
+\section{Conclusion}\label{sec:Conclusion}
+We presented a tampering detection algorithm that leverages a preliminary image partition to improve the detection of camera displacements, while at the same time can detect blurring.
+This algorithm has low-computational complexity and is suited for low-power cameras operating at low frame rates.
 
-\section{Conclusion}\label{sec:Conclusion}
-A low-complexity and sequential monitoring scheme over the average gradient norm is used in~\cite{alippi2010detecting} to detect blurring due to external disturbances on the camera lens.
+Ongoing work concern the extension of our solution to other types of tampering, as occlusions or imaging sensor degradations, as well as the integration of this one-shot technique with sequential monitoring scheme such as~\cite{alippi2010detecting}, to leverage the fact that tampering might last more than one frame in the detection.
+Moreover, we will investigate different segmentation strategy and test our algorithm on a larger dataset.
 
-Random Toughs:
-\begin{itemize}
-\item The problem of false alarms, radio module activation
-\item Other tampering attacks like obfuscation (??) which might be due to environmental phenomena such as rain, fog and mist over the camera lenses have to be detected by image analysis methods
-\item Displacement can be perceived by MEMS as well but these device alone are prone to false alarms. Visual inspection is necessary to reduce false alarms
-\end{itemize}
-Ongoing work regards the extension of our solution to other types of tampering, as occlusions or imaging sensor degradations.
-Furthermore, we are investigating strategies to improve the detection performance and reduce the number of \textit{FP}s by combination of our solution with other techniques:
-for example we could use sequential techniques on the features, or integrate the frame analysis with the data extracted from MEMS inertial sensors.
+%Ongoing work regards the.
+%Furthermore, we are investigating strategies to improve the detection performance and reduce the number of \textit{FP}s by combination of our solution with other techniques:
+%for example we could use sequential techniques on the features, or integrate the frame analysis with the data extracted from MEMS inertial sensors.
+
+%Random Toughs:
+%\begin{itemize}
+%\item Better segmentation
+%\item tests on larger datasets
+%\item The problem of false alarms, radio module activation
+%\item Other tampering attacks like obfuscation (??) which might be due to environmental phenomena such as rain, fog and mist over the camera lenses have to be detected by image analysis methods
+%\item Displacement can be perceived by MEMS as well but these device alone are prone to false alarms. Visual inspection is necessary to reduce false alarms
+%\end{itemize}
 

### 2015-05-29 paper.tex
Aggiunta caption con AUC nelle ROC curves

@@
 \subfigure[]{\label{fig:ROCdefocus} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus}}
-\caption{ROC curves for displacement detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean luma energy. \textbf{(b)} Analysis of the frame differencing.}
+\caption{\textbf{(a)} ROC curve for displacement detection.
+	There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
+	In particular, AUC of our solution is $0.9994$, while the solution which considers the whole image has an AUC of $0.9989$ and the one which considers Voronoi regions ha san AUC of $0.9985$.
+	\textbf{(b)} ROC curve for blurring detection.
+	In this case considering adaptive regions instead of considering the whole image is not convenient.
+	In particular, AUC of the solution which considers adaptive regions is $0.9895$,  while the solution which considers the whole image has an AUC of $0.9996$}
 \label{fig:ROC}

### 2015-05-29 paper.tex
Messe caption e diminuito size del pdf

@@
 \centering
-\subfigure[]{\label{fig:pioggia} \includegraphics[width=0.4\linewidth]{Immagini/pioggia}}
-\subfigure[]{\label{fig:neve} \includegraphics[width=0.4\linewidth]{Immagini/neve}}
+\subfigure[]{\label{fig:pioggia} \includegraphics[width=0.3\linewidth]{Immagini/pioggia}}
+\subfigure[]{\label{fig:neve} \includegraphics[width=0.3\linewidth]{Immagini/neve}}
 \caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena. \textbf{(a)} Blur due to rain drops on the camera lens. \textbf{(b)} Occlusion due to snow on the camera lens.}
@@
 \centering
-\includegraphics[width=1\linewidth]{Immagini/sequenze}
-\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences; \textbf{(d)} real displacement event on 4-th frame.}
+\includegraphics[width=0.8\linewidth]{Immagini/sequenze}
+\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences.}
 \label{fig:sequences}
@@
 \includegraphics[width=1\linewidth]{Immagini/indicatori}
-\caption{Example of displacement and behavior of the average value of the luma.
+\caption{Example of displacement. On the left side of the image is reported the behavior of the average value of the luma, while on the right side there is its detrending.
 The event comes at frame $157$ and the corresponding value of $l(t)$ is depicted with a red star. 
@@
 \section{Experiments}\label{sec:experiments}
-
 Experiments are meant to show the effectiveness of the proposed algorithm on real-world sequences, and to specifically assess the advantage of using segmentation for detecting camera displacements. 
@@
 In the experiments on blurring detection we compare Algorithm \ref{alg:DISPL} against monitoring the gradient separately over image regions, as done for $\partial l_k$.
-
 \subsubsection{Datasets}
@@
 %\gi{una footnote dicendo sequences can be released upon request?}
-
 \subsubsection{Performance Assessment}
@@
 where TP, FP, TN, FN represent the number of true positives (tampered frames correctly detected), of false positives (tampering-free frames detected), of true negatives (tampering-free frames not detected) and of false negatives (number of missed tampering frames), for a given value of threshold $\gamma_l$ (or $\gamma_g$).
-
 \subsubsection{Discussion}

### 2015-05-29 paper.tex
Finalizzato le ultime cose (forse!) Mancano references Claudio

@@
 \end{equation}
-where $\eta_t$ is a random variable accounting for image noise and $y_t$ are acquired from the same viewpoint and camera orientation, even though $y_t$ might show different content because of scene changes.
-
+where $\eta_t$ is a random variable accounting for image noise and $y_t$ are acquired from the same viewpoint and camera orientation, even though $y_t$ might show different content because of scene changes.\\
 When, at time $\tau^*$, an external disturbance introduces \emph{blurring}, the image $y_t$ is degraded by an unknown blur operator, and $z_t$ becomes
@@
 \end{equation}
-where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.
-
+where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.\\
 A \emph{camera displacement} at frame $\tau^*$ is instead modeled as 
@@
 \end{equation}
-where $w_t$ relates to a different viewpoint and/or camera orientation than $y_t$.
-
+where $w_t$ relates to a different viewpoint and/or camera orientation than $y_t$. \\
 The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect the time instant $\tau^*$ when tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.
@@
 The employed segmentation is detailed in Section \ref{subsec:Segmentation}.
-
+\\
 For each $z_t$, the luma is separately computed within each of the $K$ regions:
@@
 \end{equation}
-where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, $\circledast$ denotes the 2d convolution and $\#(\cdot)$ the cardinality of a set.
-
+where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, $\circledast$ denotes the 2d convolution and $\#(\cdot)$ the cardinality of a set.\\
 \begin{figure}[tb]
@@
 \end{figure}
-
 %Tampering changes the sequence of indicators:
@@
 \end{equation}
-and we similarly define $\partial g(t)$.
-
+and we similarly define $\partial g(t)$. \\
 The sequences of detrended indicators are reported in Figure~\ref{fig:indicatori}, and can be suitably monitored by defining the following confidence intervals:
@@
 A similar interval is built for $\partial g$ (line~\ref{DISPL-Tr7}).
-
+\\
 During operations, any indicator falling outside its confidence region is considered anomalous (line~\ref{DISPL-Test6}).
@@
 This threshold was set to allow a single region not to change, as it typically happens in the sky when the camera is displaced, either horizontally or vertically.
-In contrast, any outlier in $\partial g$ would yield a blurring alert (line~\ref{DISPL-Test11}).
-
+In contrast, any outlier in $\partial g$ would yield a blurring alert (line~\ref{DISPL-Test11}).\\
 It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, only the first and last tampered frames correspond to an outlier in the detrended indicators, see the detrended indicators in Figure~\ref{fig:indicatori}.
@@
 In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and column of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the average and standard deviation of the intensity at $x$ computed over time, and $\overline{g}(x)$ and $\sigma_g(x)$ are the average norm of the gradient and its standard deviation, computed overtime.
-These feature vectors are meant to cluster image regions according to their spatial appearance and temporal behavior over training frames.
-
+These feature vectors are meant to cluster image regions according to their spatial appearance and temporal behavior over training frames.\\
 Segmentation consists in clustering the feature vectors~\eqref{eq:featureVector} over the whole image by a weighted k-means~\cite{kottke1994motion}.
 In particular, since the components of the feature vector span very different ranges, we compute the distance between a feature vector and a cluster centroid by scaling the Euclidean distance of a weight inversely proportional to the standard deviation over the cluster.
-\gi{Adriano, controlla bene}
-The number of clusters is defined by the Calinski-Harabasz criterion~\cite{calinski1974dendrite}, which initializes the weighted k-means using different number of clusters, and then chooses the best one according a cost function\gi{Adriano controllare come lo chiama}.
-
+%\gi{Adriano, controlla bene}
+%The number of clusters is defined by the Calinski-Harabasz criterion~\cite{calinski1974dendrite}, which initializes the weighted k-means using different number of clusters, and then chooses the best one according a cost function\gi{Adriano controllare come lo chiama}.
+The number of clusters is defined by initializing the weighted k-means using different number of clusters and then choosing the best solution according to the Calinski-Harabasz criterion~\cite{calinski1974dendrite}.\\
 Finally, morphological image-processing operations are executed to remove boundaries between different regions, and eventually regions that are too small.
@@
 \subsubsection{Datasets}
-We acquired two datasets: the first dataset contains \gi{Adriano: how many?} sequences of \gi{Adriano: how many?} frames each that have been recorded from webcam monitoring urban areas (e.g., see Figure \ref{fig:sequences}(a)).
-The second dataset was recorded using a Raspberry Pi Model B+ \gi{Adriano: refrence} mounting a camera module.
+We acquired two datasets: the first dataset contains  sequences of  frames taken from webcams monitoring urban areas (e.g., see Figure \ref{fig:sequences}(a)). 
+The second dataset was recorded using a Raspberry Pi Model B+ \cite{raspberry} mounting a camera module. 
 In the first dataset tampering was synthetically generated by blurring frames (see, Figure \ref{fig:sequences}(b)) or concatenating frames from different sequences depicting similar environments (see Figure \ref{fig:sequences}(c)).
-Few sequences of the first dataset contained real tampering events \gi{Adriano: quanti?
-solo displacement?
-niente blur considerato qui?} (Figure \ref{fig:indicatori}).
+Few sequences of the first dataset contain real tampering events  (Figure \ref{fig:indicatori}). 
 In the second dataset, tampering was introduced by moving the device or spraying water on the camera lens. 
-%\gi{una footnote dicendo sequences can be released upon request?}
+At last we have considered $9$ different scenarios, with $10$ displacement events and $10$ blurring events each, for a total of $180$ tampering events.%\gi{una footnote dicendo sequences can be released upon request?}
 \subsubsection{Performance Assessment}
@@
 Please note that, for the visualization sake, these ROC curves have been reported in a limited range of x and y axis.
-While the improvement might seem negligible we observe that Algorithm~\ref{alg:DISPL} achieves \gi{Adriano, quanto?} false positive rates (i.e., 1-SPECIFICITY) than..
-
+While the improvement might seem negligible we observe that Algorithm~\ref{alg:DISPL}, for having a recall of $98\%$, achieves $0.3\%$ false positive rates (i.e., 1-SPECIFICITY) than $0.7\%$ of the whole image solution or $1\%$ of the Voronoi regions solution.
+Having a solution which halves the false positives is important in a low-power context, considering the high number of frames to be processed.\\
 In contrast, Figure~\ref{fig:ROCdefocus} confirms that when monitoring $g$ it is more convenient to consider the whole scene at once.
@@
 	In this case considering adaptive regions instead of considering the whole image is not convenient.
-	In particular, AUC of the solution which considers adaptive regions is $0.9895$,  while the solution which considers the whole image has an AUC of $0.9996$}
+	In particular, AUC of the solution which considers adaptive regions is $0.9895$,  while the solution which considers the whole image has an AUC of $0.9996$.}
 \label{fig:ROC}
@@
 
-\section*{Acknowledgments}\label{sec:Acknowledgments}
-Authors would like to thank ST for supporting Adriano Gaibotti.
+%\section*{Acknowledgments}\label{sec:Acknowledgments}
+%Authors would like to thank ST for supporting Adriano Gaibotti.
 
@@
 %
-\subsection{Indicators}\label{subsec:Indicators}
+%\subsection{Indicators}\label{subsec:Indicators}
 

### 2015-05-29 paper.tex
Messa l'affiliazione al Politecnico
first submission

@@
 
-\author{Adriano Gaibotti\inst{1} \and Claudio Marchisio\inst{1} \and Alexandro Sentinelli\inst{1} \and Giacomo Boracchi\inst{2}}
+\author{Adriano Gaibotti\inst{1,}\inst{2} \and Claudio Marchisio\inst{1} \and Alexandro Sentinelli\inst{1} \and \\ Giacomo Boracchi\inst{2}}
 
@@
 No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attacks (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures.
-Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at video-like frame-rates, nor use sophisticated image-analysis algorithms.
+Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at full-speed  frame-rates, nor use sophisticated image-analysis algorithms.
 
@@
 Any frame yielding anomalous indicator is detected as a tampered frame.
-Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements.
+Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to improve the detection of camera displacements.
 Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity.
@@
 % What is the problem?
-When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as shown in Figure \ref{fig:pioggia}, or in partial occlusions of the scene, as in Fig \ref{fig:neve}.
+When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as shown in Figure \ref{fig:pioggia}, or in partial occlusions of the scene, as in Figure \ref{fig:neve}.
 Similarly, intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be useless for monitoring purposes.
@@
 Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly report alerts.
-In fact, tampering prevents the correct interpretation of the scenex: even a mild blur might hinder the identification of important details such as licence plates.
-Surveillance cameras are typically connected to the power supply and operates at normal frame-rates (e.g.
+In fact, tampering prevents the correct interpretation of the scene: even a mild blur might hinder the identification of important details such as licence plates.
+Surveillance cameras are typically connected to the power supply and operate at normal frame-rates (e.g.
 around few frames per second): in these conditions, several tampering-detection algorithms have been presented.
-
 In particular, camera displacement and occlusions are typically detected by estimating the background or computing frame difference, while blurring is instead detected by monitoring the high-frequency components of images, as these are expected to drop.
@@
 
-In this work we expressly target low-power and ultra-low-power smart cameras\gi{Claudio, riesci a mettere una reference qua?}, battery-powered devices that are characterized by a constrained computational power and memory availability.
-Even though these devices are not meant for critical surveillance applications, they are becoming popular in distributed systems for monitoring wide environments due to their low cost and maintenance requirements.
+In this work we expressly target low-power and ultra-low-power smart cameras, like \emph{SecSoC} (Security System on Chip), an innovative prototype based on a cluster of ReISC (Reduced Energy Instruction Set Computer) cores, designed and producted by STMicroelectronics, operating at clock rates of 82.5 MHz at 1.2V and sub-1MHz at 0.6V.
+These devices are battery-powered, and characterized by a constrained computational power and memory availability.
+While low-power cameras are typically not employed in critical surveillance applications, they are becoming popular in distributed systems for monitoring wide environments due to their low cost and maintenance requirements.
 %As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where smart-cameras can acquire and transmit images at regular time interval or upon requests.
@@
 Smart cameras have to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames.
-Moreover, false alarms represent a serious problem in low-power smart cameras since they result in energy waste because of useless data processing and network transmissions.
+Moreover, false alarms represent a serious problem in low-power smart cameras since they result in energy waste because of useless data processing and unnecessary network transmissions.
 % We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose.
@@
 \centering
-\includegraphics[width=0.8\linewidth]{Immagini/sequenze}
+\includegraphics[width=0.9\linewidth]{Immagini/sequenze}
 \caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences.}
@@
 \end{equation}
-where $w_t$ relates to a different viewpoint and/or camera orientation than $y_t$. \\
+where $w_t$ relates to a different viewpoint and/or camera orientation than $y_t$.
+
 The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect the time instant $\tau^*$ when tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.
-We assume that $T_0$ tampering-free frames and provided for training.
+We assume that $T_0$ tampering-free frames are provided for training.
 
@@
 
-Algorithm \ref{alg:DISPL} presents the proposed tampering-detection algorithm, which relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$).
-The former is meant to detect camera displacement, the latter blurring over camera lens (because blur cases a drop in the high-frequency content).
+Algorithm~\ref{alg:DISPL} presents the proposed tampering detection, which relies on two simple indicators: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$).
+The former is meant to detect camera displacement, the latter image blurring since it causes a drop in the high-frequency content.
 As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$.
-The employed segmentation is detailed in Section \ref{subsec:Segmentation}.
-\\
+The employed segmentation is detailed in Section~\ref{subsec:Segmentation}.
+
 For each $z_t$, the luma is separately computed within each of the $K$ regions:
@@
 \end{equation}
-where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, $\circledast$ denotes the 2d convolution and $\#(\cdot)$ the cardinality of a set.\\
+where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, $\circledast$ denotes the 2d convolution and $\#(\cdot)$ the cardinality of a set.
+
 \begin{figure}[tb]
@@
 \includegraphics[width=1\linewidth]{Immagini/indicatori}
-\caption{Example of displacement. On the left side of the image is reported the behavior of the average value of the luma, while on the right side there is its detrending.
-The event comes at frame $157$ and the corresponding value of $l(t)$ is depicted with a red star. 
-\textbf{(a)} Considering the average value of the luma in the whole scene. 
-\textbf{(b)} and \textbf{(c)} Considering the average value of the luma in a specific region.}
+\caption{Example of camera displacement.
+Left plots report the luma values, while right plots depict the detrended luma: we show both indicators computed over the full image (\textbf{a}) and on two reported regions (\textbf{b} and \textbf{c}).
+Red stars indicate a single displaced frame, which occurs at $t =157$.
+Displacement yields a peak in $l$ and two outliers in the sequence of detrended indicators $\partial l$, which can be clearly detected.
+}
 \label{fig:indicatori} 
@@
 %Tampering changes the sequence of indicators:
+
 Figure~\ref{fig:indicatori}(a) shows how a camera displacement affects the image intensities.
-First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) which is more clearly visible in the sequences $l_k(t)$ corresponding to the two displayed regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)).
+First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) which is also visible in the sequences $l_k(t)$ corresponding to the two displayed regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)).
 Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here.
-In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables, while here changes in the scene or in the illumination in non-tampered frames, yield indicators that follow an unpredictable trend.
-Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by temporal derivative (line~\ref{DISPL-Test5})
+In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables, while here indicators follow an unpredictable trend because of changes in the scene or in the illumination.
+Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by a temporal derivative (line~\ref{DISPL-Test5})
 \begin{equation}\label{eq:detrending}
@@
 \end{equation}
-and we similarly define $\partial g(t)$. \\
+and we similarly define $\partial g(t)$.
+
 The sequences of detrended indicators are reported in Figure~\ref{fig:indicatori}, and can be suitably monitored by defining the following confidence intervals:
@@
 \end{equation}
-where $\overline{\partial l}_k(t)$ denotes the mean and $\sigma_{l_k}$ the standard deviation of $\partial l$ over $R_k$, computed from the tampering-free frames provided for training (i.e.
+where $\overline{\partial l}_k(t)$ denotes the mean and $\sigma_{l_k}$ the standard deviation of $\partial l$ over $R_k$ (lines \ref{DISPL-Tr1} and \ref{DISPL-Tr7}), computed from the tampering-free frames provided for training (i.e.
 $z_t\,, t = 1, \dots, T_o$) and $\gamma_l >0$ is a tuning parameter.
 A similar interval is built for $\partial g$ (line~\ref{DISPL-Tr7}).
-\\
+
 During operations, any indicator falling outside its confidence region is considered anomalous (line~\ref{DISPL-Test6}).
@@
 This threshold was set to allow a single region not to change, as it typically happens in the sky when the camera is displaced, either horizontally or vertically.
-In contrast, any outlier in $\partial g$ would yield a blurring alert (line~\ref{DISPL-Test11}).\\
-It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, only the first and last tampered frames correspond to an outlier in the detrended indicators, see the detrended indicators in Figure~\ref{fig:indicatori}.
-Therefore, detrended indicators have to be monitored in a one-shot manner, targeting the detection of the first tampered-frame.
+In contrast, any outlier in $\partial g$ would yield a blurring alert (line~\ref{DISPL-Test11}).
+
+It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, a single tampered frame yield two outliers (as shown in Figure~\ref{fig:indicatori}) and only the first and the last of few consecutive tampered frames yield an outlier in the detrended indicators.
+This is the reason why detrended indicators are monitored in a one-shot manner, targeting the detection of the first tampered-frame.
 On the one hand, this one-shot solution can provide prompt detections, and this is particularly important at the low frame-rates we consider.
@@
 To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, and possibly combined with the proposed one-shot monitoring.
-
+%
 \begin{algorithm}[t]
@@
 \end{algorithm}
-
+%
 \subsection{Scene Segmentation}\label{subsec:Segmentation}
@@
 \end{equation}
-In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and column of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the average and standard deviation of the intensity at $x$ computed over time, and $\overline{g}(x)$ and $\sigma_g(x)$ are the average norm of the gradient and its standard deviation, computed overtime.
-These feature vectors are meant to cluster image regions according to their spatial appearance and temporal behavior over training frames.\\
+In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and column of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the mean and standard deviation of the intensity at $x$, computed over time, and $\overline{g}(x)$ and $\sigma_g(x)$ the mean of gradient norm and its standard deviation, computed overtime.
+These feature vectors are meant to cluster pixels in regions having similar spatial appearance and temporal behavior over training frames.
+
 Segmentation consists in clustering the feature vectors~\eqref{eq:featureVector} over the whole image by a weighted k-means~\cite{kottke1994motion}.
@@
 %The number of clusters is defined by the Calinski-Harabasz criterion~\cite{calinski1974dendrite}, which initializes the weighted k-means using different number of clusters, and then chooses the best one according a cost function\gi{Adriano controllare come lo chiama}.
-The number of clusters is defined by initializing the weighted k-means using different number of clusters and then choosing the best solution according to the Calinski-Harabasz criterion~\cite{calinski1974dendrite}.\\
+The number of clusters is defined by initializing the weighted k-means using different number of clusters and then choosing the best solution according to the Calinski-Harabasz criterion~\cite{calinski1974dendrite}.
+
 Finally, morphological image-processing operations are executed to remove boundaries between different regions, and eventually regions that are too small.
@@
 
-The most computationally demanding operations in the algorithm consists in computing of the indicators $g(t)$.
+The most computationally demanding operations of Algorithm \ref{alg:DISPL} consists in computing of the indicators $g(t)$.
 Assuming that derivatives in~\eqref{eq:normaGradiente} are computed by the Sobel filters, this requires $34$ operations per pixels\footnote{Execution times can be however  decreased whether hardware implementations of the FFT transform are available}.
@@
 Thus, monitoring regions has a negligible impact over the overall computational complexity.
-Segmentation has not been here considered because it is executed only when training the algorithm, and can be performed on an external device connected to the smart camera during configuration.
+Segmentation has not been here considered because it is executed only during the initial training phase, and can be performed on an external device connected to the smart camera during configuration.
 
@@
 %\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
-
+\vspace{-0.2cm}
 \section{Experiments}\label{sec:experiments}
 Experiments are meant to show the effectiveness of the proposed algorithm on real-world sequences, and to specifically assess the advantage of using segmentation for detecting camera displacements. 
+\vspace{-0.3cm}
 \subsubsection{Considered Methods}
@@
 Frame difference is considered as a baseline technique for displacement detection, while comparison against \emph{Luma-Voronoi} shows the advantages of setting regions depending of the image content.
-In the experiments on blurring detection we compare Algorithm \ref{alg:DISPL} against monitoring the gradient separately over image regions, as done for $\partial l_k$.
+\\
+In the experiments on blurring detection, we compare Algorithm \ref{alg:DISPL} (where $\partial g$ is monitored on the whole image), against monitoring $\partial g$ separately in image regions, as done for $\partial l_k$.
+\vspace{-0.3cm}
 \subsubsection{Datasets}
@@
 At last we have considered $9$ different scenarios, with $10$ displacement events and $10$ blurring events each, for a total of $180$ tampering events.%\gi{una footnote dicendo sequences can be released upon request?}
+\vspace{-0.3cm}
 \subsubsection{Performance Assessment}
-The typical figure of merit for assessing detection performance in one-shot monitoring methods are the ROC curves: in our case, these can be computed by varying $\gamma_l$ and $\gamma_g$ in a range of values.
+The typical figure of merit for assessing detection performance in one-shot monitoring methods are the Receiver-Operating Characteristic (ROC) curves: in our case, these can be computed by varying $\gamma_l$ and $\gamma_g$ in a range of values.
 In particular, each point in the ROC curve depicts a pair of values (1-SPECIFICITY, RECALL) defined as
 \[1-\text{SPECIFICITY}_\gamma = \frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}, \ \text{ and } \ \text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}\]
-where TP, FP, TN, FN represent the number of true positives (tampered frames correctly detected), of false positives (tampering-free frames detected), of true negatives (tampering-free frames not detected) and of false negatives (number of missed tampering frames), for a given value of threshold $\gamma_l$ (or $\gamma_g$).
+where TP, FP, TN, FN represent the number of true positives (tampered frames correctly detected), of false positives (tampering-free frames detected), of true negatives (tampering-free frames not detected) and of false negatives (missed tampered frames), respectively, for a given value of $\gamma_l$ (or $\gamma_g$).
+\vspace{-0.3cm}
 \subsubsection{Discussion}
-ROC curves and the AUC in Figure~\ref{fig:ROCdisplacement} caption indicate that camera displacements can be effectively detected.
-However, the best performance are achieved by Algorithm \ref{alg:DISPL} the (\emph{Frame-Difference} curve has not been reported because it was below these).
-This result indicate that separately monitoring the luma over image regions is an effective strategy and that the regions have to be carefully designed.
-Please note that, for the visualization sake, these ROC curves have been reported in a limited range of x and y axis.
-While the improvement might seem negligible we observe that Algorithm~\ref{alg:DISPL}, for having a recall of $98\%$, achieves $0.3\%$ false positive rates (i.e., 1-SPECIFICITY) than $0.7\%$ of the whole image solution or $1\%$ of the Voronoi regions solution.
-Having a solution which halves the false positives is important in a low-power context, considering the high number of frames to be processed.\\
-In contrast, Figure~\ref{fig:ROCdefocus} confirms that when monitoring $g$ it is more convenient to consider the whole scene at once.
-This is probably due to the fact that typically, regions $\{R_k\}$ do not include strong edges, which are the most informative to detect blurring.
+ROC curves and the Area Under the Curve (AUC) reported in Figure~\ref{fig:ROCdisplacement} and its caption indicate that camera displacements can be effectively detected by all the considered methods, and that Algorithm~\ref{alg:DISPL} achieves the best performance (the \emph{Frame-Difference} curve has not been reported because it was quite below the others).
+This result suggests that separately monitoring the luma over image regions is an effective strategy and that the regions have to be suitably estimated.
+For the visualization sake, these ROC curves have been reported in a limited range of both axis, and in particular we focus on low values of 1-SPECIFICITY, thus of false alarms.
+We observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $0.3\%$ of false alarms, which is half of the whole image solution ($0.7\%$) and less than one third of the Voronoi solution ($1\%$).
+Having few false alarms is important in a low-power context, considering the high number of frames that have to be processed.
+Most importantly, false alerts are halved basically at no computational overhead.
+
+In contrast, Figure~\ref{fig:ROCdefocus} confirms that when monitoring $g$ it is more convenient to consider the whole scene at once, as typically the regions $\{R_k\}$ do not include the most prominent edge of the scene, which are indeed the most informative part to detect blurring.
+This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.
 
@@
 \caption{\textbf{(a)} ROC curve for displacement detection.
-	There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
-	In particular, AUC of our solution is $0.9994$, while the solution which considers the whole image has an AUC of $0.9989$ and the one which considers Voronoi regions ha san AUC of $0.9985$.
+	%There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
+	The AUC corresponding to Algorithm~\ref{alg:DISPL} (Luma-Region) is $99.94\%$, the AUC of Luma-Total is $99.89\%$ the AUC of Luma-Voronoi is $99.85\%$, and the AUC of Frame-Difference is $99.44\%$ (axis have been zoomed for the visualization sake).
+Large AUC values indicate that these displacements can be successfully detected and that Algorithm~\ref{alg:DISPL} is the most effective one.
 	\textbf{(b)} ROC curve for blurring detection.
-	In this case considering adaptive regions instead of considering the whole image is not convenient.
-	In particular, AUC of the solution which considers adaptive regions is $0.9895$,  while the solution which considers the whole image has an AUC of $0.9996$.}
+	In this case, the Gradient-Total (AUC $99.96\%$) outperforms the Gradient-Regions (AUC $98.95\%$) because this latter do not include the most prominent edges in images, which are the most important information to detect blurring.
+This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.}
 \label{fig:ROC}
@@
 %Confronto tra intervalli di confidenza: 2 confronti
+\vspace{-0.3cm}
+\section{Conclusions}\label{sec:Conclusion}
+\vspace{-0.1cm}
+We presented a tampering detection algorithm that leverages a preliminary image partitioning to improve the detection of camera displacements, and that at the same time can detect blurring.
+The algorithm has low-computational complexity and is suited for low-power cameras operating at low frame rates.
 
-\section{Conclusion}\label{sec:Conclusion}
-We presented a tampering detection algorithm that leverages a preliminary image partition to improve the detection of camera displacements, while at the same time can detect blurring.
-This algorithm has low-computational complexity and is suited for low-power cameras operating at low frame rates.
-
-Ongoing work concern the extension of our solution to other types of tampering, as occlusions or imaging sensor degradations, as well as the integration of this one-shot technique with sequential monitoring scheme such as~\cite{alippi2010detecting}, to leverage the fact that tampering might last more than one frame in the detection.
+Ongoing work concern the extension of our solution to other types of tampering, such as occlusions or imaging sensor degradations, as well as the integration of this one-shot technique with sequential monitoring scheme such as~\cite{alippi2010detecting}, to leverage the fact that tampering might last more than one frame in the detection.
 Moreover, we will investigate different segmentation strategy and test our algorithm on a larger dataset.

### 2015-06-18 paper.tex
Modificata immagine sequenze e modificata descrizione del dataset

@@
 \includegraphics[width=0.9\linewidth]{Immagini/sequenze}
-\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences.}
+\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 3-rd frame, created using gaussian filtering; \textbf{(c)} displacement event on 3-rd frame, created moving the crop of the original image from the center.}
 \label{fig:sequences}
@@
 \subsubsection{Datasets}
-We acquired two datasets: the first dataset contains  sequences of  frames taken from webcams monitoring urban areas (e.g., see Figure \ref{fig:sequences}(a)). 
-The second dataset was recorded using a Raspberry Pi Model B+ \cite{raspberry} mounting a camera module. 
-In the first dataset tampering was synthetically generated by blurring frames (see, Figure \ref{fig:sequences}(b)) or concatenating frames from different sequences depicting similar environments (see Figure \ref{fig:sequences}(c)).
-Few sequences of the first dataset contain real tampering events  (Figure \ref{fig:indicatori}). 
-In the second dataset, tampering was introduced by moving the device or spraying water on the camera lens. 
-At last we have considered $9$ different scenarios, with $10$ displacement events and $10$ blurring events each, for a total of $180$ tampering events.%\gi{una footnote dicendo sequences can be released upon request?}
+The considered dataset contains  sequences of  frames taken from webcams monitoring urban areas (e.g., see Figure \ref{fig:sequences}(a)), in which we've taken the central crop of each image (about $3/4$ of the original image).
+%The second dataset was recorded using a Raspberry Pi Model B+ \cite{raspberry} mounting a camera module. 
+In order to have a wide and sound dataset, the tampering events have been made synthetically, blurring the frames by gaussian kernels (see, Figure \ref{fig:sequences}(b)) or moving the crop of the frames in order to simulate displacements (see Figure \ref{fig:sequences}(c)).
+%Few sequences of the first dataset contain real tampering events  (Figure \ref{fig:indicatori}). 
+%In the second dataset, tampering was introduced by moving the device or spraying water on the camera lens. 
+At last we have considered $8$ different scenarios, for a total of $24410$ frames considered, in which we've inserted $5\%$ of displacement events ($1224$ frames) and $5\%$ of blurring events ($1224$ frames).\footnote{Sequences can be released upon request}
 \vspace{-0.3cm}
 \subsubsection{Performance Assessment}
-The typical figure of merit for assessing detection performance in one-shot monitoring methods are the Receiver-Operating Characteristic (ROC) curves: in our case, these can be computed by varying $\gamma_l$ and $\gamma_g$ in a range of values.
+The typical figure of merit for assessing detection performance in one-shot monitoring methods are the Receiver-Operating Characteristic (ROC) curves: in our case, these can be computed by varying $\gamma_l$, $\gamma_d$ and $\gamma_g$ in a range of values.
 In particular, each point in the ROC curve depicts a pair of values (1-SPECIFICITY, RECALL) defined as
 \[1-\text{SPECIFICITY}_\gamma = \frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}, \ \text{ and } \ \text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}\]
-where TP, FP, TN, FN represent the number of true positives (tampered frames correctly detected), of false positives (tampering-free frames detected), of true negatives (tampering-free frames not detected) and of false negatives (missed tampered frames), respectively, for a given value of $\gamma_l$ (or $\gamma_g$).
+where TP, FP, TN, FN represent the number of true positives (tampered frames correctly detected), of false positives (tampering-free frames detected), of true negatives (tampering-free frames not detected) and of false negatives (missed tampered frames), respectively, for a given value of $\gamma$.
 \vspace{-0.3cm}

### 2015-06-19 paper.tex
Modificata Fig.2
ROC nuove!
Abstract e Section 1

@@
 \begin{document}
-\title{Tampering Detection In Low-Power Smart Cameras}
+\title{Tampering Detection In Low-Power \\ Smart Cameras}
 
@@
 A desirable feature in smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene.
-No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attacks (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures.
+No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attacks (e.g., occlusions or the device displacement), these have to be promptly detected to possibly activate countermeasures.
 Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at full-speed  frame-rates, nor use sophisticated image-analysis algorithms.
@@
 %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
-\keywords{tampering detection, blurring detection, displacement detection}
+\keywords{tampering detection, smart cameras, displacement detection, blurring detection}
 \end{abstract}
@@
 When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as shown in Figure \ref{fig:pioggia}, or in partial occlusions of the scene, as in Figure \ref{fig:neve}.
-Similarly, intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be useless for monitoring purposes.
+Similarly, intentional attacks like displacing the camera Figure \ref{fig:pioggia} \gi{Adriano SISTEMA CON LE NUOVE FIGURE}, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be useless for monitoring purposes.
 We refer to these events/attacks as tampering.
-In some cases, tampering is easy to detect, e.g., when the camera integrity is affected and the device goes out-of-order.
-However, in many other situations, namely when the device is not physically damaged, tampering detection is not straightforward, and sometimes image analysis is the only viable option.
+In many situations, tampering is not straightforward to detect (e.g.
+when the camera is not physically damaged and does not go out-of-order), and image analysis techniques are the only viable option.
 
@@
 Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly report alerts.
-In fact, tampering prevents the correct interpretation of the scene: even a mild blur might hinder the identification of important details such as licence plates.
+In fact, tampering prevents the correct interpretation of the scene: for instance, even a mild blur might hinder the identification of important details such as licence plates.
 Surveillance cameras are typically connected to the power supply and operate at normal frame-rates (e.g.
 around few frames per second): in these conditions, several tampering-detection algorithms have been presented.
-In particular, camera displacement and occlusions are typically detected by estimating the background or computing frame difference, while blurring is instead detected by monitoring the high-frequency components of images, as these are expected to drop.
+In particular, camera displacement and occlusions are typically detected by comparing the current frame against an estimate of the scene background, while blurring is instead detected by monitoring the high-frequency components of images, as these are expected to drop.
 In~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain.
@@
 A frame buffer rather than an explicit background is used~\cite{ribnick2006real}.
-All these solutions, including~\cite{harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.
+All these above solutions, including~\cite{harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.
 
-In this work we expressly target low-power and ultra-low-power smart cameras, like \emph{SecSoC} (Security System on Chip), an innovative prototype based on a cluster of ReISC (Reduced Energy Instruction Set Computer) cores, designed and producted by STMicroelectronics, operating at clock rates of 82.5 MHz at 1.2V and sub-1MHz at 0.6V.
-These devices are battery-powered, and characterized by a constrained computational power and memory availability.
-While low-power cameras are typically not employed in critical surveillance applications, they are becoming popular in distributed systems for monitoring wide environments due to their low cost and maintenance requirements.
+In this work we expressly target low-power and ultra-low-power smart cameras, like \emph{SecSoC} (Security System on Chip), an innovative prototype based on a cluster of ReISC (Reduced Energy Instruction Set Computer) cores, designed and produced by STMicroelectronics.
+SecSoC is a battery-powered device, characterized by a constrained computational power (clock rates 82.5 MHz at 1.2V, and sub-1MHz at 0.6V) and reduced memory (1.25 MB).
+While low-power cameras are not meant for critical surveillance applications, they might be easily employed in distributed systems for monitoring wide environments thanks to their low cost and maintenance requirements.
 %As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where smart-cameras can acquire and transmit images at regular time interval or upon requests.
@@
 % Why is it hard? (E.g., why do naive approaches fail?)
-Tampering detection in low-power smart cameras has not been much investigated in the literature~\cite{alippi2010detecting}, also because it is more challenging than in conventional surveillance cameras~\cite{perrig2004security}.
+Tampering detection in low-power cameras (eventually organized in wireless sensor network) is more challenging than in conventional surveillance systems, and this problem has not been much investigated so far~\cite{perrig2004security,alippi2010detecting}.
 Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), thus the acquired sequence does not evolve smoothly.
-This prevents the use of learned background models and the analysis of foreground variations.
-In fact, when dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions, yielding sequences like those in Figure~\ref{fig:sequences}.
-Smart cameras have to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames.
-Moreover, false alarms represent a serious problem in low-power smart cameras since they result in energy waste because of useless data processing and unnecessary network transmissions.
+These aspects prevent the use of learned background models and the analysis of foreground variations.
+In fact, in dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions (see Figure~\ref{fig:sequences}).
+Smart cameras have then to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames.
+%Moreover, false alarms represent a serious problem in low-power smart cameras since they result in energy waste because of useless data processing and unnecessary network transmissions.
 % We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose.
@@
 We address the detection of camera blurring and displacement, which are presented in Section \ref{sec:probForm}.
-The proposed algorithm, detailed in Section~\ref{sec:propSol}, relies on two indicators: the average image intensity is used to detect displacement, and the average gradient norm to detect blurring.
-These indicators can be easily computed and monitored by an outlier-detection technique in a low-power smart camera.
-In particular, we show that separately monitoring the indicators over different regions of the image can substantially improve the displacement-detection performance.
-Remarkably, these regions can be preliminarily defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t.
+The proposed algorithm (Section~\ref{sec:propSol}) relies on indicators that can be easily computed in low-power smart cameras, such as the average image intensity or frame difference (to detect camera displacement), and the average gradient norm (to detect blurring), and detects tampering as frames yielding outliers.
+In particular, we show that separately monitoring indicators over different regions of the image can substantially improve the displacement-detection performance.
+Remarkably, image regions can be preliminarily defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t.
 monitoring the whole image (Section \ref{subsec:coputationalComplexity}).
@@
 % Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.
-
 \begin{figure}[t]

### 2015-06-19 paper.tex
nothing

@@
 \centering
-\includegraphics[width=0.9\linewidth]{Immagini/sequenze}
-\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 3-rd frame, created using gaussian filtering; \textbf{(c)} displacement event on 3-rd frame, created moving the crop of the original image from the center.}
+\includegraphics[width=1\linewidth]{Immagini/sequenze}
+\caption{Sequences taken from webcams: \textbf{(a)} defocus event on 3-rd frame, created using gaussian filtering; \textbf{(b)} displacement event on 3-rd frame, created moving the crop of the original image from the center.}
 \label{fig:sequences}
@@
 \centering
-\subfigure[]{\label{fig:ROCdisplacement}\includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement}}
-\subfigure[]{\label{fig:ROCdefocus} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus}}
+\includegraphics[width=1\linewidth]{Immagini/curve}
 \caption{\textbf{(a)} ROC curve for displacement detection.

### 2015-06-19 paper.tex
Fig 1 modificata
ho inserito un po' di modifiche e qualche nota a tutto tranne section 4. Cerco di darci un'altra passata nel WE.

@@
 % Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.
-\begin{figure}[t]
-\centering
-\includegraphics[width=1\linewidth]{Immagini/sequenze}
-\caption{Sequences taken from webcams: \textbf{(a)} defocus event on 3-rd frame, created using gaussian filtering; \textbf{(b)} displacement event on 3-rd frame, created moving the crop of the original image from the center.}
-\label{fig:sequences}
-\end{figure}
 
@@
 
+\begin{figure}[t]
+\centering
+\includegraphics[width=1\linewidth]{Immagini/sequenze}
+\caption{Sequences taken from webcams: \textbf{(a)} defocus event on 3-rd frame, created using gaussian filtering; \textbf{(b)} displacement event on 3-rd frame, created moving the crop of the original image from the center.}
+\label{fig:sequences}
+\end{figure}
+
 \section{Tampering Detection}\label{sec:propSol}
 
-Algorithm~\ref{alg:DISPL} presents the proposed tampering detection, which relies on two simple indicators: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$).
-The former is meant to detect camera displacement, the latter image blurring since it causes a drop in the high-frequency content.
+Algorithm~\ref{alg:DISPL} presents the proposed tampering detection, which relies on simple indicators such as the average intensity (the \emph{luma}, denoted by $l$), the average frame difference (denoted by $d$) and the average norm of the gradient (denoted by $g$).
+The first two are meant to detect camera displacement, while the latter image blurring, as this is expected to cause a drop in the high-frequency content.
 As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$.
@@
 \end{equation}
+and, similarly, we define the average frame difference
+\begin{equation}\label{eq:FDRegions}
+d_k(t) =\frac{1}{\#R_k} \sum_{x \in R_k} \left(z_t(x) - z_{t-1}(x) \right)^2, \ \ k = 1, \dots, K\,,
+\end{equation}
 while the average gradient norm is computed over the whole image
@@
 where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, $\circledast$ denotes the 2d convolution and $\#(\cdot)$ the cardinality of a set.
+The choice of using luma or frame difference also depends on the memory availability to store the frame.
+In what follows and in Algorithm~\ref{alg:DISPL} we will consider only the luma $l$ even though the same applies to frame difference $d$.
 
@@
 %Tampering changes the sequence of indicators:
-
-Figure~\ref{fig:indicatori}(a) shows how a camera displacement affects the image intensities.
-First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) which is also visible in the sequences $l_k(t)$ corresponding to the two displayed regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)).
+Figure~\ref{fig:indicatori}(a) shows how a camera displacement affects the luma.
+First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) and that this change is also visible in the indicators $l_k(t)$ for the regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)).
 Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here.
@@
 \end{equation}
-and we similarly define $\partial g(t)$.
+and we similarly define $\overline{\partial d}_k(t)$ and $\partial g(t)$.
 
-The sequences of detrended indicators are reported in Figure~\ref{fig:indicatori}, and can be suitably monitored by defining the following confidence intervals:
-\begin{equation}\label{eq:confidenceRegions}
- [\overline{\partial l}_k(t) - \gamma_l \sigma_{l_k}, \overline{\partial l}_k(t) + \gamma_l \sigma_{l_k}], \ \  k = 1,\dots,K\,,
-\end{equation}
-where $\overline{\partial l}_k(t)$ denotes the mean and $\sigma_{l_k}$ the standard deviation of $\partial l$ over $R_k$ (lines \ref{DISPL-Tr1} and \ref{DISPL-Tr7}), computed from the tampering-free frames provided for training (i.e.
-$z_t\,, t = 1, \dots, T_o$) and $\gamma_l >0$ is a tuning parameter.
-A similar interval is built for $\partial g$ (line~\ref{DISPL-Tr7}).
-
-During operations, any indicator falling outside its confidence region is considered anomalous (line~\ref{DISPL-Test6}).
-To raise a camera-displacement alert we require at least $K-1$ indicators $\partial l_k$ to yield an outlier (line~\ref{DISPL-Test8}).
-This threshold was set to allow a single region not to change, as it typically happens in the sky when the camera is displaced, either horizontally or vertically.
-In contrast, any outlier in $\partial g$ would yield a blurring alert (line~\ref{DISPL-Test11}).
-
-It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, a single tampered frame yield two outliers (as shown in Figure~\ref{fig:indicatori}) and only the first and the last of few consecutive tampered frames yield an outlier in the detrended indicators.
-This is the reason why detrended indicators are monitored in a one-shot manner, targeting the detection of the first tampered-frame.
-On the one hand, this one-shot solution can provide prompt detections, and this is particularly important at the low frame-rates we consider.
-On the other hand, one-shot monitoring disregards the fact that tampering might persist over several frames.
-To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, and possibly combined with the proposed one-shot monitoring.
-%
 \begin{algorithm}[t]
@@
 		\lnl{DISPL-Test2} Get frame $z_t$, set $n_l =0$;\\
+		\lnl{DISPL-Test10} Compute $\partial g(t)$\\
+			\lnl{DISPL-Test11} \If{$\partial g(t) < -\gamma_g \sigma_{g} \vee \partial l(t) > \gamma_g \sigma_{g} $}{
+					\lnl{DISPL-Test9} tampering (probably blurring) has occurred in $z_t$ \\
+									}
 		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
@@
 		}
-		\lnl{DISPL-Test8} \If{$n_l\geq K-1$}{
+		\lnl{DISPL-Test8} \If{$n_l\geq \Gamma_l$}{
 			\lnl{DISPL-Test9} tampering (probably displacement or occlusion) has occurred in $z_t$ \\
-		}
-		\lnl{DISPL-Test10} Compute $\partial g(t)$\\
-		\lnl{DISPL-Test11} \If{$\partial g(t) < -\gamma_g \sigma_{g} \vee \partial l(t) > \gamma_g \sigma_{g} $}{
-				\lnl{DISPL-Test9} tampering (probably blurring) has occurred in $z_t$ \\
 		}
@@
 \end{algorithm}
+
+The sequences of detrended indicators (reported in Figure~\ref{fig:indicatori}) can be suitably monitored by the following confidence intervals:
+\begin{equation}\label{eq:confidenceRegions}
+ [\overline{\partial l}_k(t) - \gamma_l \sigma_{l_k}, \overline{\partial l}_k(t) + \gamma_l \sigma_{l_k}], \ \  k = 1,\dots,K\,,
+\end{equation}
+where $\overline{\partial l}_k(t)$ denotes the mean and $\sigma_{l_k}$ the standard deviation of $\partial l$ over $R_k$ (lines \ref{DISPL-Tr1} and \ref{DISPL-Tr7}), computed from tampering-free frames provided for training (i.e.
+$z_t\,, t = 1, \dots, T_o$) and $\gamma_l >0$ is a tuning parameter.
+A similar interval is built for  $\overline{\partial d}_k(t)$ and $\partial g$ (line~\ref{DISPL-Tr7}).
+
+During operations, any indicator falling outside its confidence region is considered an outlier (line~\ref{DISPL-Test6}): any outlier in $\partial g$ yields a blurring alert (line~\ref{DISPL-Test11}).
+Camera displacements alerts are raised when at least $\Gamma_l$ indicators $\partial l_k$  simultaneously yield outliers (line~\ref{DISPL-Test8}).
+The threshold $\Gamma_l$ together with $\gamma_l$ determine the displacement-detection promptness.
+The extreme cases are $\Gamma_l = 1$, where a displacement alert raises at any region firing an outlier, and $\Gamma_l = K-1$ where all but one region have to simultaneously fire an outlier (it is worth excluding a region since, for instance the sky-region typically does not change when the camera is displaced, either horizontally or vertically).
+
+It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, a single tampered frame yield two outliers (as shown in Figure~\ref{fig:indicatori}) and only the first and the last of few consecutive tampered frames yield an outlier in the detrended indicators.
+This is the reason why detrended indicators are monitored in a one-shot manner, targeting the detection of the first tampered-frame.
+On the one hand, this one-shot solution can provide prompt detections, and this is particularly important at the low frame-rates we consider.
+On the other hand, one-shot monitoring disregards the fact that tampering might persist over several frames.
+To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, and possibly combined with the proposed one-shot monitoring.
+%
 %
@@
 
-Segmentation consists in clustering the feature vectors~\eqref{eq:featureVector} over the whole image by a weighted k-means~\cite{kottke1994motion}.
-In particular, since the components of the feature vector span very different ranges, we compute the distance between a feature vector and a cluster centroid by scaling the Euclidean distance of a weight inversely proportional to the standard deviation over the cluster.
+As in superpixel methods \cite{Susstrunk2012}, segmentation is performed by k-means clustering.
+In this case feature vectors~\eqref{eq:featureVector} over the whole image are clustered by a weighted k-means~\cite{kottke1994motion} that scales the Euclidean distance between a feature vector and a cluster centroid of a weight that is inversely proportional to the standard deviation over the cluster.
+This scaling compensates the fact that different components of the feature vector might span very different ranges.
+
 %\gi{Adriano, controlla bene}
@@
 
+\gi{ADD REFERENCE TO LOW-POWER, MEMORIA PER FARE FD}
 The most computationally demanding operations of Algorithm \ref{alg:DISPL} consists in computing of the indicators $g(t)$.
@@
 Segmentation has not been here considered because it is executed only during the initial training phase, and can be performed on an external device connected to the smart camera during configuration.
-
-%FRAME DIFFERENCE:
-%\begin{equation}
-%	\label{eq:frameDiffReg}
-%	\varphi^k(t) = \frac{\sum_{x \in R_k}(z_t(x) - z_{t-1}(x))^2}{|R_k|}, k=1,\dots,K.
-%\end{equation}
-
-%\begin{algorithm}[tp]
-%	% \SetAlgoNoLine
-%	\LinesNumbered
-%	\SetAlgoNlRelativeSize{0}
-%	\SetNlSty{small}{}{.}
-%	\textbf{Configuration}:\\
-%	\lnl{DEF-Tr0} Extract regions $\{R_k\}, k=1,\dots,K$  \\
-%	\lnl{DEF-Tr1} \For{$t=1,\dots,T_{o}$}
-%	{	\lnl{DEF-Tr2} Acquire frame $z_t$ \\
-%		\lnl{DEF-Tr2a} \For{$k=1,\dots,K$}{
-%			\lnl{DEF-Tr3a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
-%		}
-%		\lnl{DEF-Tr3} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
-%	}
-%	\lnl{DEF-Tr6a} \For{$k=1,\dots,K$}{
-%		\lnl{DEF-Tr7a} Define thresholds $\Gamma_{min}^k$ and $\Gamma_{max}^k$\\
-%	}
-%	\lnl{DEF-Tr5} Define CDT parameters on $g(t)$ variance\\
-%	\textbf{Operational phase}:\\
-%	\lnl{DEF-Test1} \For{$t=T_{o},\dots,\infty$}{
-%		\lnl{DEF-Test2} Acquire frame $z_t$ \\
-%		\lnl{DEF-Tr3b} Compute $g(t)$, $\frac{\partial g}{\partial t}(t)$ \\
-%		\lnl{DEF-Test2a} $n=0$\\
-%		\lnl{DEF-Test4a} \For{$k=1,\dots,K$}{
-%			\lnl{DEF-Test5a} Compute $g^k(t)$, $\frac{\partial g^k}{\partial t}(t)$ for the region $R_k$\\
-%			\lnl{DEF-Test6a} \If{$\frac{\partial g^k}{\partial t}(t) < \Gamma_{min}^k \vee \frac{\partial g^k}{\partial t}(t) > \Gamma_{max}^k $}{
-%				\lnl{DEF-Test7a} $n=n+1$\\
-%			}
-%		}
-%		\lnl{DEF-Test81} \If{$n \geq K-1$}{
-%			\lnl{DEF-Test91} $z_t$ is a defocused frame\\
-%		}
-%		\lnl{DEF-Test8} \If{CDT detect a change on $g(t)$ variance}{
-%			\lnl{DEF-Test9} $z_t$ is a defocused frame\\
-%		}
-%	}   
-%	\caption{Blur detection algorithm}
-%	\label{alg:DEFOCUS}
-%\end{algorithm}
 
@@
 
-Ongoing work concern the extension of our solution to other types of tampering, such as occlusions or imaging sensor degradations, as well as the integration of this one-shot technique with sequential monitoring scheme such as~\cite{alippi2010detecting}, to leverage the fact that tampering might last more than one frame in the detection.
-Moreover, we will investigate different segmentation strategy and test our algorithm on a larger dataset.
+Ongoing work concern the extension of our solution to other types of tampering, like degradations of the imaging sensor, as well as the integration of this one-shot technique with sequential monitoring scheme such as~\cite{alippi2010detecting}, to leverage the fact that tampering might last more than one frame in the detection.
+Moreover, we will investigate different segmentation strategies and the application of superpixels methods~\cite{Susstrunk2012} to video sequences by including temporal features in the clustered vectors.
 

### 2015-06-20 paper.tex
Sections 1-3

@@
 A desirable feature in smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene.
-No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attacks (e.g., occlusions or the device displacement), these have to be promptly detected to possibly activate countermeasures.
-Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at full-speed  frame-rates, nor use sophisticated image-analysis algorithms.
+No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attacks (e.g., occlusions or device displacements), these have to be promptly detected to possibly activate countermeasures.
+Tampering detection is particularly challenging in battery-powered cameras, where it is not possible to acquire images at full-speed frame-rates, nor use sophisticated image-analysis algorithms.
 
-We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme.
-Any frame yielding anomalous indicator is detected as a tampered frame.
+We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras.
+The algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme: any frame yielding an outlier is detected as tampered.
 Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to improve the detection of camera displacements.
-Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity.
+Experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity.
 %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
@@
 % What is the problem?
-When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as shown in Figure \ref{fig:pioggia}, or in partial occlusions of the scene, as in Figure \ref{fig:neve}.
-Similarly, intentional attacks like displacing the camera Figure \ref{fig:pioggia} \gi{Adriano SISTEMA CON LE NUOVE FIGURE}, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be useless for monitoring purposes.
+Cameras operating outdoor and in harsh environments are exposed to atmospheric phenomena and intentional attacks that might prevent the correct image acquisition.
+Rain, snow, dust lying on the camera lens cause blurry (Figure \ref{fig:pioggia}) or partially occluded (Figure \ref{fig:neve}) pictures, while wind might displace the camera (Figure \ref{fig:displ1} -\ref{fig:displ2}).
+Similarly, an attacker can intentionally change the camera focus, spray some opaque or glossy liquid, displace or occlude the camera.
 We refer to these events/attacks as tampering.
@@
 \centering
-\subfigure[]{\label{fig:pioggia} \includegraphics[width=0.3\linewidth]{Immagini/pioggia}}
-\subfigure[]{\label{fig:neve} \includegraphics[width=0.3\linewidth]{Immagini/neve}}
-\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena. \textbf{(a)} Blur due to rain drops on the camera lens. \textbf{(b)} Occlusion due to snow on the camera lens.}
+\subfigure[]{\label{fig:pioggia} \includegraphics[width=0.225\linewidth]{Immagini/pioggia}}
+\subfigure[]{\label{fig:neve} \includegraphics[width=0.225\linewidth]{Immagini/neve}}
+\subfigure[]{\label{fig:displ1} \includegraphics[width=0.225\linewidth]{Immagini/displ1}}
+\subfigure[]{\label{fig:displ2} \includegraphics[width=0.225\linewidth]{Immagini/displ2}}
+\caption{Examples of tampering events due to atmospheric phenomena.
+\textbf{(a)} rain drops resulting in a blurry picture, \textbf{(b)} snow partially occluding the camera view, \textbf{(c)}-\textbf{(d)} wind displacing the camera.}
 \label{fig:tampering}
@@
 % Why is it interesting and important?
-Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly report alerts.
-In fact, tampering prevents the correct interpretation of the scene: for instance, even a mild blur might hinder the identification of important details such as licence plates.
-Surveillance cameras are typically connected to the power supply and operate at normal frame-rates (e.g.
-around few frames per second): in these conditions, several tampering-detection algorithms have been presented.
-In particular, camera displacement and occlusions are typically detected by comparing the current frame against an estimate of the scene background, while blurring is instead detected by monitoring the high-frequency components of images, as these are expected to drop.
-In~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain.
+Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, which are expected to autonomously detect any tampering, and promptly report alerts.
+Tampering, in fact, might result in images that are useless for monitoring purposes: even a mild blur might hinder the identification of important details such as licence plates.
+Surveillance cameras most often operate at normal frame-rates (e.g.
+around few frames per second) and are connected to the power supply.
+Tampering-detection for surveillance cameras have been quite investigated in the literature; in particular, camera displacements and occlusions are typically detected by comparing the current frame against an estimate of the scene background, while blurring is detected by monitoring the high-frequency components of images, as these are expected to drop.
+In~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing the histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain.
 Two background models, estimated over different time intervals, are used in~\cite{saglam2009real}.
@@
 Background estimates and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while the number of SURF~\cite{bay2006surf} keypoints is used to detect blurring.
-A frame buffer rather than an explicit background is used~\cite{ribnick2006real}.
-All these above solutions, including~\cite{harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.
+A buffer of recent frames rather than an explicit background is used~\cite{ribnick2006real}.
+All the above solutions, including~\cite{harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.
 
@@
 SecSoC is a battery-powered device, characterized by a constrained computational power (clock rates 82.5 MHz at 1.2V, and sub-1MHz at 0.6V) and reduced memory (1.25 MB).
-While low-power cameras are not meant for critical surveillance applications, they might be easily employed in distributed systems for monitoring wide environments thanks to their low cost and maintenance requirements.
+While low-power cameras are not meant for critical surveillance applications, they might be easily employed for monitoring wide environments thanks to their low cost and maintenance requirements.
 %As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where smart-cameras can acquire and transmit images at regular time interval or upon requests.
@@
 These aspects prevent the use of learned background models and the analysis of foreground variations.
-In fact, in dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions (see Figure~\ref{fig:sequences}).
-Smart cameras have then to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames.
+In fact, when dynamic environments are monitored at low frame-rates, changes in the scene and in the light conditions might produce consecutive frames that are very different (see Figure~\ref{fig:sequences}).
+Smart cameras have to promptly distinguish between \emph{normal} changes (due to illumination or movements in the scene), and changes due to camera tampering, to raise an alert and eventually avoid the transmission of tampered frames.
 %Moreover, false alarms represent a serious problem in low-power smart cameras since they result in energy waste because of useless data processing and unnecessary network transmissions.
@@
 %What are the key components of my approach and results? Also include any specific limitations.
-We address the detection of camera blurring and displacement, which are presented in Section \ref{sec:probForm}.
-The proposed algorithm (Section~\ref{sec:propSol}) relies on indicators that can be easily computed in low-power smart cameras, such as the average image intensity or frame difference (to detect camera displacement), and the average gradient norm (to detect blurring), and detects tampering as frames yielding outliers.
-In particular, we show that separately monitoring indicators over different regions of the image can substantially improve the displacement-detection performance.
-Remarkably, image regions can be preliminarily defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t.
+We address the detection of camera blurring and displacement (Section \ref{sec:probForm}), and propose an algorithm (Section~\ref{sec:propSol}) that relies on indicators that can be easily computed in low-power smart cameras.
+We monitor the average image intensity or frame difference (to detect camera displacement), and the average gradient norm (to detect blurring), and consider any frame yielding an outlier as tampered.
+In particular, we show that separately monitoring these indicators over different regions of the image can substantially improve the displacement-detection performance.
+Image regions are defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t.
 monitoring the whole image (Section \ref{subsec:coputationalComplexity}).
-Thus, our solution represents a prompt trigger that can be possibly combined with other sequential monitoring techniques.
-Experiments in Section \ref{sec:experiments} show that leveraging image regions can substantially improve the displacement-detection performance.
+Our algorithm thus represents a prompt trigger, to be possibly combined with other sequential monitoring techniques.
+Experiments (Section \ref{sec:experiments}) show that leveraging image regions can substantially improve the displacement-detection performance.
 % Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.
@@
 \label{eq:observationModel}
-z_t(x)=\mathcal{D}_t[y_t](x) \quad \forall x \in X
+z_t(x)=\mathcal{D}_t[y_t](x), \quad \forall x \in X
 \end{equation}
-where $\mathcal{D}_t$ denotes an operator that transforms the original image $y_t$ in the acquired frame $z_t$; $X \subset \mathbb{Z}^2$ denotes the regular pixels grid and $x\in \mathbb{Z}^2$ indicates the pixel coordinates.
+where $\mathcal{D}_t$ denotes an operator transforming the original image $y_t$, and $x\in \mathbb{Z}^2$ indicates the pixel coordinates belonging to the regular pixel grid $X \subset \mathbb{Z}^2$.
 As far as there are no tampering attacks/events,
@@
 \label{eq:no_tampering}
-\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x) \quad \forall x \in X
+\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x), \quad \forall x \in X
 \end{equation}
-where $\eta_t$ is a random variable accounting for image noise and $y_t$ are acquired from the same viewpoint and camera orientation, even though $y_t$ might show different content because of scene changes.\\
+where $\eta_t$ is a random variable accounting for image noise and $y_t$ are acquired from the same viewpoint and camera orientation, even though typically $y_t \neq y_{t-1}$ because the depicted scene changes.\\
 When, at time $\tau^*$, an external disturbance introduces \emph{blurring}, the image $y_t$ is degraded by an unknown blur operator, and $z_t$ becomes
@@
 \end{equation}
-where $w_t$ relates to a different viewpoint and/or camera orientation than $y_t$.
+where $w_t$ and $y_t$ refer to different viewpoints and/or camera orientations.
 
@@
 \includegraphics[width=1\linewidth]{Immagini/sequenze}
-\caption{Sequences taken from webcams: \textbf{(a)} defocus event on 3-rd frame, created using gaussian filtering; \textbf{(b)} displacement event on 3-rd frame, created moving the crop of the original image from the center.}
+\caption{Examples of blurring and displacement synthetically introduced: \textbf{(a)} the 3-rd frame is blurred by a convolution against a Gaussian filter; \textbf{(b)} the 3-rd frame has been displaced by cropping a different region in the original image.}
 \label{fig:sequences}
@@
 Algorithm~\ref{alg:DISPL} presents the proposed tampering detection, which relies on simple indicators such as the average intensity (the \emph{luma}, denoted by $l$), the average frame difference (denoted by $d$) and the average norm of the gradient (denoted by $g$).
-The first two are meant to detect camera displacement, while the latter image blurring, as this is expected to cause a drop in the high-frequency content.
-As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$.
+The first two are meant to detect camera displacements, which would substantially change the image content, while the latter the blurring, which would cause a drop in the high-frequency content.
+As anticipated in Section~\ref{sec:introduction}, during the initial configuration the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$.
 The employed segmentation is detailed in Section~\ref{subsec:Segmentation}.
 
-For each $z_t$, the luma is separately computed within each of the $K$ regions:
-\begin{equation}\label{eq:lumaRegions}
-l_k(t) =\frac{1}{\#R_k} \sum_{x \in R_k} z_t(x), \ \ k = 1, \dots, K\,,
-\end{equation}
-and, similarly, we define the average frame difference
-\begin{equation}\label{eq:FDRegions}
-d_k(t) =\frac{1}{\#R_k} \sum_{x \in R_k} \left(z_t(x) - z_{t-1}(x) \right)^2, \ \ k = 1, \dots, K\,,
-\end{equation}
-while the average gradient norm is computed over the whole image
+For each $z_t$, we compute the luma and frame difference separately on each of the $K$ regions,
+\begin{eqnarray}
+\vspace{-0.5cm}
+l_k(t) &=& \frac{1}{\#R_k} \sum_{x \in R_k} z_t(x) \label{eq:lumaRegions}\\
+d_k(t) &=&\frac{1}{\#R_k} \sum_{x \in R_k} \left(z_t(x) - z_{t-1}(x) \right)^2, \ \ k = 1, \dots, K\,,\label{eq:FDRegions}
+\vspace{-0.5cm}
+\end{eqnarray}
+where $\#(\cdot)$ denotes the cardinality of a set.
+We can simultaneously monitor both luma and frame difference, even though computing $d_k$ requires to store the previous frame and this also depends on memory availability of the device.
+In what follows, including Algorithm~\ref{alg:DISPL}, we consider only the luma $l$ but the same procedures apply to the frame difference $d$ as well.
+The average gradient norm is instead computed over the whole image
 \begin{equation}
@@
 \end{equation}
-where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, $\circledast$ denotes the 2d convolution and $\#(\cdot)$ the cardinality of a set.
-The choice of using luma or frame difference also depends on the memory availability to store the frame.
-In what follows and in Algorithm~\ref{alg:DISPL} we will consider only the luma $l$ even though the same applies to frame difference $d$.
+where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, and $\circledast$ denotes the 2d convolution.
 
@@
 In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables, while here indicators follow an unpredictable trend because of changes in the scene or in the illumination.
-Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by a temporal derivative (line~\ref{DISPL-Test5})
+Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by a temporal derivative
 \begin{equation}\label{eq:detrending}
@@
 \end{equation}
-and we similarly define $\overline{\partial d}_k(t)$ and $\partial g(t)$.
+and we similarly define $\partial d_k(t)$ and $\partial g(t)$.
 
@@
 			\lnl{DISPL-Test11} \If{$\partial g(t) < -\gamma_g \sigma_{g} \vee \partial l(t) > \gamma_g \sigma_{g} $}{
-					\lnl{DISPL-Test9} tampering (probably blurring) has occurred in $z_t$ \\
+					\lnl{DISPL-Test9} raise an alert: tampering (probably blurring) in $z_t$ \\
 									}
@@
 		\lnl{DISPL-Test8} \If{$n_l\geq \Gamma_l$}{
-			\lnl{DISPL-Test9} tampering (probably displacement or occlusion) has occurred in $z_t$ \\
+			\lnl{DISPL-Test9} raise an alert: tampering (probably displacement or occlusion) in $z_t$ \\
 		}
@@
 \end{equation}
-where $\overline{\partial l}_k(t)$ denotes the mean and $\sigma_{l_k}$ the standard deviation of $\partial l$ over $R_k$ (lines \ref{DISPL-Tr1} and \ref{DISPL-Tr7}), computed from tampering-free frames provided for training (i.e.
+where $\overline{\partial l}_k(t)$ denotes the mean and $\sigma_{l_k}$ the standard deviation of $\partial l$ over $R_k$ (Algorithm~\ref{alg:DISPL}, line \ref{DISPL-Tr1}), computed from tampering-free frames provided for training (i.e.
 $z_t\,, t = 1, \dots, T_o$) and $\gamma_l >0$ is a tuning parameter.
-A similar interval is built for  $\overline{\partial d}_k(t)$ and $\partial g$ (line~\ref{DISPL-Tr7}).
+Similar intervals are built for  $\partial d_k(t)$ and $\partial g$ (line~\ref{DISPL-Tr7}).
 
-During operations, any indicator falling outside its confidence region is considered an outlier (line~\ref{DISPL-Test6}): any outlier in $\partial g$ yields a blurring alert (line~\ref{DISPL-Test11}).
-Camera displacements alerts are raised when at least $\Gamma_l$ indicators $\partial l_k$  simultaneously yield outliers (line~\ref{DISPL-Test8}).
+During operations, indicators are computed (lines~\ref{DISPL-Test10} and~\ref{DISPL-Test5}), and any indicator falling outside its confidence region is considered an outlier.
+In particular, any outlier in $\partial g$ yields a blurring alert (line~\ref{DISPL-Test11}), while camera-displacement alerts are raised when at least $\Gamma_l$ indicators $\partial l_k$  simultaneously yield outliers (line~\ref{DISPL-Test8}).
 The threshold $\Gamma_l$ together with $\gamma_l$ determine the displacement-detection promptness.
-The extreme cases are $\Gamma_l = 1$, where a displacement alert raises at any region firing an outlier, and $\Gamma_l = K-1$ where all but one region have to simultaneously fire an outlier (it is worth excluding a region since, for instance the sky-region typically does not change when the camera is displaced, either horizontally or vertically).
+The extreme cases are $\Gamma_l = 1$, where it is enough that a single region fires an outlier to raise a camera-displacement alert, and $\Gamma_l = K-1$ where all but one region\footnote{It is better to exclude a region since, for instance, the sky-region typically does not change when the camera is displaced, either horizontally or vertically} have to simultaneously fire an outlier to raise the alert.
 
-It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, a single tampered frame yield two outliers (as shown in Figure~\ref{fig:indicatori}) and only the first and the last of few consecutive tampered frames yield an outlier in the detrended indicators.
+It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, a single tampered frame yields two outliers (as shown in Figure~\ref{fig:indicatori}) and only the first and the last of a sequence of consecutive tampered frames yield outliers in the detrended indicators.
 This is the reason why detrended indicators are monitored in a one-shot manner, targeting the detection of the first tampered-frame.
-On the one hand, this one-shot solution can provide prompt detections, and this is particularly important at the low frame-rates we consider.
-On the other hand, one-shot monitoring disregards the fact that tampering might persist over several frames.
-To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, and possibly combined with the proposed one-shot monitoring.
+On the one hand, this one-shot solution can provide prompt detections, which is particularly important at the low frame-rates we consider, on the other hand, the persistence of tampering is disregarded.
+To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, possibly combined with the proposed one-shot monitoring.
 %
@@
 \subsection{Scene Segmentation}\label{subsec:Segmentation}
-Scene has to be preliminarily segmented to define regions that our algorithm takes as input (see Algorithm \ref{alg:DISPL}).
-To this purpose, we use part of the training frames before $T_0$ to compute a feature vector $\textbf{f}(x)\in \mathbb{R}^5$ for each pixel $x\in\mathcal{X}$
+Scene has to be preliminarily segmented to define regions that Algorithm \ref{alg:DISPL} takes as input.
+To this purpose, we use part of the training frames before $T_0$ to compute the feature vector $\textbf{f}(x)\in \mathbb{R}^5$ for each pixel $x\in\mathcal{X}$
 \begin{equation}
@@
 In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and column of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the mean and standard deviation of the intensity at $x$, computed over time, and $\overline{g}(x)$ and $\sigma_g(x)$ the mean of gradient norm and its standard deviation, computed overtime.
-These feature vectors are meant to cluster pixels in regions having similar spatial appearance and temporal behavior over training frames.
+These feature vectors are meant to cluster pixels in regions having, over training frames, similar spatial appearance and temporal behavior.
 
 As in superpixel methods \cite{Susstrunk2012}, segmentation is performed by k-means clustering.
-In this case feature vectors~\eqref{eq:featureVector} over the whole image are clustered by a weighted k-means~\cite{kottke1994motion} that scales the Euclidean distance between a feature vector and a cluster centroid of a weight that is inversely proportional to the standard deviation over the cluster.
-This scaling compensates the fact that different components of the feature vector might span very different ranges.
+Feature vectors~\eqref{eq:featureVector} over the whole image are clustered by a weighted k-means~\cite{kottke1994motion} that scales each component of the Euclidean distance between a feature vector and a cluster centroid of a weight that is inversely proportional to the standard deviation over the cluster.
+This scaling compensates the fact that feature vector components might span very different ranges.
+The number of clusters is defined by testing several values and then choosing the best solution according to the Calinski-Harabasz criterion~\cite{calinski1974dendrite}.
 
@@
 %The number of clusters is defined by the Calinski-Harabasz criterion~\cite{calinski1974dendrite}, which initializes the weighted k-means using different number of clusters, and then chooses the best one according a cost function\gi{Adriano controllare come lo chiama}.
-The number of clusters is defined by initializing the weighted k-means using different number of clusters and then choosing the best solution according to the Calinski-Harabasz criterion~\cite{calinski1974dendrite}.
 
@@
 
-\gi{ADD REFERENCE TO LOW-POWER, MEMORIA PER FARE FD}
-The most computationally demanding operations of Algorithm \ref{alg:DISPL} consists in computing of the indicators $g(t)$.
-Assuming that derivatives in~\eqref{eq:normaGradiente} are computed by the Sobel filters, this requires $34$ operations per pixels\footnote{Execution times can be however  decreased whether hardware implementations of the FFT transform are available}.
-Computing $l(t)$ requires 1 operation per pixel and the outlier detection over the indicators require two comparisons per frame.
-Thus, monitoring regions has a negligible impact over the overall computational complexity.
-Segmentation has not been here considered because it is executed only during the initial training phase, and can be performed on an external device connected to the smart camera during configuration.
+The most computationally demanding operations of Algorithm \ref{alg:DISPL} consists in computing of the indicators, in particular $g$.
+Assuming that $f_v$ and $f_h$ in~\eqref{eq:normaGradiente} are Sobel filters, this requires $34$ operations per pixels\footnote{Execution can be accelerated whether hardware implementations of the FFT transform are available}.
+Computing $l$ and $d$ requires 1 and 3 operations per pixel, respectively, while detecting outliers in the indicators requires two comparisons per frame.
+Thus, monitoring regions instead of the whole image has a negligible impact over the overall computational complexity.
+Algorithm~\ref{alg:DISPL} has a very low memory requirement: to compute $d$ it is necessary to store a single frame, while other indicators can be computed online.
+Segmentation is executed only during the initial configuration and can be performed on an external device connected to the smart camera.
+As such, Algorithm~\ref{alg:DISPL} can be properly executed on low-power and ultra low-power cameras.
 

### 2015-06-20 paper.tex
edits inizio Section 4

@@
 \section{Experiments}\label{sec:experiments}
-Experiments are meant to show the effectiveness of the proposed algorithm on real-world sequences, and to specifically assess the advantage of using segmentation for detecting camera displacements. 
-\vspace{-0.3cm}
-\subsubsection{Considered Methods}
-As far as camera-displacement is concerned, Algorithm \eqref{alg:DISPL} is compared against three variants: 1) monitoring the luma on the whole image (\emph{Luma-Total}), 2) monitoring the total energy of the difference between two adjacent frames (\emph{Frame Difference}), and 3) monitoring the luma on $K$ regions obtained by clustering $[r(x), c(x)]$, which yields a Voronoi partitioning of the scene (\emph{Luma-Voronoi}).
-Frame difference is considered as a baseline technique for displacement detection, while comparison against \emph{Luma-Voronoi} shows the advantages of setting regions depending of the image content.
-\\
-In the experiments on blurring detection, we compare Algorithm \ref{alg:DISPL} (where $\partial g$ is monitored on the whole image), against monitoring $\partial g$ separately in image regions, as done for $\partial l_k$.
-\vspace{-0.3cm}
+Experiments are meant to assess the advantages of using a segmentation in a tampering-detection framework like Algorithm~\eqref{alg:DISPL}.
+In particular, to assess the importance of partitioning the scene we compare Algorithm~\eqref{alg:DISPL} against a solution monitoring the indicators on the whole image (\emph{Total} in Figure~\ref{fig:ROC}).
+To show that regions have to be defined depending on the image content, we consider also Algorithm~\eqref{alg:DISPL} when the $K$ regions are obtained by clustering $[r(x), c(x)]$, which yields a Voronoi partitioning of the scene (\emph{Voronoi} in Figure~\ref{fig:ROC}).
+For both regions defined as in Section~\ref{subsec:Segmentation} and Voronoi regions, we test the two extreme cases where $\Gamma_l = 1$ and $\Gamma_l = K-1$ ($\Gamma_d = 1$ and $\Gamma_d = K-1$).
+
 \subsubsection{Datasets}
-The considered dataset contains  sequences of  frames taken from webcams monitoring urban areas (e.g., see Figure \ref{fig:sequences}(a)), in which we've taken the central crop of each image (about $3/4$ of the original image).
+We recorded $8$ sequences from webcams monitoring different urban areas and we synthetically introduced $5\%$ of tampered frames in each.
+In each sequence we have cropped the central region, removing 50 rows from the top and bottom, and 50 columns from the left and right of the image.
+Blurring was obtained by convolution against a Gaussian kernels (as in Figure \ref{fig:sequences}(a)), while displacements were simulated by moving the central cropping regions of a random shift between 20 and 50 pixels in each direction (see Figure \ref{fig:sequences}(b)).
+Overall, the dataset contains $24410$ frames, in which we've inserted $5\%$ of displacement events ($1224$ frames) and $5\%$ of blurring events ($1224$ frames).\footnote{Sequences can be released upon request.}
+\gi{image sizes? blur kernel? quanti pixel il displacemnet?}
 %The second dataset was recorded using a Raspberry Pi Model B+ \cite{raspberry} mounting a camera module.
-In order to have a wide and sound dataset, the tampering events have been made synthetically, blurring the frames by gaussian kernels (see, Figure \ref{fig:sequences}(b)) or moving the crop of the frames in order to simulate displacements (see Figure \ref{fig:sequences}(c)).
+
 %Few sequences of the first dataset contain real tampering events  (Figure \ref{fig:indicatori}). 
 %In the second dataset, tampering was introduced by moving the device or spraying water on the camera lens. 
-At last we have considered $8$ different scenarios, for a total of $24410$ frames considered, in which we've inserted $5\%$ of displacement events ($1224$ frames) and $5\%$ of blurring events ($1224$ frames).\footnote{Sequences can be released upon request}
 \vspace{-0.3cm}

### 2015-06-21 paper.tex
ripassata sect 4

@@
 
-The most computationally demanding operations of Algorithm \ref{alg:DISPL} consists in computing of the indicators, in particular $g$.
-Assuming that $f_v$ and $f_h$ in~\eqref{eq:normaGradiente} are Sobel filters, this requires $34$ operations per pixels\footnote{Execution can be accelerated whether hardware implementations of the FFT transform are available}.
-Computing $l$ and $d$ requires 1 and 3 operations per pixel, respectively, while detecting outliers in the indicators requires two comparisons per frame.
-Thus, monitoring regions instead of the whole image has a negligible impact over the overall computational complexity.
-Algorithm~\ref{alg:DISPL} has a very low memory requirement: to compute $d$ it is necessary to store a single frame, while other indicators can be computed online.
+The most computationally demanding operations of Algorithm \ref{alg:DISPL} consist in computing of the indicators, in particular $g$.
+Computing $g$ requires, when using the Sobel filters in~\eqref{eq:normaGradiente}, $34$ operations per pixels\footnote{Execution can be accelerated whether hardware implementations of the FFT transform are available}, while computing $l$ and $d$ requires 1 and 3 operations per pixel, respectively.
+Detecting outliers in the indicators requires two comparisons per frame: thus, monitoring regions instead of the whole image has a negligible impact on the overall computational complexity.
+Algorithm~\ref{alg:DISPL} has very low memory requirements: $l$ and $g$ indicators can be computed online, while $d$ requires to store a single frame in memory.
 Segmentation is executed only during the initial configuration and can be performed on an external device connected to the smart camera.
@@
 Experiments are meant to assess the advantages of using a segmentation in a tampering-detection framework like Algorithm~\eqref{alg:DISPL}.
-In particular, to assess the importance of partitioning the scene we compare Algorithm~\eqref{alg:DISPL} against a solution monitoring the indicators on the whole image (\emph{Total} in Figure~\ref{fig:ROC}).
-To show that regions have to be defined depending on the image content, we consider also Algorithm~\eqref{alg:DISPL} when the $K$ regions are obtained by clustering $[r(x), c(x)]$, which yields a Voronoi partitioning of the scene (\emph{Voronoi} in Figure~\ref{fig:ROC}).
-For both regions defined as in Section~\ref{subsec:Segmentation} and Voronoi regions, we test the two extreme cases where $\Gamma_l = 1$ and $\Gamma_l = K-1$ ($\Gamma_d = 1$ and $\Gamma_d = K-1$).
+In particular, to assess the importance of partitioning the scene, we compare Algorithm~\eqref{alg:DISPL} against a solution monitoring the indicators on the whole image (\emph{Total} in Figure~\ref{fig:ROC}).
+To show that regions have to follow the image content, we consider Algorithm~\eqref{alg:DISPL} when the $K$ regions are obtained by clustering $[r(x), c(x)]$, which yields a Voronoi partitioning of the scene (\emph{Voronoi} in Figure~\ref{fig:ROC}).
+For both regions defined as in Section~\ref{subsec:Segmentation} and Voronoi regions, we test the two extreme cases where $\Gamma_l = 1$ and $\Gamma_l = K-1$ (respectively, $\Gamma_d = 1$ and $\Gamma_d = K-1$).
 
@@
 In each sequence we have cropped the central region, removing 50 rows from the top and bottom, and 50 columns from the left and right of the image.
-Blurring was obtained by convolution against a Gaussian kernels (as in Figure \ref{fig:sequences}(a)), while displacements were simulated by moving the central cropping regions of a random shift between 20 and 50 pixels in each direction (see Figure \ref{fig:sequences}(b)).
-Overall, the dataset contains $24410$ frames, in which we've inserted $5\%$ of displacement events ($1224$ frames) and $5\%$ of blurring events ($1224$ frames).\footnote{Sequences can be released upon request.}
-\gi{image sizes? blur kernel? quanti pixel il displacemnet?}
+Blurring was obtained by convolution against a Gaussian kernel (as in Figure \ref{fig:sequences}(a)), while displacements were simulated by moving the central cropping regions of a random shift between 20 and 50 pixels in each direction (see Figure \ref{fig:sequences}(b)).
+Overall, the dataset contains $24410$ frames, in which we inserted $1224$ displacement frames ($5\%$) and $1224$ blurry frames ($5\%$)\footnote{Sequences can be released upon request.}.
+\gi{Adriano: image sizes? blur kernel std? quanti pixel il displacement?}
 %The second dataset was recorded using a Raspberry Pi Model B+ \cite{raspberry} mounting a camera module.
@@
 \subsubsection{Performance Assessment}
-The typical figure of merit for assessing detection performance in one-shot monitoring methods are the Receiver-Operating Characteristic (ROC) curves: in our case, these can be computed by varying $\gamma_l$, $\gamma_d$ and $\gamma_g$ in a range of values.
-In particular, each point in the ROC curve depicts a pair of values (1-SPECIFICITY, RECALL) defined as
+The typical figure of merit for assessing detection performance in one-shot monitoring methods are the Receiver-Operating Characteristic (ROC) curves: in our case, these can be computed by varying $\gamma_l$, $\gamma_d$ and $\gamma_g$ in a range of values \gi{quale?}.
+In particular, each point in the ROC curve report a pair of values (1-SPECIFICITY, RECALL) defined as
 \[1-\text{SPECIFICITY}_\gamma = \frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}, \ \text{ and } \ \text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}\]
-where TP, FP, TN, FN represent the number of true positives (tampered frames correctly detected), of false positives (tampering-free frames detected), of true negatives (tampering-free frames not detected) and of false negatives (missed tampered frames), respectively, for a given value of $\gamma$.
+where TP$_\gamma$ represents the number of true positives (tampered frames correctly detected), FP$_\gamma$ of false positives (tampering-free frames detected), TN$_\gamma$ of true negatives (tampering-free frames not detected) and FN$_\gamma$ of false negatives (missed tampered frames), for a given value of $\gamma$.
 \vspace{-0.3cm}
 \subsubsection{Discussion}
-ROC curves and the Area Under the Curve (AUC) reported in Figure~\ref{fig:ROCdisplacement} and its caption indicate that camera displacements can be effectively detected by all the considered methods, and that Algorithm~\ref{alg:DISPL} achieves the best performance (the \emph{Frame-Difference} curve has not been reported because it was quite below the others).
+ROC curves in Figure~\ref{fig:ROCdisplacement} and the corresponding values of Area Under the Curve (AUC) reported in the caption, indicate that camera displacements can be effectively detected by all the considered methods, and that Algorithm~\ref{alg:DISPL} achieves the best performance when $\Gamma_l = \Gamma_d = 1$.
 This result suggests that separately monitoring the luma over image regions is an effective strategy and that the regions have to be suitably estimated.
-For the visualization sake, these ROC curves have been reported in a limited range of both axis, and in particular we focus on low values of 1-SPECIFICITY, thus of false alarms.
-We observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $0.3\%$ of false alarms, which is half of the whole image solution ($0.7\%$) and less than one third of the Voronoi solution ($1\%$).
+
+\emph{We observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $0.3\%$ of false alarms, which is half of the whole image solution ($0.7\%$) and less than one third of the Voronoi solution ($1\%$).
 Having few false alarms is important in a low-power context, considering the high number of frames that have to be processed.
-Most importantly, false alerts are halved basically at no computational overhead.
-
-In contrast, Figure~\ref{fig:ROCdefocus} confirms that when monitoring $g$ it is more convenient to consider the whole scene at once, as typically the regions $\{R_k\}$ do not include the most prominent edge of the scene, which are indeed the most informative part to detect blurring.
+Most importantly, false alerts are halved basically at no computational overhead.} \gi{Aggiornare con i nuovi dati}
+In contrast, Figure~\ref{fig:ROC}(c) confirms that when monitoring $g$ it is more convenient to consider the whole scene at once, as typically the regions $\{R_k\}$ do not include the most prominent edge of the scene, which are indeed the most informative part to detect blurring.
 This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.
@@
 	In this case, the Gradient-Total (AUC $99.96\%$) outperforms the Gradient-Regions (AUC $98.95\%$) because this latter do not include the most prominent edges in images, which are the most important information to detect blurring.
-This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.}
+This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.
+	\textbf{TODO} For the visualization sake, these ROC curves have been reported in a limited range of both axis, and in particular we focus on low values of 1-SPECIFICITY, thus of false alarms.
+	}
 \label{fig:ROC}

### 2015-06-22 paper.tex
Section 4 e Conclusions

@@
 \section{Experiments}\label{sec:experiments}
-Experiments are meant to assess the advantages of using a segmentation in a tampering-detection framework like Algorithm~\eqref{alg:DISPL}.
-In particular, to assess the importance of partitioning the scene, we compare Algorithm~\eqref{alg:DISPL} against a solution monitoring the indicators on the whole image (\emph{Total} in Figure~\ref{fig:ROC}).
-To show that regions have to follow the image content, we consider Algorithm~\eqref{alg:DISPL} when the $K$ regions are obtained by clustering $[r(x), c(x)]$, which yields a Voronoi partitioning of the scene (\emph{Voronoi} in Figure~\ref{fig:ROC}).
+Experiments are meant to assess the advantages of separately monitoring regions in a tampering-detection framework like Algorithm~\ref{alg:DISPL}.
+To this purpose, we show that Algorithm~\ref{alg:DISPL} outperforms a solution monitoring the indicators on the whole image (\emph{Total} in Figure~\ref{fig:ROC}).
+We also show that image regions have to be adaptively defined on the image content, and that operating on a Voronoi partition of the scene in $K$ regions obtained by clustering $[r(x), c(x)]$ leads to performance loss (\emph{Voronoi} in Figure~\ref{fig:ROC}).
 For both regions defined as in Section~\ref{subsec:Segmentation} and Voronoi regions, we test the two extreme cases where $\Gamma_l = 1$ and $\Gamma_l = K-1$ (respectively, $\Gamma_d = 1$ and $\Gamma_d = K-1$).
 
-\subsubsection{Datasets}
-We recorded $8$ sequences from webcams monitoring different urban areas and we synthetically introduced $5\%$ of tampered frames in each.
-In each sequence we have cropped the central region, removing 50 rows from the top and bottom, and 50 columns from the left and right of the image.
-Blurring was obtained by convolution against a Gaussian kernel (as in Figure \ref{fig:sequences}(a)), while displacements were simulated by moving the central cropping regions of a random shift between 20 and 50 pixels in each direction (see Figure \ref{fig:sequences}(b)).
-Overall, the dataset contains $24410$ frames, in which we inserted $1224$ displacement frames ($5\%$) and $1224$ blurry frames ($5\%$)\footnote{Sequences can be released upon request.}.
+We recorded $8$ sequences from webcams monitoring different urban areas, yielding overall $24410$ frames.
+From each frame we have cropped the central area, removing the 50 top-most and bottom-most rows, and the 50 left-most and right-most columns.
+We have synthetically introduced $5\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and coulumns (as in Figure \ref{fig:sequences}(b)).
+Blurring was simulated by convolution against a Gaussian kernel (as in Figure \ref{fig:sequences}(a)).
+\footnote{Sequences can be released upon request.}.
 \gi{Adriano: image sizes? blur kernel std? quanti pixel il displacement?}
+
 %The second dataset was recorded using a Raspberry Pi Model B+ \cite{raspberry} mounting a camera module. 
-
 %Few sequences of the first dataset contain real tampering events  (Figure \ref{fig:indicatori}). 
 %In the second dataset, tampering was introduced by moving the device or spraying water on the camera lens.
-\vspace{-0.3cm}
-\subsubsection{Performance Assessment}
+
 The typical figure of merit for assessing detection performance in one-shot monitoring methods are the Receiver-Operating Characteristic (ROC) curves: in our case, these can be computed by varying $\gamma_l$, $\gamma_d$ and $\gamma_g$ in a range of values \gi{quale?}.
-In particular, each point in the ROC curve report a pair of values (1-SPECIFICITY, RECALL) defined as
+In particular, each point in the ROC curve report a pair of values (1-SPECIFICITY$_\gamma$, RECALL$_\gamma$) defined as
 \[1-\text{SPECIFICITY}_\gamma = \frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}, \ \text{ and } \ \text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}\]
@@
 \subsubsection{Discussion}
-ROC curves in Figure~\ref{fig:ROCdisplacement} and the corresponding values of Area Under the Curve (AUC) reported in the caption, indicate that camera displacements can be effectively detected by all the considered methods, and that Algorithm~\ref{alg:DISPL} achieves the best performance when $\Gamma_l = \Gamma_d = 1$.
-This result suggests that separately monitoring the luma over image regions is an effective strategy and that the regions have to be suitably estimated.
+ROC curves in Figure~\ref{fig:ROC}(a)-(b) and the Area Under the Curve (AUC) values reported in the caption indicate that camera displacements can be effectively detected by all the considered methods; the comparison against Voronoi regions suggests that a more substantial advantage is guaranteed when regions fit the scene content.
+As far as the values of $\Gamma_l$ and $\Gamma_d $ are concerned, the best option is to raise an alert as soon as a single region provides a very clear outlier (i.e., interval \eqref{eq:confidenceRegions} are quite wide).
+This option is preferable with respect to requiring $K-1$ regions to simultaneously fire an outlier, because it also enable the detection of partial occlusions of the scene.
+Overall, the ROC curves confirm that $d$ is a far more effective indicator than $l$ in detecting camera displacements.
 
@@
 Most importantly, false alerts are halved basically at no computational overhead.} \gi{Aggiornare con i nuovi dati}
-In contrast, Figure~\ref{fig:ROC}(c) confirms that when monitoring $g$ it is more convenient to consider the whole scene at once, as typically the regions $\{R_k\}$ do not include the most prominent edge of the scene, which are indeed the most informative part to detect blurring.
-This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.
+
+In contrast, Figure~\ref{fig:ROC}(c) confirms that blurring are more effectively detected by monitoring the whole scene at once.
+This is probably due to the fact that typically, regions $\{R_k\}$ do not include the most prominent edge of the scene, which are indeed the most informative part to detect blurring.
+This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.\gi{aggiungere commento su K-1 vornoi che va meglio?}
 
@@
 \vspace{-0.1cm}
-We presented a tampering detection algorithm that leverages a preliminary image partitioning to improve the detection of camera displacements, and that at the same time can detect blurring.
-The algorithm has low-computational complexity and is suited for low-power cameras operating at low frame rates.
+We presented a tampering-detection algorithm that leverages an image segmentation to improve the detection of camera displacements, and that at the same time can detect blurring.
+The algorithm has low-computational complexity and has to be considered a prompt trigger for detecting tampering in low-power and ultra low-power cameras operating at low frame rates.
 
-Ongoing work concern the extension of our solution to other types of tampering, like degradations of the imaging sensor, as well as the integration of this one-shot technique with sequential monitoring scheme such as~\cite{alippi2010detecting}, to leverage the fact that tampering might last more than one frame in the detection.
-Moreover, we will investigate different segmentation strategies and the application of superpixels methods~\cite{Susstrunk2012} to video sequences by including temporal features in the clustered vectors.
+Ongoing work concerns investigations towards other types of tampering, like degradations of the imaging sensor, as well as the integration of sequential monitoring schemes~\cite{alippi2010detecting} to detect subtle tampering that persists over time.
+Moreover, we will investigate the use of superpixels methods~\cite{Susstrunk2012} to perform the preliminary scene segmentation, by including suitable temporal information to the feature vectors.
 

### 2015-06-22 paper.tex
SIstemato didascalie e immagini (tranne fig 3 che è in fase di elaborazione)

@@
 \caption{Example of camera displacement. 
-Left plots report the luma values, while right plots depict the detrended luma: we show both indicators computed over the full image (\textbf{a}) and on two reported regions (\textbf{b} and \textbf{c}).
-Red stars indicate a single displaced frame, which occurs at $t =157$.
-Displacement yields a peak in $l$ and two outliers in the sequence of detrended indicators $\partial l$, which can be clearly detected.
-}
+Left plots report the luma values, while right plots depict the detrended luma: 
+we show both indicators computed over the full image (\textbf{a}) and on two reported regions (\textbf{b} and \textbf{c}).
+Red dots indicate a single displaced frame.
+Values in the highlighted areas refer to the frames depicted on top of the fugure.
+Displacement yields a peak in $l$ and two outliers in the sequence of detrended indicators $\partial l$, which can be clearly detected. }
 \label{fig:indicatori} 
@@
 From each frame we have cropped the central area, removing the 50 top-most and bottom-most rows, and the 50 left-most and right-most columns.
-We have synthetically introduced $5\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and coulumns (as in Figure \ref{fig:sequences}(b)).
+We have synthetically introduced $10\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and coulumns (as in Figure \ref{fig:sequences}(b)).
 Blurring was simulated by convolution against a Gaussian kernel (as in Figure \ref{fig:sequences}(a)).
 \footnote{Sequences can be released upon request.}.
-\gi{Adriano: image sizes? blur kernel std? quanti pixel il displacement?}
+%\gi{Adriano: image sizes? blur kernel std? quanti pixel il displacement?}
 
@@
 
-The typical figure of merit for assessing detection performance in one-shot monitoring methods are the Receiver-Operating Characteristic (ROC) curves: in our case, these can be computed by varying $\gamma_l$, $\gamma_d$ and $\gamma_g$ in a range of values \gi{quale?}.
+The typical figure of merit for assessing detection performance in one-shot monitoring methods are the Receiver-Operating Characteristic (ROC) curves: in our case, these can be computed by varying $\gamma_l$, $\gamma_d$ and $\gamma_g$ in a range of values between $0.1$ and $50$, with step $0.1$.
 In particular, each point in the ROC curve report a pair of values (1-SPECIFICITY$_\gamma$, RECALL$_\gamma$) defined as
@@
 
-\emph{We observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $0.3\%$ of false alarms, which is half of the whole image solution ($0.7\%$) and less than one third of the Voronoi solution ($1\%$).
+\emph{Considering $d$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $0.4\%$ of false alarms, with $\gamma = 24.7$, which is less than one fifth of the whole image solution ($2.2\%$ with $\gamma = 4.6$) and the Voronoi solution ($2.3\%$ with $\gamma = 12.9$).
+On the other side, considering $l$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $13\%$ of false alarms, with $\gamma = 2.5$, which is about one sixth of the whole image solution ($75\%$ with $\gamma = 0.1$) and about three fifty than the Voronoi solution ($22\%$ with $\gamma = 1.7 $).
 Having few false alarms is important in a low-power context, considering the high number of frames that have to be processed.
@@
 \includegraphics[width=1\linewidth]{Immagini/curve}
-\caption{\textbf{(a)} ROC curve for displacement detection.
+\caption{\textbf{(a)} ROC curve for displacement detection considering Frame Difference (FD) as indicator.
 	%There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
-	The AUC corresponding to Algorithm~\ref{alg:DISPL} (Luma-Region) is $99.94\%$, the AUC of Luma-Total is $99.89\%$ the AUC of Luma-Voronoi is $99.85\%$, and the AUC of Frame-Difference is $99.44\%$ (axis have been zoomed for the visualization sake).
+	The AUC corresponding to Algorithm~\ref{alg:DISPL} ($1$ Region - Algorithm $1$) is $99.89\%$, while the AUC of considering the whole image (Total) is $99.65\%$. 
+	%the AUC of Luma-Voronoi is $99.85\%$, and the AUC of Frame-Difference is $99.44\%$ (axis have been zoomed for the visualization sake).
 Large AUC values indicate that these displacements can be successfully detected and that Algorithm~\ref{alg:DISPL} is the most effective one.
-	\textbf{(b)} ROC curve for blurring detection.
-	In this case, the Gradient-Total (AUC $99.96\%$) outperforms the Gradient-Regions (AUC $98.95\%$) because this latter do not include the most prominent edges in images, which are the most important information to detect blurring.
+	\textbf{(b)} ROC curve for displacement detection considering Luma as indicator.
+	%There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
+	The AUC corresponding to Algorithm~\ref{alg:DISPL} ($1$ Region - Algorithm $1$) is $98.44\%$, while the AUC of considering the whole image (Total) is $84.07\%$. 
+	\textbf{(c)} ROC curve for blurring detection.
+	In this case, the Gradient-Total (AUC $99.13\%$) outperforms the Gradient-Regions (AUC $98.4\%$) because this latter do not include the most prominent edges in images, which are the most important information to detect blurring.
 This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.
-	\textbf{TODO} For the visualization sake, these ROC curves have been reported in a limited range of both axis, and in particular we focus on low values of 1-SPECIFICITY, thus of false alarms.
+	For the visualization sake, ROC curves \textbf{(a)} and \textbf{(c)} have been reported in a limited range of both axis, and in particular we focus on low values of 1-SPECIFICITY, thus of false alarms.
 	}

### 2015-06-22 paper.tex
Finito capitolo 4

@@
 As such, Algorithm~\ref{alg:DISPL} can be properly executed on low-power and ultra low-power cameras.
-
 %\gi{Adriano: inserisci qui l'algoritmo e traducilo in inglese. Se riusciamo lo spostiamo prima di tutte le sottosezioni}
@@
 We have synthetically introduced $10\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and coulumns (as in Figure \ref{fig:sequences}(b)).
-Blurring was simulated by convolution against a Gaussian kernel (as in Figure \ref{fig:sequences}(a)).
+Blurring  (as in Figure \ref{fig:sequences}(a)) was simulated by convolution against a Gaussian kernel, with $1\leq\sigma\leq5$ randomly taken.
 \footnote{Sequences can be released upon request.}.
 %\gi{Adriano: image sizes? blur kernel std? quanti pixel il displacement?}
-
 %The second dataset was recorded using a Raspberry Pi Model B+ \cite{raspberry} mounting a camera module. 
@@
 This option is preferable with respect to requiring $K-1$ regions to simultaneously fire an outlier, because it also enable the detection of partial occlusions of the scene.
-Overall, the ROC curves confirm that $d$ is a far more effective indicator than $l$ in detecting camera displacements.
-
-\emph{Considering $d$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $0.4\%$ of false alarms, with $\gamma = 24.7$, which is less than one fifth of the whole image solution ($2.2\%$ with $\gamma = 4.6$) and the Voronoi solution ($2.3\%$ with $\gamma = 12.9$).
-On the other side, considering $l$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $13\%$ of false alarms, with $\gamma = 2.5$, which is about one sixth of the whole image solution ($75\%$ with $\gamma = 0.1$) and about three fifty than the Voronoi solution ($22\%$ with $\gamma = 1.7 $).
+Overall, the ROC curves confirm that $d$ is a far more effective indicator than $l$ in detecting camera displacements.\\
+Considering $d$ as indicator, we observe that to limit FPs to $1\%$, Algorithm~\ref{alg:DISPL} have a recall of $99.92\%$, with $\gamma = 15.6$, while considering the whole image solution decreases TPs to $91.67\%$ ($\gamma = 6.5$), as for considering Voronoi's regions ($89.05\%$, $\gamma = 18.64$).
+Considering $l$ as indicator, instead, limit FPs to $1\%$ brings a recall of $73.98\%$ when using Algorithm~\ref{alg:DISPL} ($\gamma=5.8$), more than considering the whole image solution ($17.85\%$, $\gamma=3.7$), and Voronoi's regions ($53.02\%$, $\gamma=6$).
 Having few false alarms is important in a low-power context, considering the high number of frames that have to be processed.
+Most importantly, false alerts are halved basically at no computational overhead.\\ %\gi{Aggiornare con i nuovi dati}
+%\emph{Considering $d$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $0.4\%$ of false alarms, with $\gamma = 24.7$, which is less than one fifth of the whole image solution ($2.2\%$ with $\gamma = 4.6$) and the Voronoi solution ($2.3\%$ with $\gamma = 12.9$).
+%On the other side, considering $l$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $13\%$ of false alarms, with $\gamma = 2.5$, which is about one sixth of the whole image solution ($75\%$ with $\gamma = 0.1$) and about three fifty than the Voronoi solution ($22\%$ with $\gamma = 1.7 $).
+%Having few false alarms is important in a low-power context, considering the high number of frames that have to be processed.
 Most importantly, false alerts are halved basically at no computational overhead.} \gi{Aggiornare con i nuovi dati}
-
-In contrast, Figure~\ref{fig:ROC}(c) confirms that blurring are more effectively detected by monitoring the whole scene at once.
+In contrast, Figure~\ref{fig:ROC}(c) confirms that blurring \ad{Giacomo: blurs?} are more effectively detected by monitoring the whole scene at once.
 This is probably due to the fact that typically, regions $\{R_k\}$ do not include the most prominent edge of the scene, which are indeed the most informative part to detect blurring.
-This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.\gi{aggiungere commento su K-1 vornoi che va meglio?}
+This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.%\gi{aggiungere commento su K-1 vornoi che va meglio?}
 

### 2015-06-22 paper.tex
Sistemata didascalia

@@
 We address the detection of camera blurring and displacement (Section \ref{sec:probForm}), and propose an algorithm (Section~\ref{sec:propSol}) that relies on indicators that can be easily computed in low-power smart cameras.
-We monitor the average image intensity or frame difference (to detect camera displacement), and the average gradient norm (to detect blurring), and consider any frame yielding an outlier as tampered.
+We monitor the average image intensity or frame difference (FD) (to detect camera displacement), and the average gradient norm (to detect blurring), and consider any frame yielding an outlier as tampered.
 In particular, we show that separately monitoring these indicators over different regions of the image can substantially improve the displacement-detection performance.
@@
 Most importantly, false alerts are halved basically at no computational overhead.} \gi{Aggiornare con i nuovi dati}
-In contrast, Figure~\ref{fig:ROC}(c) confirms that blurring \ad{Giacomo: blurs?} are more effectively detected by monitoring the whole scene at once.
+In contrast, Figure~\ref{fig:ROC}(c) confirms that blurring is more effectively detected by monitoring the whole scene at once.
 This is probably due to the fact that typically, regions $\{R_k\}$ do not include the most prominent edge of the scene, which are indeed the most informative part to detect blurring.
@@
 \includegraphics[width=1\linewidth]{Immagini/curve}
-\caption{\textbf{(a)} ROC curve for displacement detection considering Frame Difference (FD) as indicator.
-	%There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
-	The AUC corresponding to Algorithm~\ref{alg:DISPL} ($1$ Region - Algorithm $1$) is $99.89\%$, while the AUC of considering the whole image (Total) is $99.65\%$. 
-	%the AUC of Luma-Voronoi is $99.85\%$, and the AUC of Frame-Difference is $99.44\%$ (axis have been zoomed for the visualization sake).
+\caption{\textbf{(a)} ROC curves relative to displacement detection based on FD: the AUC obtained by 
+"Algo 1 on one region" is $99.89\%$, while the AUC obtained by "Total" (on the whole 
+image) is $99.65\%$.
+\textbf{(b)} ROC curves relative to displacement detection based on Luma: the AUC obtained by 
+"Algo 1 on one region" is $98.44\%$, while the AUC obtained by "Total" (on the whole 
+image) is $84.07\%$.
+\textbf{(c)} ROC curves relative to blurring detection based on Gradient: as opposed to \textbf{(a)} and 
+\textbf{(b)}, "Algo 1 on one region" (AUC $98.4\%$) is outperformed by "Total" (AUC $99.13\%$).
+Note that, in order to highlight the details, the ROC curves are drawn with different 
+scales.}
+%\caption{\textbf{(a)} ROC curve for displacement detection considering Frame Difference (FD) as indicator.
+%	%There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
+%	The AUC corresponding to Algorithm~\ref{alg:DISPL} ($1$ region - Algorithm $1$) is $99.89\%$, while the AUC of considering the whole image (Total) is $99.65\%$. 
+%	%the AUC of Luma-Voronoi is $99.85\%$, and the AUC of Frame-Difference is $99.44\%$ (axis have been zoomed for the visualization sake).
 Large AUC values indicate that these displacements can be successfully detected and that Algorithm~\ref{alg:DISPL} is the most effective one.
-	\textbf{(b)} ROC curve for displacement detection considering Luma as indicator.
-	%There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
-	The AUC corresponding to Algorithm~\ref{alg:DISPL} ($1$ Region - Algorithm $1$) is $98.44\%$, while the AUC of considering the whole image (Total) is $84.07\%$. 
-	\textbf{(c)} ROC curve for blurring detection.
-	In this case, the Gradient-Total (AUC $99.13\%$) outperforms the Gradient-Regions (AUC $98.4\%$) because this latter do not include the most prominent edges in images, which are the most important information to detect blurring.
-This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.
-	For the visualization sake, ROC curves \textbf{(a)} and \textbf{(c)} have been reported in a limited range of both axis, and in particular we focus on low values of 1-SPECIFICITY, thus of false alarms.
-	}
+%	\textbf{(b)} ROC curve for displacement detection considering Luma as indicator.
+%	%There is an improvement in the performance considering adaptive regions w.r.t. considering the whole image or Voronoi regions.
+%	The AUC corresponding to Algorithm~\ref{alg:DISPL} ($1$ region - Algorithm $1$) is $98.44\%$, while the AUC of considering the whole image (Total) is $84.07\%$. 
+%	\textbf{(c)} ROC curve for blurring detection.
+%	In this case, the Gradient-Total (AUC $99.13\%$) outperforms the Gradient-Regions (AUC $98.4\%$) because this latter do not include the most prominent edges in images, which are the most important information to detect blurring.
+%This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.
+%	ROC curves \textbf{(a)} and \textbf{(c)}
+%	%For the visualization sake, ROC curves \textbf{(a)} and \textbf{(c)} have been reported in a limited range of both axis, and in particular we focus on low values of 1-SPECIFICITY, thus of false alarms.
+	%}
 \label{fig:ROC}

### 2015-06-22 paper.tex
Sistemati caratteri ROC e sistemati apici caption

@@
 \caption{\textbf{(a)} ROC curves relative to displacement detection based on FD: the AUC obtained by 
-"Algo 1 on one region" is $99.89\%$, while the AUC obtained by "Total" (on the whole 
+``One region - Algorithm 1'' is $99.89\%$, while the AUC obtained by ``Total'' (on the whole 
 image) is $99.65\%$.
 \textbf{(b)} ROC curves relative to displacement detection based on Luma: the AUC obtained by 
-"Algo 1 on one region" is $98.44\%$, while the AUC obtained by "Total" (on the whole 
+``One region - Algorithm 1'' is $98.44\%$, while the AUC obtained by ``Total'' (on the whole 
 image) is $84.07\%$.
 \textbf{(c)} ROC curves relative to blurring detection based on Gradient: as opposed to \textbf{(a)} and 
-\textbf{(b)}, "Algo 1 on one region" (AUC $98.4\%$) is outperformed by "Total" (AUC $99.13\%$).
+\textbf{(b)}, ``One region - Algorithm 1'' (AUC $98.4\%$) is outperformed by ``Total'' (AUC $99.13\%$).
 Note that, in order to highlight the details, the ROC curves are drawn with different 

### 2015-06-22 paper.tex
Modificata frase capitolo 4

@@
 As far as the values of $\Gamma_l$ and $\Gamma_d $ are concerned, the best option is to raise an alert as soon as a single region provides a very clear outlier (i.e., interval \eqref{eq:confidenceRegions} are quite wide).
-This option is preferable with respect to requiring $K-1$ regions to simultaneously fire an outlier, because it also enable the detection of partial occlusions of the scene.
-Overall, the ROC curves confirm that $d$ is a far more effective indicator than $l$ in detecting camera displacements.\\
+This option is preferable with respect to requiring $K-1$ regions to simultaneously fire an outlier, because it also enable the detection of partial occlusions of the scene.\\
 Considering $d$ as indicator, we observe that to limit FPs to $1\%$, Algorithm~\ref{alg:DISPL} have a recall of $99.92\%$, with $\gamma = 15.6$, while considering the whole image solution decreases TPs to $91.67\%$ ($\gamma = 6.5$), as for considering Voronoi's regions ($89.05\%$, $\gamma = 18.64$).
@@
 Having few false alarms is important in a low-power context, considering the high number of frames that have to be processed.
-Most importantly, false alerts are halved basically at no computational overhead.\\ %\gi{Aggiornare con i nuovi dati}
+Most importantly, false alerts are halved basically at no computational overhead.
+Finally, ROC curves confirm that $d$ is a far more effective indicator than $l$ in detecting camera displacements.\\ %\gi{Aggiornare con i nuovi dati}
 %\emph{Considering $d$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $0.4\%$ of false alarms, with $\gamma = 24.7$, which is less than one fifth of the whole image solution ($2.2\%$ with $\gamma = 4.6$) and the Voronoi solution ($2.3\%$ with $\gamma = 12.9$).

### 2015-06-22 paper.tex
Cambiati assi ROC

@@
 Cameras operating outdoor and in harsh environments are exposed to atmospheric phenomena and intentional attacks that might prevent the correct image acquisition.
-Rain, snow, dust lying on the camera lens cause blurry (Figure \ref{fig:pioggia}) or partially occluded (Figure \ref{fig:neve}) pictures, while wind might displace the camera (Figure \ref{fig:displ1} -\ref{fig:displ2}).
+Rain, snow, dust lying on the camera lens cause blurry (Figure \ref{fig:pioggia}) or partially occluded (Figure \ref{fig:neve}) pictures, while wind might displace the camera (Figure \ref{fig:displ1} - \ref{fig:displ2}).
 Similarly, an attacker can intentionally change the camera focus, spray some opaque or glossy liquid, displace or occlude the camera.
@@
 around few frames per second) and are connected to the power supply.
-Tampering-detection for surveillance cameras have been quite investigated in the literature; in particular, camera displacements and occlusions are typically detected by comparing the current frame against an estimate of the scene background, while blurring is detected by monitoring the high-frequency components of images, as these are expected to drop.
+Tampering detection for surveillance cameras have been quite investigated in the literature; in particular, camera displacements and occlusions are typically detected by comparing the current frame against an estimate of the scene background, while blurring is detected by monitoring the high-frequency components of images, as these are expected to drop.
 In~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing the histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain.

### 2015-06-22 paper.tex
Section IV, including caption

--- 2015-06-22 paper.tex
@@
 \section{Experiments}\label{sec:experiments}
-Experiments are meant to assess the advantages of separately monitoring regions in a tampering-detection framework like Algorithm~\ref{alg:DISPL}.
-To this purpose, we show that Algorithm~\ref{alg:DISPL} outperforms a solution monitoring the indicators on the whole image (\emph{Total} in Figure~\ref{fig:ROC}).
-We also show that image regions have to be adaptively defined on the image content, and that operating on a Voronoi partition of the scene in $K$ regions obtained by clustering $[r(x), c(x)]$ leads to performance loss (\emph{Voronoi} in Figure~\ref{fig:ROC}).
-For both regions defined as in Section~\ref{subsec:Segmentation} and Voronoi regions, we test the two extreme cases where $\Gamma_l = 1$ and $\Gamma_l = K-1$ (respectively, $\Gamma_d = 1$ and $\Gamma_d = K-1$).
+Experiments are meant to assess the advantages of separately monitoring regions in the considered tampering-detection framework.
+To this purpose, we show that Algorithm~\ref{alg:DISPL} detects camera displacements better than monitoring indicators on the whole image (\emph{Whole Image} in Figure~\ref{fig:ROC}).
+We also show that it is important to define regions considering the scene content, since operating on $K$ regions obtained by clustering $[r(x), c(x)]$ leads to performance loss (\emph{Voronoi} in Figure~\ref{fig:ROC}).
+For both Voronoi and regions defined as in Section~\ref{subsec:Segmentation}, we consider the two extreme configurations where alerts are raised at the first region firing an outlier or when $K-1$ regions simultaneously fire outliers.
 
 We recorded $8$ sequences from webcams monitoring different urban areas, yielding overall $24410$ frames.
-From each frame we have cropped the central area, removing the 50 top-most and bottom-most rows, and the 50 left-most and right-most columns.
-We have synthetically introduced $10\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and coulumns (as in Figure \ref{fig:sequences}(b)).
-Blurring  (as in Figure \ref{fig:sequences}(a)) was simulated by convolution against a Gaussian kernel, with $1\leq\sigma\leq5$ randomly taken.
-\footnote{Sequences can be released upon request.}.
+From each frame we have cropped the central area, removing the 50 top-most and bottom-most rows, and the 50 left-most and right-most columns\footnote{Sequences can be provided upon request.}.
+We have synthetically introduced $5\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and coulumns (as in Figure \ref{fig:sequences}(b)).
+Blurring  (as in Figure \ref{fig:sequences}(a)) was simulated by convolution against a Gaussian kernel, with $1\leq\sigma\leq5$ randomly defined.
 %\gi{Adriano: image sizes? blur kernel std? quanti pixel il displacement?}
@@
 
-The typical figure of merit for assessing detection performance in one-shot monitoring methods are the Receiver-Operating Characteristic (ROC) curves: in our case, these can be computed by varying $\gamma_l$, $\gamma_d$ and $\gamma_g$ in a range of values between $0.1$ and $50$, with step $0.1$.
-In particular, each point in the ROC curve report a pair of values (1-SPECIFICITY$_\gamma$, RECALL$_\gamma$) defined as
-\[1-\text{SPECIFICITY}_\gamma = \frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}, \ \text{ and } \ \text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}\]
-where TP$_\gamma$ represents the number of true positives (tampered frames correctly detected), FP$_\gamma$ of false positives (tampering-free frames detected), TN$_\gamma$ of true negatives (tampering-free frames not detected) and FN$_\gamma$ of false negatives (missed tampered frames), for a given value of $\gamma$.
+A typical figure of merit for assessing detection performance is the Receiver-Operating Characteristic (ROC) curve, where each point corresponds to a pair (FPR$_\gamma$, TPR$_\gamma$) defined as
+%
+\[\text{FPR}_\gamma = \frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}, \ \text{ and } \ \text{TPR}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
+%
+Here, TP$_\gamma$ represents the number of true positives (tampered frames correctly detected), FP$_\gamma$ of false positives (tampering-free frames detected), TN$_\gamma$ of true negatives (tampering-free frames not detected) and FN$_\gamma$ of false negatives (missed tampered frames), for a given value of $\gamma$.
+ROC curves have been computed by varying $\gamma_l$, $\gamma_d$ and $\gamma_g$ in a range of values between $0.1$ and $50$.
 \vspace{-0.3cm}
 \subsubsection{Discussion}
-ROC curves in Figure~\ref{fig:ROC}(a)-(b) and the Area Under the Curve (AUC) values reported in the caption indicate that camera displacements can be effectively detected by all the considered methods; the comparison against Voronoi regions suggests that a more substantial advantage is guaranteed when regions fit the scene content.
-As far as the values of $\Gamma_l$ and $\Gamma_d $ are concerned, the best option is to raise an alert as soon as a single region provides a very clear outlier (i.e., interval \eqref{eq:confidenceRegions} are quite wide).
-This option is preferable with respect to requiring $K-1$ regions to simultaneously fire an outlier, because it also enable the detection of partial occlusions of the scene.\\
-Considering $d$ as indicator, we observe that to limit FPs to $1\%$, Algorithm~\ref{alg:DISPL} have a recall of $99.92\%$, with $\gamma = 15.6$, while considering the whole image solution decreases TPs to $91.67\%$ ($\gamma = 6.5$), as for considering Voronoi's regions ($89.05\%$, $\gamma = 18.64$).
-Considering $l$ as indicator, instead, limit FPs to $1\%$ brings a recall of $73.98\%$ when using Algorithm~\ref{alg:DISPL} ($\gamma=5.8$), more than considering the whole image solution ($17.85\%$, $\gamma=3.7$), and Voronoi's regions ($53.02\%$, $\gamma=6$).
-Having few false alarms is important in a low-power context, considering the high number of frames that have to be processed.
-Most importantly, false alerts are halved basically at no computational overhead.
-Finally, ROC curves confirm that $d$ is a far more effective indicator than $l$ in detecting camera displacements.\\ %\gi{Aggiornare con i nuovi dati}
+ROC curves in Figure~\ref{fig:ROC}(a)-(b) and the Area Under the Curve (AUC) values reported in the caption confirm that camera displacements can be better detected by monitoring FD than luma, and that it is convenient to define regions segmenting the scene content.
+As an example, if we set $\gamma_d = 15.6$, Algorithm~\ref{alg:DISPL} operates at $\text{FPR} = 1\%$ and detects nearly all tampering since $\text{TPR} = 99.92\%$.
+In contrast, to achieve $\text{FPR} = 1\%$ when monitoring the whole image, FPR drops to $91.67\%$ ($\gamma_d = 6.5$), and when considering Voronoi to $\text{TPR} = 89.05\%$ ($\gamma_d = 18.64$).
+When monitoring $l$, Algorithm~\ref{alg:DISPL} achieves TPR~=~$73.98\%$ at $\text{FPR}=1\%$  ($\gamma_d=5.8$), while monitoring the whole image is quite ineffective ($\text{TPR} = 17.85\%$, $\gamma_d=3.7$), and similarly perform Voronoi's regions ($\text{FPR} = 53.02\%$, $\gamma_d=6$).
+We remark that, in the considered low-power scenario, it is important to operate at low FPR, given the large number of frames that smart cameras have to process over time and the cost of data-transmission in terms of battery power.
+As far as the values of $\Gamma_d$ and $\Gamma_l$ are concerned, the most effective configuration consists in raising an alert as soon as a single region fires a very clear outlier, which implies setting quite wide intervals~\eqref{eq:confidenceRegions}.
+This option is also preferable because it enables the detection of partial occlusions of the scene.
+
+%\gi{Aggiornare con i nuovi dati}
 %\emph{Considering $d$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $0.4\%$ of false alarms, with $\gamma = 24.7$, which is less than one fifth of the whole image solution ($2.2\%$ with $\gamma = 4.6$) and the Voronoi solution ($2.3\%$ with $\gamma = 12.9$).
@@
 Most importantly, false alerts are halved basically at no computational overhead.} \gi{Aggiornare con i nuovi dati}
-In contrast, Figure~\ref{fig:ROC}(c) confirms that blurring is more effectively detected by monitoring the whole scene at once.
-This is probably due to the fact that typically, regions $\{R_k\}$ do not include the most prominent edge of the scene, which are indeed the most informative part to detect blurring.
+Figure~\ref{fig:ROC}(c) confirms that blurring is more effectively detected by monitoring the whole image at once.
+This is probably due to the fact that, typically, regions do not include the most prominent edge of the scene, which are indeed the most informative parts to detect blurring.
 This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.%\gi{aggiungere commento su K-1 vornoi che va meglio?}
@@
 \textbf{(b)}, ``One region - Algorithm 1'' (AUC $98.4\%$) is outperformed by ``Total'' (AUC $99.13\%$).
-Note that, in order to highlight the details, the ROC curves are drawn with different 
-scales.}
+To highlight differences, the ROC curves in \textbf{(a)} and \textbf{(c)} are drawn in a different TPR, FPR interval.}
 %\caption{\textbf{(a)} ROC curve for displacement detection considering Frame Difference (FD) as indicator.

### 2015-06-22 paper.tex
piccole modifiche

--- 2015-06-22 paper.tex
@@
 
-The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect the time instant $\tau^*$ when tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.
+The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t, t=1,\dots\}$ to detect the time instant $\tau^*$ when tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.
 We assume that $T_0$ tampering-free frames are provided for training.
@@
 
-The most computationally demanding operations of Algorithm \ref{alg:DISPL} consist in computing of the indicators, in particular $g$.
+The most computationally demanding operations of Algorithm \ref{alg:DISPL} consist in computing the indicators, in particular $g$.
 Computing $g$ requires, when using the Sobel filters in~\eqref{eq:normaGradiente}, $34$ operations per pixels\footnote{Execution can be accelerated whether hardware implementations of the FFT transform are available}, while computing $l$ and $d$ requires 1 and 3 operations per pixel, respectively.

### 2015-06-22 paper.tex
minor edits
edits di Claudio

@@
 Background estimates and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while the number of SURF~\cite{bay2006surf} keypoints is used to detect blurring.
-A buffer of recent frames rather than an explicit background is used~\cite{ribnick2006real}.
-All the above solutions, including~\cite{harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.
+A buffer of recent frames rather than an explicit background can be used as in~\cite{ribnick2006real}.
+All the above solutions, including~\cite{harasse2004automated,kryjak2012fpga}, are computationally demanding and are not viable options for low-power cameras.
 
-In this work we expressly target low-power and ultra-low-power smart cameras, like \emph{SecSoC} (Security System on Chip), an innovative prototype based on a cluster of ReISC (Reduced Energy Instruction Set Computer) cores, designed and produced by STMicroelectronics.
+In this work, we expressly target low-power and ultra-low-power smart cameras, like \emph{SecSoC} (Security System on Chip), an innovative prototype based on a cluster of ReISC (Reduced Energy Instruction Set Computer) cores, designed and produced by STMicroelectronics.
 SecSoC is a battery-powered device, characterized by a constrained computational power (clock rates 82.5 MHz at 1.2V, and sub-1MHz at 0.6V) and reduced memory (1.25 MB).
@@
 %
-Let $z_t$ be the frame acquired at time $t$
+Let $z_t$ be frame acquired at time $t$
 \begin{equation}
@@
 We assume that $T_0$ tampering-free frames are provided for training.
+For simplicity, we consider grayscale frames: extensions to color images are straightforward.
 
@@
 \includegraphics[width=1\linewidth]{Immagini/sequenze}
-\caption{Examples of blurring and displacement synthetically introduced: \textbf{(a)} the 3-rd frame is blurred by a convolution against a Gaussian filter; \textbf{(b)} the 3-rd frame has been displaced by cropping a different region in the original image.}
+\caption{Examples of synthetically generated blurring (3-rd frame in \textbf{(a)}) and camera displacement (3-rd frame in \textbf{(b)}).}
 \label{fig:sequences}
@@
 Algorithm~\ref{alg:DISPL} presents the proposed tampering detection, which relies on simple indicators such as the average intensity (the \emph{luma}, denoted by $l$), the average frame difference (denoted by $d$) and the average norm of the gradient (denoted by $g$).
-The first two are meant to detect camera displacements, which would substantially change the image content, while the latter the blurring, which would cause a drop in the high-frequency content.
-As anticipated in Section~\ref{sec:introduction}, during the initial configuration the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$.
+The first two are meant to detect camera displacements, which would substantially change the image content, while the latter the blurring, which would cause a drop of the energy in the high-frequencies.
+As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$.
 The employed segmentation is detailed in Section~\ref{subsec:Segmentation}.
@@
 We can simultaneously monitor both luma and frame difference, even though computing $d_k$ requires to store the previous frame and this also depends on memory availability of the device.
-In what follows, including Algorithm~\ref{alg:DISPL}, we consider only the luma $l$ but the same procedures apply to the frame difference $d$ as well.
+In what follows, including Algorithm~\ref{alg:DISPL}, we consider only the luma $l$, but the same procedures apply to the frame difference $d$ as well.
 The average gradient norm is instead computed over the whole image
@@
 \includegraphics[width=1\linewidth]{Immagini/indicatori}
-\caption{Example of camera displacement. 
+\caption{Example of camera displacements. 
 Left plots report the luma values, while right plots depict the detrended luma: 
@@
 Red dots indicate a single displaced frame.
-Values in the highlighted areas refer to the frames depicted on top of the fugure.
-Displacement yields a peak in $l$ and two outliers in the sequence of detrended indicators $\partial l$, which can be clearly detected. }
+Values in the highlighted areas refer to the frames depicted on top of the figure.
+Displacement yields a peak in $l$ and two outliers in the sequence of detrended indicators $\partial l$, which can be clearly detected.}
 \label{fig:indicatori} 
@@
 %Tampering changes the sequence of indicators:
-Figure~\ref{fig:indicatori}(a) shows how a camera displacement affects the luma.
-First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) and that this change is also visible in the indicators $l_k(t)$ for the regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)).
+Figure~\ref{fig:indicatori} shows how a camera displacement affects the luma.
+First of all, we observe that a displaced frame changes $l(t)$ (introducing a peak in Figure~\ref{fig:indicatori}(a)) and that this change is also visible in the indicators $l_k(t)$ for the regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)).
 Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here.
@@
 			\lnl{DISPL-Test11} \If{$\partial g(t) < -\gamma_g \sigma_{g} \vee \partial l(t) > \gamma_g \sigma_{g} $}{
-					\lnl{DISPL-Test9} raise an alert: tampering (probably blurring) in $z_t$ \\
+					\lnl{DISPL-Test9} raise a blurring alert in $z_t$ \\
 									}
@@
 		\lnl{DISPL-Test8} \If{$n_l\geq \Gamma_l$}{
-			\lnl{DISPL-Test9} raise an alert: tampering (probably displacement or occlusion) in $z_t$ \\
+			\lnl{DISPL-Test9} raise a camera displacement alert in $z_t$ \\
 		}
@@
 \textbf{(b)}, ``One region - Algorithm 1'' (AUC $98.4\%$) is outperformed by ``Total'' (AUC $99.13\%$).
-To highlight differences, the ROC curves in \textbf{(a)} and \textbf{(c)} are drawn in a different TPR, FPR interval.}
+To highlight differences, the ROC curves in \textbf{(a)} and \textbf{(c)} are drawn over a smaller FPR, TPR range.}
 %\caption{\textbf{(a)} ROC curve for displacement detection considering Frame Difference (FD) as indicator.
@@
 
-Ongoing work concerns investigations towards other types of tampering, like degradations of the imaging sensor, as well as the integration of sequential monitoring schemes~\cite{alippi2010detecting} to detect subtle tampering that persists over time.
+Ongoing work concerns approaching other types of tampering, such as degradations of the imaging sensor, and the integration of sequential monitoring schemes~\cite{alippi2010detecting} to detect subtle tampering that persists over time.
 Moreover, we will investigate the use of superpixels methods~\cite{Susstrunk2012} to perform the preliminary scene segmentation, by including suitable temporal information to the feature vectors.

### 2015-06-22 paper.tex
modifiche di Claudio

@@
 	\SetNlSty{small}{}{.}
-	\textbf{Input}: $\gamma_l, \gamma_g$, $\{z_t, t = 1, \dots, T_{0}\}$, $\{R_k, \ k=1,\dots,K\}$ \\
-	\textbf{Configuration}:\\
+	\textbf{Input}: $\gamma_l, \gamma_g, \Gamma_l$, training frames $\{z_t, t = 1, \dots, T_{0}\}$, regions $\{R_k, \ k=1,\dots,K\}$ \\
+	\textbf{Training phase}:\\
 	\lnl{DISPL-Tr1} Compute $\partial l_k(t)$ and $\partial g(t), t = 1, \cdots, T_o$\\ 
@@
 The threshold $\Gamma_l$ together with $\gamma_l$ determine the displacement-detection promptness.
-The extreme cases are $\Gamma_l = 1$, where it is enough that a single region fires an outlier to raise a camera-displacement alert, and $\Gamma_l = K-1$ where all but one region\footnote{It is better to exclude a region since, for instance, the sky-region typically does not change when the camera is displaced, either horizontally or vertically} have to simultaneously fire an outlier to raise the alert.
+The extreme configurations correspond to $\Gamma_l = 1$, where it is sufficient that a single region fires an outlier to raise a camera-displacement alert, and $\Gamma_l = K-1$ where all but one region\footnote{It is better to exclude a region since, for instance, the sky-region typically does not change when the camera is displaced, either horizontally or vertically} have to simultaneously fire an outlier to raise an alert.
 
@@
 \end{equation}
-In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and column of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the mean and standard deviation of the intensity at $x$, computed over time, and $\overline{g}(x)$ and $\sigma_g(x)$ the mean of gradient norm and its standard deviation, computed overtime.
+In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and column of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the mean and standard deviation of the intensity at $x$, computed over time, and $\overline{g}(x)$ and $\sigma_g(x)$ the mean of gradient norm and its standard deviation, computed over time.
 These feature vectors are meant to cluster pixels in regions having, over training frames, similar spatial appearance and temporal behavior.

### 2015-06-22 paper.tex
minor edits nell'intro e didascalia 4

@@
 We address the detection of camera blurring and displacement (Section \ref{sec:probForm}), and propose an algorithm (Section~\ref{sec:propSol}) that relies on indicators that can be easily computed in low-power smart cameras.
-We monitor the average image intensity or frame difference (FD) (to detect camera displacement), and the average gradient norm (to detect blurring), and consider any frame yielding an outlier as tampered.
+We monitor the average image intensity or frame difference (to detect camera displacement), and the average gradient norm (to detect blurring), and detect tampered frames by analyzing outliers in these indicators.
 In particular, we show that separately monitoring these indicators over different regions of the image can substantially improve the displacement-detection performance.
-Image regions are defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t.
-monitoring the whole image (Section \ref{subsec:coputationalComplexity}).
+Image regions are defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead with respect to monitoring the whole image (Section \ref{subsec:coputationalComplexity}).
 Our algorithm thus represents a prompt trigger, to be possibly combined with other sequential monitoring techniques.
@@
 
-For each $z_t$, we compute the luma and frame difference separately on each of the $K$ regions,
+For each $z_t$, we compute the luma and frame difference (FD) separately on each of the $K$ regions,
 \begin{eqnarray}
@@
 
-We recorded $8$ sequences from webcams monitoring different urban areas, yielding overall $24410$ frames.
+We recorded $8$ sequences from webcams monitoring different urban areas, yielding overall $12200$ frames.
 From each frame we have cropped the central area, removing the 50 top-most and bottom-most rows, and the 50 left-most and right-most columns\footnote{Sequences can be provided upon request.}.
-We have synthetically introduced $5\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and coulumns (as in Figure \ref{fig:sequences}(b)).
+We have synthetically introduced $10\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and coulumns (as in Figure \ref{fig:sequences}(b)).
 Blurring  (as in Figure \ref{fig:sequences}(a)) was simulated by convolution against a Gaussian kernel, with $1\leq\sigma\leq5$ randomly defined.
@@
 
-A typical figure of merit for assessing detection performance is the Receiver-Operating Characteristic (ROC) curve, where each point corresponds to a pair (FPR$_\gamma$, TPR$_\gamma$) defined as
+A typical figure of merit for assessing detection performance is the Receiver Operating Characteristic (ROC) curve, where each point corresponds to a pair (FPR$_\gamma$, TPR$_\gamma$) defined as
 %
@@
 This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.%\gi{aggiungere commento su K-1 vornoi che va meglio?}
-
+x
 \begin{figure}[t]
@@
 \caption{\textbf{(a)} ROC curves relative to displacement detection based on FD: the AUC obtained by 
-``One region - Algorithm 1'' is $99.89\%$, while the AUC obtained by ``Total'' (on the whole 
-image) is $99.65\%$.
-\textbf{(b)} ROC curves relative to displacement detection based on Luma: the AUC obtained by 
-``One region - Algorithm 1'' is $98.44\%$, while the AUC obtained by ``Total'' (on the whole 
-image) is $84.07\%$.
-\textbf{(c)} ROC curves relative to blurring detection based on Gradient: as opposed to \textbf{(a)} and 
-\textbf{(b)}, ``One region - Algorithm 1'' (AUC $98.4\%$) is outperformed by ``Total'' (AUC $99.13\%$).
+``Algorithm 1 - One region'' is $99.89\%$, while the AUC obtained by on the whole 
+image is $99.65\%$.
+\textbf{(b)} ROC curves relative to displacement detection based on luma: the AUC obtained by 
+``Algorithm 1 - One region'' is $98.44\%$, while the AUC obtained on the whole 
+image is $84.07\%$.
+\textbf{(c)} ROC curves relative to blurring detection based on gradient: as opposed to \textbf{(a)} and 
+\textbf{(b)}, ``Algorithm 1 - One region'' (AUC $98.4\%$) is outperformed by monitoring the whole image, AUC = $99.13\%$.
 To highlight differences, the ROC curves in \textbf{(a)} and \textbf{(c)} are drawn over a smaller FPR, TPR range.}

### 2015-06-22 paper.tex
final edits

@@
 %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
-\keywords{tampering detection, smart cameras, displacement detection, blurring detection}
+\keywords{tampering detection, smart cameras, displacement detection, blurring detection, low-power cameras, low-frame rate.}
 \end{abstract}
@@
 
-Algorithm~\ref{alg:DISPL} presents the proposed tampering detection, which relies on simple indicators such as the average intensity (the \emph{luma}, denoted by $l$), the average frame difference (denoted by $d$) and the average norm of the gradient (denoted by $g$).
-The first two are meant to detect camera displacements, which would substantially change the image content, while the latter the blurring, which would cause a drop of the energy in the high-frequencies.
+Algorithm~\ref{alg:DISPL} presents the proposed tampering detection, which relies on simple indicators such as the average intensity (the \emph{luma}, denoted by $l$), the average frame difference (\emph{FD}, denoted by $d$) and the average norm of the gradient (denoted by $g$).
+The first two are meant to detect camera displacements, which would substantially change the image content, while the latter the blurring, which would attenuate the high frequencies of the image.
 As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$.
@@
 
-For each $z_t$, we compute the luma and frame difference (FD) separately on each of the $K$ regions,
+For each $z_t$, we compute the luma and frame difference separately on each of the $K$ regions,
 \begin{eqnarray}
@@
 Values in the highlighted areas refer to the frames depicted on top of the figure.
-Displacement yields a peak in $l$ and two outliers in the sequence of detrended indicators $\partial l$, which can be clearly detected.}
+The highlighted displacement yields a small peak in $l(t)$ and two outliers in the sequence of detrended indicators $\partial l(t)$, which can be clearly detected.}
 \label{fig:indicatori} 
@@
 It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, a single tampered frame yields two outliers (as shown in Figure~\ref{fig:indicatori}) and only the first and the last of a sequence of consecutive tampered frames yield outliers in the detrended indicators.
-This is the reason why detrended indicators are monitored in a one-shot manner, targeting the detection of the first tampered-frame.
-On the one hand, this one-shot solution can provide prompt detections, which is particularly important at the low frame-rates we consider, on the other hand, the persistence of tampering is disregarded.
-To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, possibly combined with the proposed one-shot monitoring.
+This is the reason why detrended indicators are monitored in a \emph{one-shot} manner, targeting the detection of the first tampered-frame.
+On the one hand, this one-shot monitoring can provide prompt detections, which is particularly important at the low frame-rates we consider, on the other hand, the persistence of tampering is disregarded.
+To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, possibly combined with Algorithm~\ref{alg:DISPL}.
 %
@@
 Feature vectors~\eqref{eq:featureVector} over the whole image are clustered by a weighted k-means~\cite{kottke1994motion} that scales each component of the Euclidean distance between a feature vector and a cluster centroid of a weight that is inversely proportional to the standard deviation over the cluster.
-This scaling compensates the fact that feature vector components might span very different ranges.
+This scaling compensates the fact that the components of the feature vector might span very different ranges.
 The number of clusters is defined by testing several values and then choosing the best solution according to the Calinski-Harabasz criterion~\cite{calinski1974dendrite}.
@@
 To this purpose, we show that Algorithm~\ref{alg:DISPL} detects camera displacements better than monitoring indicators on the whole image (\emph{Whole Image} in Figure~\ref{fig:ROC}).
-We also show that it is important to define regions considering the scene content, since operating on $K$ regions obtained by clustering $[r(x), c(x)]$ leads to performance loss (\emph{Voronoi} in Figure~\ref{fig:ROC}).
+We also show that it is important to adapt regions to the image content, since operating on $K$ regions obtained by clustering $[r(x), c(x)]$ leads to performance loss (\emph{Voronoi} in Figure~\ref{fig:ROC}).
 For both Voronoi and regions defined as in Section~\ref{subsec:Segmentation}, we consider the two extreme configurations where alerts are raised at the first region firing an outlier or when $K-1$ regions simultaneously fire outliers.
@@
 We recorded $8$ sequences from webcams monitoring different urban areas, yielding overall $12200$ frames.
-From each frame we have cropped the central area, removing the 50 top-most and bottom-most rows, and the 50 left-most and right-most columns\footnote{Sequences can be provided upon request.}.
-We have synthetically introduced $10\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and coulumns (as in Figure \ref{fig:sequences}(b)).
-Blurring  (as in Figure \ref{fig:sequences}(a)) was simulated by convolution against a Gaussian kernel, with $1\leq\sigma\leq5$ randomly defined.
+From each frame, we have cropped the central area, removing the 50 top-most and bottom-most rows, and the 50 left-most and right-most columns\footnote{Sequences can be provided upon request.}.
+We have synthetically introduced $10\%$ of tampered frames: displacements have been simulated by moving the central cropping area of a random shift having magnitude between 20 and 50 rows and columns (as in Figure \ref{fig:sequences}(b)), while blurring (as in Figure \ref{fig:sequences}(a)) was simulated by convolution against a Gaussian kernel, with standard deviation randomly defined in the range $[1,5]$.
 %\gi{Adriano: image sizes? blur kernel std? quanti pixel il displacement?}
@@
 \subsubsection{Discussion}
-ROC curves in Figure~\ref{fig:ROC}(a)-(b) and the Area Under the Curve (AUC) values reported in the caption confirm that camera displacements can be better detected by monitoring FD than luma, and that it is convenient to define regions segmenting the scene content.
+ROC curves in Figure~\ref{fig:ROC}(a)-(b) and the Area Under the Curve (AUC) values reported in the caption confirm that camera displacements can be better detected by monitoring FD than luma, and that it is convenient to define regions by segmenting the scene.
 As an example, if we set $\gamma_d = 15.6$, Algorithm~\ref{alg:DISPL} operates at $\text{FPR} = 1\%$ and detects nearly all tampering since $\text{TPR} = 99.92\%$.
-In contrast, to achieve $\text{FPR} = 1\%$ when monitoring the whole image, FPR drops to $91.67\%$ ($\gamma_d = 6.5$), and when considering Voronoi to $\text{TPR} = 89.05\%$ ($\gamma_d = 18.64$).
-When monitoring $l$, Algorithm~\ref{alg:DISPL} achieves TPR~=~$73.98\%$ at $\text{FPR}=1\%$  ($\gamma_d=5.8$), while monitoring the whole image is quite ineffective ($\text{TPR} = 17.85\%$, $\gamma_d=3.7$), and similarly perform Voronoi's regions ($\text{FPR} = 53.02\%$, $\gamma_d=6$).
-We remark that, in the considered low-power scenario, it is important to operate at low FPR, given the large number of frames that smart cameras have to process over time and the cost of data-transmission in terms of battery power.
+In contrast, to operate at $\text{FPR} = 1\%$ when monitoring the whole image or Voronoi regions we obtain $\text{TPR} = 91.67\%$ ($\gamma_d = 6.5$) and $\text{TPR} = 89.05\%$ ($\gamma_d = 18.64$), respectively.
+When monitoring $l$, Algorithm~\ref{alg:DISPL} achieves TPR~=~$73.98\%$ at $\text{FPR}=1\%$  ($\gamma_d=5.8$), while monitoring the whole image is quite ineffective ($\text{TPR} = 17.85\%$, $\gamma_d=3.7$), and similarly perform Voronoi's regions ($\text{TPR} = 53.02\%$, $\gamma_d=6$).
+We remark that, in the considered low-power scenario, it is important to operate at low FPR to prevent useless data transmission (that would reduce the battery time-life) and the activation of eventual countermeasures.
 As far as the values of $\Gamma_d$ and $\Gamma_l$ are concerned, the most effective configuration consists in raising an alert as soon as a single region fires a very clear outlier, which implies setting quite wide intervals~\eqref{eq:confidenceRegions}.
 This option is also preferable because it enables the detection of partial occlusions of the scene.
+We observe that displacements typically change all the Voronoi regions, and this is why Voronoi outperforms Algorithm~\ref{alg:DISPL} in the configuration $K-1$ in both luma and FD.
 
-%\gi{Aggiornare con i nuovi dati}
-%\emph{Considering $d$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $0.4\%$ of false alarms, with $\gamma = 24.7$, which is less than one fifth of the whole image solution ($2.2\%$ with $\gamma = 4.6$) and the Voronoi solution ($2.3\%$ with $\gamma = 12.9$).
-%On the other side, considering $l$ as indicator, we observe that to achieve a recall of $98\%$, Algorithm~\ref{alg:DISPL} operates at $13\%$ of false alarms, with $\gamma = 2.5$, which is about one sixth of the whole image solution ($75\%$ with $\gamma = 0.1$) and about three fifty than the Voronoi solution ($22\%$ with $\gamma = 1.7 $).
-%Having few false alarms is important in a low-power context, considering the high number of frames that have to be processed.
-Most importantly, false alerts are halved basically at no computational overhead.} \gi{Aggiornare con i nuovi dati}
 Figure~\ref{fig:ROC}(c) confirms that blurring is more effectively detected by monitoring the whole image at once.
 This is probably due to the fact that, typically, regions do not include the most prominent edge of the scene, which are indeed the most informative parts to detect blurring.
-This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.%\gi{aggiungere commento su K-1 vornoi che va meglio?}
-x
+% This is the reason why, in Algorithm~\ref{alg:DISPL}, $g$ is monitored on the whole image.%\gi{aggiungere commento su K-1 vornoi che va meglio?}
+
 \begin{figure}[t]
@@
 Ongoing work concerns approaching other types of tampering, such as degradations of the imaging sensor, and the integration of sequential monitoring schemes~\cite{alippi2010detecting} to detect subtle tampering that persists over time.
-Moreover, we will investigate the use of superpixels methods~\cite{Susstrunk2012} to perform the preliminary scene segmentation, by including suitable temporal information to the feature vectors.
+Moreover, we will investigate the use of superpixels methods~\cite{Susstrunk2012} to segment the scene, by including suitable temporal information to the feature vectors.
 
